### 2022-06-13 outputs/paper/paper.qmd
Create paper.qmd

---
title: "paper.qmd"
author: "Sarah Mansoor"
format: pdf
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

### 2022-06-13 outputs/paper/paper.qmd
Update paper.qmd

@@
 ---
-title: "paper.qmd"
+title: "Disparities in Climate-Induced Health Outcomes in the Greater Toronto Area."
 author: "Sarah Mansoor"
+thanks: "Code and data are available at: https://github.com/sarahmansoorr/SUDS2022"
+date: "`r format(Sys.time(), '%d %B %Y')`"
 format: pdf
 editor: visual
+bstract: ""
 ---
 
-## Quarto
+## Introduction
 
-Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.
+## Data
 
-## Running Code
+## Visualizing the Data and the Implications
 
-When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:
-
-```{r}
-1 + 1
-```
-
-You can add options to executable code like this
-
-```{r}
-#| echo: false
-2 * 2
-```
-
-The `echo: false` option disables the printing of code (only output is displayed).
+## Next Steps

### 2022-06-13 outputs/paper/paper.qmd
Update paper.qmd

@@
 editor: visual
-bstract: ""
+abstract: ""
+toc: true
+bibliography: references.bib
 ---
@@
 
+#### Data Collection
+
+#### Variables of Interest
+
+#### Data Processing
+
+#### Population, Frame or Sample
+
 ## Visualizing the Data and the Implications
 
+#### Visual 1
+
+#### Implications 1
+
+#### Visual 2
+
+#### Implications 2
+
+#### Visuals 3
+
+#### Implications 3
+
+## Limitations
+
+## Results
+
 ## Next Steps
+
+## Conclusion

### 2022-06-14 outputs/paper/paper.qmd
Update paper.qmd

@@
 ---
+#### Preamble ####
 title: "Disparities in Climate-Induced Health Outcomes in the Greater Toronto Area."
 author: "Sarah Mansoor"
+email: "sarah.mansoor@mail.utoronto.ca"
 thanks: "Code and data are available at: https://github.com/sarahmansoorr/SUDS2022"

### 2022-06-14 scripts/00-download_data.R
Downloading temperature data 2002-2022
Jan 2002 - May 2022

@@ -14,3 +14,30 @@
 library(tidyverse)
 library(dplyr)
 
+#### Load datasets ####
+temp_jan_may_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2022_P1D")
+temp_jan_dec_2021 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2021_P1D.csv")
+temp_jan_dec_2020 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2020_P1D.csv")
+temp_jan_dec_2019 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2019_P1D.csv")
+temp_jan_dec_2018 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2018_P1D.csv")
+temp_jan_dec_2017 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2017_P1D.csv")
+temp_jan_dec_2016 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2016_P1D.csv")
+temp_jan_dec_2015 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2015_P1D.csv")
+temp_jan_dec_2014 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2014_P1D.csv")
+temp_jan_dec_2013 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2013_P1D.csv")
+temp_jan_may_2012 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2012_P1D.csv")
+temp_jan_dec_2011 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2011_P1D.csv")
+temp_jan_dec_2010 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2010_P1D.csv")
[…]

### 2022-06-14 scripts/00-download_data.R
Download and rename files
Change file names to a more readable name

@@ -15,7 +15,7 @@ library(tidyverse)
 library(dplyr)
 
 #### Load datasets ####
-temp_jan_may_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2022_P1D")
+temp_jan_may_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2022_P1D.csv")
 temp_jan_dec_2021 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2021_P1D.csv")
 temp_jan_dec_2020 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2020_P1D.csv")
 temp_jan_dec_2019 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2019_P1D.csv")
@@ -37,7 +37,28 @@ temp_jan_dec_2004 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_O
 temp_jan_dec_2003 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2003_P1D.csv")
 temp_jan_dec_2002 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2002_P1D.csv")
 
-
+# Rename files to be more readable
+file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2022_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_may_2022.csv")
+file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2021_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_dec_2021.csv")
+file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2020_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_dec_2020.csv")
+file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2019_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_dec_2019.csv")
[…]

### 2022-06-15 scripts/00-download_data.R
Cleaning temperature data

@@ -48,7 +48,7 @@ file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2016_P1D
 file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2015_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_dec_2015.csv")
 file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2014_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_dec_2014.csv")
 file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2013_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_dec_2013.csv")
-file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2012_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_may_2012.csv")
+file.rename("~/Desktop/SUDS2022/inputs/data/temp_jan_may_2012.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_dec_2012.csv")
 file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2011_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_dec_2011.csv")
 file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2010_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_dec_2010.csv")
 file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2009_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_dec_2009.csv")

### 2022-06-15 scripts/01-data_cleaning.R

@@ -15,3 +15,37 @@ library(tidyverse)
 library(dplyr)
 library(ggplot2)
 library(janitor)
+
+#### Cleaning Data ####
+
+# Combine 2002 - 2022 data
+
+temp_2002_2022 <- rbind(temp_jan_dec_2002, temp_jan_dec_2003, temp_jan_dec_2004, 
+                        temp_jan_dec_2005, temp_jan_dec_2006, temp_jan_dec_2007,
+                        temp_jan_dec_2008, temp_jan_dec_2009, temp_jan_dec_2010,
+                        temp_jan_dec_2011, temp_jan_dec_2012, temp_jan_dec_2013, 
+                        temp_jan_dec_2014, temp_jan_dec_2015, temp_jan_dec_2016,
+                        temp_jan_dec_2017, temp_jan_dec_2018, temp_jan_dec_2019,
+                        temp_jan_dec_2020, temp_jan_dec_2021, temp_jan_may_2022)
+temp_2002_2022_5yr_interval <- rbind(temp_jan_dec_2002, temp_jan_dec_2007, 
+                                     temp_jan_dec_2012, temp_jan_dec_2017, 
+                                     temp_jan_may_2022)
+temp_2002_2022_10yr_interval <- rbind(temp_jan_dec_2002,
+                                     temp_jan_dec_2012,
+                                     temp_jan_may_2022)
+
+### 2002 ###
+# change Date.Time from chr to Date type
+temp_2002_2022 <- temp_2002_2022 %>% select(Station.Name, Date.Time, Year, 
+                                            Month, Day, Max.Temp...C., 
+                                            Min.Temp...C., Mean.Temp...C., 
+                                            Heat.Deg.Days...C., Cool.Deg.Days...C., 
+                                            Total.Rain..mm., Total.Snow..cm.,
+                                            Total.Precip..mm.)
+
+my_column_names <- c("station.name", "date", "year", "month", "day", "max.temp", "min.temp", 
+                     "mean.temp", "heat.deg.days", "cool.deg.days", "total.rain.mm", 
+                     "total.snow.cm", "total.percip.mm")
+
+colnames(temp_2002_2022) <- my_column_names
+temp_2002_2022 <- clean_names(temp_2002_2022)

### 2022-06-15 scripts/01-data_cleaning.R
Update 01-data_cleaning.R

@@ -19,7 +19,6 @@ library(janitor)
 #### Cleaning Data ####
 
 # Combine 2002 - 2022 data
-
 temp_2002_2022 <- rbind(temp_jan_dec_2002, temp_jan_dec_2003, temp_jan_dec_2004, 
                         temp_jan_dec_2005, temp_jan_dec_2006, temp_jan_dec_2007,
                         temp_jan_dec_2008, temp_jan_dec_2009, temp_jan_dec_2010,
@@ -27,15 +26,15 @@ temp_2002_2022 <- rbind(temp_jan_dec_2002, temp_jan_dec_2003, temp_jan_dec_2004,
                         temp_jan_dec_2014, temp_jan_dec_2015, temp_jan_dec_2016,
                         temp_jan_dec_2017, temp_jan_dec_2018, temp_jan_dec_2019,
                         temp_jan_dec_2020, temp_jan_dec_2021, temp_jan_may_2022)
+
 temp_2002_2022_5yr_interval <- rbind(temp_jan_dec_2002, temp_jan_dec_2007, 
                                      temp_jan_dec_2012, temp_jan_dec_2017, 
                                      temp_jan_may_2022)
+
 temp_2002_2022_10yr_interval <- rbind(temp_jan_dec_2002,
                                      temp_jan_dec_2012,
                                      temp_jan_may_2022)
 
-### 2002 ###
-# change Date.Time from chr to Date type
 temp_2002_2022 <- temp_2002_2022 %>% select(Station.Name, Date.Time, Year, 
                                             Month, Day, Max.Temp...C., 
                                             Min.Temp...C., Mean.Temp...C., 
@@ -49,3 +48,4 @@ my_column_names <- c("station.name", "date", "year", "month", "day", "max.temp",
 
 colnames(temp_2002_2022) <- my_column_names
 temp_2002_2022 <- clean_names(temp_2002_2022)
+

### 2022-06-15 scripts/01-data_cleaning.R
Data cleaning

@@ -27,6 +27,12 @@ temp_2002_2022 <- rbind(temp_jan_dec_2002, temp_jan_dec_2003, temp_jan_dec_2004,
                         temp_jan_dec_2017, temp_jan_dec_2018, temp_jan_dec_2019,
                         temp_jan_dec_2020, temp_jan_dec_2021, temp_jan_may_2022)
 
+temp_2012_2022 <- rbind(temp_jan_dec_2012, temp_jan_dec_2013, 
+                        temp_jan_dec_2014, temp_jan_dec_2015, temp_jan_dec_2016,
+                        temp_jan_dec_2017, temp_jan_dec_2018, temp_jan_dec_2019,
+                        temp_jan_dec_2020, temp_jan_dec_2021, temp_jan_may_2022)
+
+
 temp_2002_2022_5yr_interval <- rbind(temp_jan_dec_2002, temp_jan_dec_2007, 
                                      temp_jan_dec_2012, temp_jan_dec_2017, 
                                      temp_jan_may_2022)

### 2022-06-15 scripts/plots.R

@@ -1,0 +1,52 @@
+#### ggplots ####
+library(ggplot2)
+
+ggplot(temp_2002_2022, aes(x=year, y=max_temp, color=month)) + geom_point()
+
+ggplot(temp_2002_2022, aes(x=year, y=max_temp, color=month)) + geom_line()
+
+ggplot(temp_2002_2022, aes(x=year, y=min_temp, color=month)) + geom_point()
+
+ggplot(temp_2002_2022, aes(x=year, y=mean_temp, color=month)) + geom_point()
+
+ggplot(temp_2002_2022, aes(x=month, y=max_temp, color=year)) + geom_point()
+
+ggplot(temp_2002_2022, aes(x=month, y=min_temp, color=year)) + geom_point()
+
+ggplot(temp_2002_2022, aes(x=month, y=mean_temp, color=year)) + geom_point()
+
+
+#### tables ####
+library(knitr)
+install.packages("kableExtra")
+library(kableExtra)
+
+max_temp_2002_2022 <- temp_2002_2022 %>%
+  select(year, month, day, max_temp)
+max_temp_2002_2022 %>%
+  slice(c(162, 393, 745, 1318, 1819, 2115, 3222, 4222, 5222, 6333)) %>%
+  kable(
+    caption = "Extracting rows from the Temperature data from 2002 - 2022 for Maximum Temperature",
+    col.names=c("Year", "Month", "Day", "Max Temp"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
+
+min_temp_2002_2022 <- temp_2002_2022 %>%
+  select(year, month, day, min_temp)
+min_temp_2002_2022 %>%
+  slice(c(162, 393, 745, 1318, 1819, 2115, 3222, 4222, 5222, 6333)) %>%
+  kable(
+    caption = "Extracting rows from the Temperature data from 2002 - 2022 for Minimum Temperature",
+    col.names=c("Year", "Month", "Day", "Min Temp"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
+
+mean_temp_2002_2022 <- temp_2002_2022 %>%
+  select(year, month, day, mean_temp)
+mean_temp_2002_2022 %>%
+  slice(c(162, 393, 745, 1318, 1819, 2115, 3222, 4222, 5222, 6333)) %>%
+  kable(
[…]

### 2022-06-16 scripts/01-data_cleaning.R
Adding column for the month names

@@ -55,3 +55,21 @@ my_column_names <- c("station.name", "date", "year", "month", "day", "max.temp",
 colnames(temp_2002_2022) <- my_column_names
 temp_2002_2022 <- clean_names(temp_2002_2022)
 
+
+### Add column for Month Names
+temp_2002_2022 <- temp_2002_2022 %>%
+  mutate(month_name = case_when(
+    month == 1 ~ "Jan",
+    month == 2 ~ "Feb", 
+    month == 3 ~ "Mar", 
+    month == 4 ~ "Apr",
+    month == 5 ~ "May", 
+    month == 6 ~ "Jun",
+    month == 7 ~ "Jul",
+    month == 8 ~ "Aug", 
+    month == 9 ~ "Sep",
+    month == 10 ~ "Oct",
+    month == 11 ~ "Nov", 
+    month == 12 ~ "Dec"
+  ))
+

### 2022-06-16 scripts/01-data_cleaning.R
Cleaning to drop NA's in max and min temps

@@ -18,6 +18,29 @@ library(janitor)
 
 #### Cleaning Data ####
 
+### Attach data
+attach(temp_jan_dec_2002)
+attach(temp_jan_dec_2003)
+attach(temp_jan_dec_2004)
+attach(temp_jan_dec_2005)
+attach(temp_jan_dec_2006)
+attach(temp_jan_dec_2007)
+attach(temp_jan_dec_2008)
+attach(temp_jan_dec_2009)
+attach(temp_jan_dec_2010)
+attach(temp_jan_dec_2011)
+attach(temp_jan_dec_2012)
+attach(temp_jan_dec_2013)
+attach(temp_jan_dec_2014)
+attach(temp_jan_dec_2015)
+attach(temp_jan_dec_2016)
+attach(temp_jan_dec_2017)
+attach(temp_jan_dec_2018)
+attach(temp_jan_dec_2019)
+attach(temp_jan_dec_2020)
+attach(temp_jan_dec_2021)
+attach(temp_jan_may_2022)
+
 # Combine 2002 - 2022 data
 temp_2002_2022 <- rbind(temp_jan_dec_2002, temp_jan_dec_2003, temp_jan_dec_2004, 
                         temp_jan_dec_2005, temp_jan_dec_2006, temp_jan_dec_2007,
@@ -45,12 +68,10 @@ temp_2002_2022 <- temp_2002_2022 %>% select(Station.Name, Date.Time, Year,
                                             Month, Day, Max.Temp...C., 
                                             Min.Temp...C., Mean.Temp...C., 
                                             Heat.Deg.Days...C., Cool.Deg.Days...C., 
-                                            Total.Rain..mm., Total.Snow..cm.,
                                             Total.Precip..mm.)
 
 my_column_names <- c("station.name", "date", "year", "month", "day", "max.temp", "min.temp", 
-                     "mean.temp", "heat.deg.days", "cool.deg.days", "total.rain.mm", 
-                     "total.snow.cm", "total.percip.mm")
+                     "mean.temp", "heat.deg.days", "cool.deg.days", "total.percip.mm")
 
 colnames(temp_2002_2022) <- my_column_names
 temp_2002_2022 <- clean_names(temp_2002_2022)
@@ -72,4 +93,11 @@ temp_2002_2022 <- temp_2002_2022 %>%
     month == 11 ~ "Nov", 
     month == 12 ~ "Dec"
   ))
+
[…]

### 2022-06-16 scripts/plots.R

@@ -1,7 +1,7 @@
 #### ggplots ####
 library(ggplot2)
 
-ggplot(temp_2002_2022, aes(x=year, y=max_temp, color=month)) + geom_point()
+ggplot(temp_2002_2022, aes(x=year, y=max_temp, color=month.name)) + geom_point()
 
 ggplot(temp_2002_2022, aes(x=year, y=max_temp, color=month)) + geom_line()
 

### 2022-06-16 scripts/01-data_cleaning.R
Update 01-data_cleaning.R

@@ -95,9 +95,8 @@ temp_2002_2022 <- temp_2002_2022 %>%
   ))
 
 ### Drop NA for max_temp
-df <- temp_2002_2022
-df <- df %>% drop_na(max_temp)
-df <- df %>% drop_na(min_temp)
-df <- df %>% drop_na(mean_temp)
-df <- df %>% drop_na(total_percip_mm)
+temp_2002_2022 <- temp_2002_2022 %>% drop_na(max_temp)
+temp_2002_2022 <- temp_2002_2022 %>% drop_na(min_temp)
+temp_2002_2022 <- temp_2002_2022 %>% drop_na(mean_temp)
+temp_2002_2022 <- temp_2002_2022 %>% drop_na(total_percip_mm)
 

### 2022-06-16 scripts/plots.R
Update plots.R

@@ -1,6 +1,8 @@
 #### ggplots ####
 library(ggplot2)
 
+ggplot(temp_2002_2022, aes(x=year, y=max_temp, color=month)) + geom_point()
+
 ggplot(temp_2002_2022, aes(x=year, y=max_temp, color=month.name)) + geom_point()
 
 ggplot(temp_2002_2022, aes(x=year, y=max_temp, color=month)) + geom_line()

### 2022-06-16 scripts/01-data_cleaning.R
Update 01-data_cleaning.R

@@ -67,11 +67,10 @@ temp_2002_2022_10yr_interval <- rbind(temp_jan_dec_2002,
 temp_2002_2022 <- temp_2002_2022 %>% select(Station.Name, Date.Time, Year, 
                                             Month, Day, Max.Temp...C., 
                                             Min.Temp...C., Mean.Temp...C., 
-                                            Heat.Deg.Days...C., Cool.Deg.Days...C., 
                                             Total.Precip..mm.)
 
 my_column_names <- c("station.name", "date", "year", "month", "day", "max.temp", "min.temp", 
-                     "mean.temp", "heat.deg.days", "cool.deg.days", "total.percip.mm")
+                     "mean.temp", "total.percip.mm")
 
 colnames(temp_2002_2022) <- my_column_names
 temp_2002_2022 <- clean_names(temp_2002_2022)

### 2022-06-20 scripts/00-download_data.R
Updates to analysis of temperature data

@@ -14,7 +14,7 @@
 library(tidyverse)
 library(dplyr)
 
-#### Load datasets ####
+#### Load temperature datasets ####
 temp_jan_may_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2022_P1D.csv")
 temp_jan_dec_2021 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2021_P1D.csv")
 temp_jan_dec_2020 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2020_P1D.csv")

### 2022-06-20 scripts/01-data_cleaning.R

@@ -99,3 +99,5 @@ temp_2002_2022 <- temp_2002_2022 %>% drop_na(min_temp)
 temp_2002_2022 <- temp_2002_2022 %>% drop_na(mean_temp)
 temp_2002_2022 <- temp_2002_2022 %>% drop_na(total_percip_mm)
 
+####### clarify in readme to use the cleaned data to download from github
+#### start a script for analysis that starts with clean data

### 2022-06-20 scripts/plots.R

@@ -86,8 +86,25 @@ ggplot(winter, aes(x=year, y=max_temp, color=month_name)) + geom_point()
 spring <- temp_2002_2022 %>%
   filter(month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul")
 
+ggplot(spring, aes(x=year, y=max_temp, color=month_name)) + geom_point()
+
 summer <- temp_2002_2022 %>%
   filter(month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")
 
+ggplot(summer, aes(x=year, y=max_temp, color=month_name)) + geom_point()
 
+
+## using if else
+winter_2002_2022 <- ifelse(temp_2002_2022$date < "")
+
+"Econometrics
+Check differences to a different city over time
+Freedom of information
+Uoft databases
+Causation models
+Regression models
+
+looking at rising temperatures and deaths
+looking at rising temperatures and lung diseases
+looking at rising temperatures and pollen and deaths"
 

### 2022-06-20 scripts/00-download_data.R
Download pollution data

@@ -61,4 +61,8 @@ file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2003_P1D
 file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2002_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_dec_2002.csv")
 
 
+#
+
+
+#### Load pollution datasets ####
 

### 2022-06-20 scripts/00-download_data.R
Downloading and cleaning Pollution data

@@ -13,29 +13,31 @@
 # Use R Projects, not setwd().
 library(tidyverse)
 library(dplyr)
+library(readr)
+
 
 #### Load temperature datasets ####
-temp_jan_may_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2022_P1D.csv")
-temp_jan_dec_2021 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2021_P1D.csv")
-temp_jan_dec_2020 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2020_P1D.csv")
-temp_jan_dec_2019 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2019_P1D.csv")
-temp_jan_dec_2018 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2018_P1D.csv")
-temp_jan_dec_2017 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2017_P1D.csv")
-temp_jan_dec_2016 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2016_P1D.csv")
-temp_jan_dec_2015 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2015_P1D.csv")
-temp_jan_dec_2014 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2014_P1D.csv")
-temp_jan_dec_2013 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2013_P1D.csv")
-temp_jan_may_2012 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2012_P1D.csv")
-temp_jan_dec_2011 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2011_P1D.csv")
-temp_jan_dec_2010 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2010_P1D.csv")
-temp_jan_dec_2009 <- read.csv("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2009_P1D.csv")
[…]

### 2022-06-20 scripts/01-data_cleaning.R

@@ -1,5 +1,5 @@
-#### Preamble ####
-# Purpose: Clean the data downloaded
+#### Preamble ----
+# Purpose: Clean the tempreature and pollution data 
 # Author: Sarah Mansoor
 # Date: May 2022
 # Contact: sarah.mansoor@mail.utoronto.ca
@@ -9,16 +9,15 @@
 # - Don't forget to gitignore it!
 
 
-#### Workspace setup ####
-# Use R Projects, not setwd().
+#### Workspace setup ----
 library(tidyverse)
 library(dplyr)
 library(ggplot2)
 library(janitor)
 
-#### Cleaning Data ####
+#### Cleaning Temperature Data ----
 
-### Attach data
+### Attach temperature data
 attach(temp_jan_dec_2002)
 attach(temp_jan_dec_2003)
 attach(temp_jan_dec_2004)
@@ -93,11 +92,108 @@ temp_2002_2022 <- temp_2002_2022 %>%
     month == 12 ~ "Dec"
   ))
 
-### Drop NA for max_temp
+### Drop NA
 temp_2002_2022 <- temp_2002_2022 %>% drop_na(max_temp)
 temp_2002_2022 <- temp_2002_2022 %>% drop_na(min_temp)
 temp_2002_2022 <- temp_2002_2022 %>% drop_na(mean_temp)
 temp_2002_2022 <- temp_2002_2022 %>% drop_na(total_percip_mm)
 
-####### clarify in readme to use the cleaned data to download from github
-#### start a script for analysis that starts with clean data
+
+
+
+#### Cleaning Pollution Data ----
+
+### Attach pollution data
+attach(pollution_2020)
+attach(pollution_2019)
+attach(pollution_2018)
+attach(pollution_2017)
+attach(pollution_2016)
+attach(pollution_2015)
+attach(pollution_2014)
+attach(pollution_2013)
+attach(pollution_2012)
+attach(pollution_2011)
+
+### Filter for Toronto and reduce to variables of interest and add year
+
+pollution_2020 <- pollution_2020 %>%
+  filter(City == "Toronto") %>% select(City, Pollutant, January, February, March, April, May, 
+                                       June, July, August, September, October, November, December, 
+                                       Average, StdDeviation) %>%
+  mutate(Year = 2020)
+
+pollution_2019 <- pollution_2019 %>%
+  filter(City == "Toronto") %>% select(City, Pollutant, January, February, March, April, May, 
+                                       June, July, August, September, October, November, December, 
[…]

### 2022-06-20 scripts/02-data_analysis.R

@@ -1,0 +1,12 @@
+#### Preamble ####
+# Purpose: Clean the tempreature and pollution data 
+# Author: Sarah Mansoor
+# Date: May 2022
+# Contact: sarah.mansoor@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+# - Need to have downloaded the ACS data and saved it to inputs/data
+# - Don't forget to gitignore it!
+
+
+#### Workspace setup ####

### 2022-06-20 scripts/plots.R

@@ -56,7 +56,7 @@ mean_temp_2002_2022 %>%
   kable_styling(latex_options = "HOLD_position", font_size = 10)
 
 
-### summarising data
+# summarising data ----
 summary_max_data <- temp_2002_2022 %>%
   group_by(month_name) %>%
   summarise(mean_max_temp = mean(max_temp))
@@ -108,3 +108,6 @@ looking at rising temperatures and deaths
 looking at rising temperatures and lung diseases
 looking at rising temperatures and pollen and deaths"
 
+# pollution data ----
+
+ggplot(pollution_2011_2020, aes(Year, January, color=Pollutant)) + geom_point()

### 2022-06-20 outputs/paper/paper.qmd
Notes in paper.qmd
Update paper.qmd

--- 2022-06-13 outputs/paper/paper.qmd
@@
 editor: visual
-abstract: ""
+abstract: "The growing impact of climate change in Canada demonstrates a clear and current threat to human health.
+These threats are becoming revealed in Canada as the country continues to face the consequences of warming almost at twice the global rate.
+In this report, I will explore how tempreture increases, lower air quality relate to a rise in threats to human healht in Toronto.
+Once the data for this research was collected, the data cleaning process begun which entailed of using packages in R like tidyverse, janitor, dplyr, ….
+Duplicate and irrelevant observations were removed, as well as unwanted outliers.
+The process was complete once the data was correct, consistent and usable.
+"
 toc: true
@@
 
+Why do we care?
+
+What is in this paper?
+
+Results?
+
 ## Data
+
+table of variables used?
 
@@
 
-## Visualizing the Data and the Implications
+## Visualizing the Data and the Implications (needs to be reworded)
+
+use other graphs beside boxplot. log scales. (dotpoint, barchart, etc)
+
+avoid drastic, dramatic, etc
 
@@
 
+need updated data?
+
 ## Conclusion

### 2022-06-21 scripts/01-data_cleaning.R
Updates to paper.qmd

@@ -121,31 +121,31 @@ pollution_2020 <- pollution_2020 %>%
   filter(City == "Toronto") %>% select(City, Pollutant, January, February, March, April, May, 
                                        June, July, August, September, October, November, December, 
                                        Average, StdDeviation) %>%
-  mutate(Year = 2020)
+  mutate(Year = "2020")
 
 pollution_2019 <- pollution_2019 %>%
   filter(City == "Toronto") %>% select(City, Pollutant, January, February, March, April, May, 
                                        June, July, August, September, October, November, December, 
                                        Average, StdDeviation) %>%
-  mutate(Year = 2019)
+  mutate(Year = "2019")
 
 pollution_2018 <- pollution_2018 %>%
   filter(City == "Toronto") %>% select(City, Pollutant, January, February, March, April, May, 
                                        June, July, August, September, October, November, December, 
                                        Average, StdDeviation) %>%
-  mutate(Year = 2018)
+  mutate(Year = "2018")
 
 pollution_2017 <- pollution_2017 %>%
   filter(City == "Toronto") %>% select(City, Pollutant, January, February, March, April, May, 
                                        June, July, August, September, October, November, December, 
                                        Average, StdDeviation) %>%
-  mutate(Year = 2017)
+  mutate(Year = "2017")
 
 pollution_2016 <- pollution_2016 %>%
   filter(City == "Toronto") %>% select(City, Pollutant, January, February, March, April, May, 
                                        June, July, August, September, October, November, December, 
                                        Average, StdDeviation) %>%
-  mutate(Year = 2016)
+  mutate(Year = "2016")
 
 pollution_2015 <- pollution_2015 %>%
   filter(City == "Toronto") %>% select(City, Pollutant, January, February, March, April, May, 
@@ -157,25 +157,25 @@ pollution_2014 <- pollution_2014 %>%
   filter(City == "Toronto") %>% select(City, Pollutant, January, February, March, April, May, 
                                        June, July, August, September, October, November, December, 
                                        Average, StdDeviation) %>%
-  mutate(Year = 2014)
+  mutate(Year = "2014")
 
 pollution_2013 <- pollution_2013 %>%
   filter(City == "Toronto") %>% select(City, Pollutant, January, February, March, April, May, 
                                        June, July, August, September, October, November, December, 
                                        Average, StdDeviation) %>%
-  mutate(Year = 2013)
+  mutate(Year = "2013")
 
 pollution_2012 <- pollution_2012 %>%
[…]

### 2022-06-21 outputs/paper/paper.qmd
Update paper.qmd
Updates to paper.qmd

@@
 These threats are becoming revealed in Canada as the country continues to face the consequences of warming almost at twice the global rate.
-In this report, I will explore how tempreture increases, lower air quality relate to a rise in threats to human healht in Toronto.
+The focus for this paper is on tempreture increases, lower air quality and more pollution and a decline in health for Toronto.
 Once the data for this research was collected, the data cleaning process begun which entailed of using packages in R like tidyverse, janitor, dplyr, ….
@@
 
-Why do we care?
-
-What is in this paper?
-
-Results?
-
 ## Data
-
-table of variables used?
 
@@
 
-## Visualizing the Data and the Implications (needs to be reworded)
-
-use other graphs beside boxplot. log scales. (dotpoint, barchart, etc)
-
-avoid drastic, dramatic, etc
+## Visualizing the Data and the Implications
 
@@
 
-need updated data?
-
 ## Conclusion

### 2022-06-21 outputs/paper/paper.qmd
Updates to paper.qmd and cleaning population data

@@
 
+As climate change worsens, the danger it presents to human health worsens as well.
+The health impacts from climate change are already being demonstrated in parts of Canada.
+Canada is warming at almost twice the global rate and if this continues, emissions will exceed 2 degrees celcius worldwide.
+
 ## Data
@@
 
+Temperature data was collected from ...
+
+Pollution data was collected from ...
+
 #### Variables of Interest
+
+Tempreature variables of interest: Station Name, Date, Maximum temperature, Minimum temperature, Mean temperature.
+
+Pollution varibales of interest: Pollutant, Month, Year, Average, StdDeviation
 

### 2022-06-21 scripts/01-data_cleaning.R

@@ -65,11 +65,10 @@ temp_2002_2022_10yr_interval <- rbind(temp_jan_dec_2002,
 
 temp_2002_2022 <- temp_2002_2022 %>% select(Station.Name, Date.Time, Year, 
                                             Month, Day, Max.Temp...C., 
-                                            Min.Temp...C., Mean.Temp...C., 
-                                            Total.Precip..mm.)
+                                            Min.Temp...C., Mean.Temp...C.)
 
 my_column_names <- c("station.name", "date", "year", "month", "day", "max.temp", "min.temp", 
-                     "mean.temp", "total.percip.mm")
+                     "mean.temp")
 
 colnames(temp_2002_2022) <- my_column_names
 temp_2002_2022 <- clean_names(temp_2002_2022)
@@ -96,9 +95,6 @@ temp_2002_2022 <- temp_2002_2022 %>%
 temp_2002_2022 <- temp_2002_2022 %>% drop_na(max_temp)
 temp_2002_2022 <- temp_2002_2022 %>% drop_na(min_temp)
 temp_2002_2022 <- temp_2002_2022 %>% drop_na(mean_temp)
-temp_2002_2022 <- temp_2002_2022 %>% drop_na(total_percip_mm)
-
-
 
 
 #### Cleaning Pollution Data ----

### 2022-06-21 scripts/02-data_analysis.R

@@ -10,3 +10,13 @@
 
 
 #### Workspace setup ####
+library(tidyverse)
+library(ggplot2)
+
+### Plots ----
+
+### Models ----
+
+### Anova Tests ----
+
+### t-tests ----

### 2022-06-21 scripts/plots.R

@@ -111,3 +111,4 @@ looking at rising temperatures and pollen and deaths"
 # pollution data ----
 
 ggplot(pollution_2011_2020, aes(Year, January, color=Pollutant)) + geom_point()
+

### 2022-06-22 outputs/paper/paper.qmd
Updating paper to reflect current findings

--- 2022-06-21 outputs/paper/paper.qmd
@@
 toc: true
-bibliography: references.bib
+# bibliography: references.bib
 ---
 
+\newpage
+
 ## Introduction
+
+As climate change worsens, the danger it presents to human health worsens as well.
+The health impacts due to climate change are already being demonstrated in parts of Canada.
+Canada is warming at almost twice the global rate and if this continues, emissions will exceed 2 degrees Celsius worldwide.
+
+The Paris Agreement within the United Nations Framework Convention on Climate Change includes 175 countries that have agreed to limit the average global temperature rise to below 2 degress Celcius.
+Canada had signed to the Paris Agreement as well and committed to cutting emission 30% below the 2005 levels by 2030.
+However, Canada's emission increased in 2016 compared to 1990.
+As a response to this increase, the Government of Canada developed a national climate change plan called *The Pan-Canadian Framework on Clean Growth and Climate Change*.
+This framework notes that human activities are driving severe changes to the Earth's climate which poses substantial
+risks to human health.
 
@@
 
+Temperature data was collected from the Government of Canada's website on Weather, Climate and Hazard's [page](https://climate.weather.gc.ca/historical_data/search_historic_data_e.html) containing past weather and climate data for the station "Toronto City".
+This data was collected from January 2002 to May 2022.
+The original data contains daily data on maximum temperature, minimum temperature, mean temperature, heat degree days, cool degree days, total rain (mm), total snow (cm), total precipitation (mm), snow on ground (cm), direction of maximum gust, and speed of maximum gust (km/h).
+For this research, I will focus on only maximum temperature, minimum temperature, and mean temperature.
+
+Pollution data was collected from the Government of Canada's National Air Pollution Surveillance (NAPS) Program [website](https://data-donnees.ec.gc.ca/data/air/monitor/national-air-pollution-surveillance-naps-program/Data-Donnees/?lang=en).
+This website contains pollution data from 2002 to 2020.
+Each year contains a file with continuous data and integrated data.
+I will be using continuous data for each year.
+The continuous data contains annual summaries and hourly data as well.
+I will be looking at the means for each year in the annual summaries data.
+These summaries include the city (which is Toronto in our case), the pollutant (these include: NO, NO2, NOX, O3, CO, PM2.5, and SO2), as well as the levels of each pollutant for each month of the year.
+
 #### Variables of Interest
+
+Temperature variables of interest:
+
+-   Date
+
+-   Maximum temperature
+
+-   Minimum temperature
+
+-   Mean temperature.
+
+Pollution variables of interest:
+
+-   Pollutant
+
+-   Month
+
+-   Year
+
+-   Average
+
+-   StdDeviation
 
@@
 
+#### The Data
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+
+# Load libraries
+library(tidyverse)
+library(knitr)
+library(kableExtra)
+library(ggplot2)
+library(dplyr)
+
+# Load temperature data
+
+temperature_2002_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/clean_temp_data/temperature_2002_2022.csv")
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="", fig.width=9, fig.height=3.2}
+
+temperature_2002_2022 %>%
+  select(year, month, day, max_temp, min_temp, mean_temp) %>%
+  slice(c(162, 393, 745, 1318, 1819, 2115, 3222, 4222, 5222, 6333)) %>%
+  kable(
+    caption = "Extracting rows from the Temperature data from 2002 - 2022",
+    col.names=c("Year", "Month", "Day", "Max Temp", "Min Temp", "Mean Temp"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+
+# Load pollution data
+
+poll_2011_2020 <- read.csv("~/Desktop/SUDS2022/inputs/data/clean_pollution_data/pollution_2011_2020.csv")
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="", fig.width=9, fig.height=3.2}
+
+poll_2011_2020 %>%
+  select(Year, Pollutant, January, February, March, April, May, June, July, August, September, October, 
+         November, December) %>%
+  slice(c(2, 34, 74, 91, 117, 147, 184, 223, 252, 308)) %>%
+  kable(
+    caption = "Extracting rows from the Pollution data from 2011 - 2020",
+    col.names=c("Year", "Pollutant", "Jan", "Feb", "Mar", "Apr", "May",
+                "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
+
+```
+
 ## Visualizing the Data and the Implications
 
-#### Visual 1
+#### Visualizing Temperature from 2002-2022
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2022 for December to March", fig.width=9, fig.height=3.2}
+
+dec_mar <- temperature_2002_2022 %>%
+  filter(month_name == "Dec" | month_name == "Jan" | month_name == "Feb" | month_name == "Mar")
+
+ggplot(dec_mar, aes(x=year, y=max_temp, color=month_name)) + geom_point()
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2022 for April to July", fig.width=9, fig.height=3.2}
+
+apr_jul <- temperature_2002_2022 %>%
+  filter(month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul")
+
+ggplot(apr_jul, aes(x=year, y=max_temp, color=month_name)) + geom_point()
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2022 for August to November", fig.width=9, fig.height=3.2}
+
+aug_nov <- temperature_2002_2022 %>%
+  filter(month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")
+
+ggplot(aug_nov, aes(x=year, y=max_temp, color=month_name)) + geom_point()
+
+```
 

### 2022-06-22 scripts/01-data_cleaning.R

@@ -73,6 +73,12 @@ my_column_names <- c("station.name", "date", "year", "month", "day", "max.temp",
 colnames(temp_2002_2022) <- my_column_names
 temp_2002_2022 <- clean_names(temp_2002_2022)
 
+# Save temperature data 2002-2022
+write_csv(
+  x = temp_2002_2022,
+  file = "~/Desktop/SUDS2022/inputs/data/clean_temp_data/temperature_2002_2022.csv"
+)
+
 
 ### Add column for Month Names
 temp_2002_2022 <- temp_2002_2022 %>%
@@ -191,5 +197,10 @@ df <- df %>% select(Pollutant, January, February, March, April, May,
                     December, Year)
 df <- t(df)
 
+# Save pollution data 2011-2020
+write_csv(
+  x = pollution_2011_2020,
+  file = "~/Desktop/SUDS2022/inputs/data/clean_pollution_data/pollution_2011_2020.csv"
+)
 
 #### Combine Temperature and Pollution Data ----

### 2022-06-22 scripts/plots.R

@@ -55,6 +55,25 @@ mean_temp_2002_2022 %>%
     linesep = "", digits = 2, booktabs=TRUE) %>%
   kable_styling(latex_options = "HOLD_position", font_size = 10)
 
+mean_temp_2002_2022 <- temp_2002_2022 %>%
+  select(year, month, day, max_temp, min_temp, mean_temp)
+mean_temp_2002_2022 %>%
+  slice(c(162, 393, 745, 1318, 1819, 2115, 3222, 4222, 5222, 6333)) %>%
+  kable(
+    caption = "Extracting rows from the Temperature data from 2002 - 2022",
+    col.names=c("Year", "Month", "Day", "Max Temp", "Min Temp", "Mean Temp"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
+
+temp_2002_2022 %>%
+  select(year, month, day, max_temp, min_temp, mean_temp) %>%
+  slice(c(162, 393, 745, 1318, 1819, 2115, 3222, 4222, 5222, 6333)) %>%
+  kable(
+    caption = "Extracting rows from the Temperature data from 2002 - 2022",
+    col.names=c("Year", "Month", "Day", "Max Temp", "Min Temp", "Mean Temp"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
+
 
 # summarising data ----
 summary_max_data <- temp_2002_2022 %>%
@@ -110,5 +129,18 @@ looking at rising temperatures and pollen and deaths"
 
 # pollution data ----
 
+## ggplot ----
 ggplot(pollution_2011_2020, aes(Year, January, color=Pollutant)) + geom_point()
 
+
+## tables ----
+pollution_2011_2020 %>%
+  select(Year, Pollutant, January, February, March, April, May, June, July, August, September, October, 
+         November, December) %>%
+  slice(c(2, 34, 74, 91, 117, 147, 184, 223, 252, 308)) %>%
+  kable(
+    caption = "Extracting rows from the Pollution data from 2011 - 2020",
+    col.names=c("Year", "Pollutant", "Jan", "Feb", "Mar", "Apr", "May",
+                "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)

### 2022-06-22 outputs/paper/paper.qmd
Update paper.qmd
Updating paper.qmd

@@
 
-## Visualizing the Data and the Implications
+## Visualizing Temperature from 2002-2022
 
-#### Visualizing Temperature from 2002-2022
+#### Visualizing Temperature by Months
 
@@
 
-ggplot(dec_mar, aes(x=year, y=max_temp, color=month_name)) + geom_point()
+ggplot(dec_mar, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
+  labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
 
@@
 
-ggplot(apr_jul, aes(x=year, y=max_temp, color=month_name)) + geom_point()
+ggplot(apr_jul, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
+  labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
 
@@
 
-ggplot(aug_nov, aes(x=year, y=max_temp, color=month_name)) + geom_point()
+ggplot(aug_nov, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
+  labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
 
 ```
-
-#### Implications 1
-
-#### Visual 2
-
-#### Implications 2
-
-#### Visuals 3
-
-#### Implications 3
 

### 2022-06-22 scripts/pollution.R
Updates to pollution data

@@ -1,0 +1,51 @@
+library(tidyverse)
+library(dplyr)
+library(ggplot2)
+library(janitor)
+library(naniar)
+
+CO_2020 <- read_csv("inputs/data/raw_pollution_data/2020/CO_2020.csv", 
+                    skip = 7)
+NO_2020 <- read_csv("inputs/data/raw_pollution_data/2020/NO_2020.csv", 
+                    skip = 7)
+NO2_2020 <- read_csv("inputs/data/raw_pollution_data/2020/NO2_2020.csv", 
+                     skip = 7)
+O3_2020 <- read_csv("inputs/data/raw_pollution_data/2020/O3_2020.csv", 
+                                                  skip = 7)
+SO2_2020 <- read_csv("inputs/data/raw_pollution_data/2020/SO2_2020.csv", 
+                     skip = 7)
+
+pollu_2020 <- rbind(CO_2020, NO_2020, NO2_2020, O3_2020, SO2_2020)
+my_column_names <- c("Pollutant", "NAPS ID", "City", "Province", "Latitude", "Longitude", "Date", "H1",
+                     "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12", "H13",
+                     "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
+                     "H24")
+colnames(pollu_2020) <- my_column_names
+pollu_2020 <- pollu_2020 %>% select("Pollutant", "City", "Date", "H1",
+                                    "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12",
+                                    "H13", "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
+                                    "H24")
+
+pollu_2020 <- pollu_2020 %>% filter(City == "Toronto")
+
+pollu_2020 <- pollu_2020 %>% distinct(Date, .keep_all = TRUE)
+
+pollu_2020 <- pollu_2020 %>% replace_with_na(replace = list(H1 = -999, H2 = -999, H3 = -999, 
+                                                            H4 = -999, H5 = -999, H6 = -999,
+                                                            H7 = -999, H8 = -999, H9 = -999, 
+                                                            H10 = -999, H11 = -999, H12 = -999, 
+                                                            H13 = -999, H14 = -999, H15 = -999,
[…]

### 2022-06-23 scripts/pollution.R
Adding new pollution data and cleaning it

@@ -4,6 +4,9 @@ library(ggplot2)
 library(janitor)
 library(naniar)
 
+## 2020 Data
+
+# read data
 CO_2020 <- read_csv("inputs/data/raw_pollution_data/2020/CO_2020.csv", 
                     skip = 7)
 NO_2020 <- read_csv("inputs/data/raw_pollution_data/2020/NO_2020.csv", 
@@ -15,21 +18,30 @@ O3_2020 <- read_csv("inputs/data/raw_pollution_data/2020/O3_2020.csv",
 SO2_2020 <- read_csv("inputs/data/raw_pollution_data/2020/SO2_2020.csv", 
                      skip = 7)
 
-pollu_2020 <- rbind(CO_2020, NO_2020, NO2_2020, O3_2020, SO2_2020)
+# change column names
 my_column_names <- c("Pollutant", "NAPS ID", "City", "Province", "Latitude", "Longitude", "Date", "H1",
                      "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12", "H13",
                      "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
                      "H24")
+colnames(CO_2020) <- my_column_names
+colnames(NO_2020) <- my_column_names
+colnames(NO2_2020) <- my_column_names
+colnames(O3_2020) <- my_column_names
+colnames(SO2_2020) <- my_column_names
+# combine data
+pollu_2020 <- rbind(CO_2020, NO_2020, NO2_2020, O3_2020, SO2_2020)
 colnames(pollu_2020) <- my_column_names
+# select variables
 pollu_2020 <- pollu_2020 %>% select("Pollutant", "City", "Date", "H1",
                                     "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12",
                                     "H13", "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
                                     "H24")
-
+# filter for Toronto
 pollu_2020 <- pollu_2020 %>% filter(City == "Toronto")
-
-pollu_2020 <- pollu_2020 %>% distinct(Date, .keep_all = TRUE)
-
+# remove duplicate dates
+unique_rows <- !duplicated(pollu_2020[c("Pollutant", "Date")])
+pollu_2020 <- pollu_2020[unique_rows,]
[…]

### 2022-06-23 scripts/plots.R
Updates to cleaning pollution data

@@ -133,6 +133,12 @@ looking at rising temperatures and pollen and deaths"
 ggplot(pollution_2011_2020, aes(Year, January, color=Pollutant)) + geom_point()
 
 
+
+
+ggplot(pollution, aes(Year, Mean, color=Pollutant)) + geom_point()
+
+
+
 ## tables ----
 pollution_2011_2020 %>%
   select(Year, Pollutant, January, February, March, April, May, June, July, August, September, October, 
@@ -144,3 +150,9 @@ pollution_2011_2020 %>%
                 "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"),
     linesep = "", digits = 2, booktabs=TRUE) %>%
   kable_styling(latex_options = "HOLD_position", font_size = 10)
+
+
+
+
+
+

### 2022-06-23 scripts/pollution.R

@@ -458,3 +458,107 @@ pollu_2012 <- pollu_2012 %>% mutate(Mean = rowMeans(pollu_2012[ ,c(4:27)], na.rm
 pollu_2012 <- pollu_2012 %>% mutate(Year = "2012")
 # Remove hourly data columns
 pollu_2012 <- pollu_2012 %>% select("Pollutant", "City", "Date", "Mean", "Year")
+
+## 2011 Data
+
+# read data
+CO_2011 <- read_csv("inputs/data/raw_pollution_data/2011/CO_2011.csv", 
+                    skip = 7)
+NO_2011 <- read_csv("inputs/data/raw_pollution_data/2011/NO_2011.csv", 
+                    skip = 7)
+NO2_2011 <- read_csv("inputs/data/raw_pollution_data/2011/NO2_2011.csv", 
+                     skip = 7)
+O3_2011 <- read_csv("inputs/data/raw_pollution_data/2011/O3_2011.csv", 
+                    skip = 7)
+SO2_2011 <- read_csv("inputs/data/raw_pollution_data/2011/SO2_2011.csv", 
+                     skip = 7)
+colnames(CO_2011) <- my_column_names
+colnames(NO_2011) <- my_column_names
+colnames(NO2_2011) <- my_column_names
+colnames(O3_2011) <- my_column_names
+colnames(SO2_2011) <- my_column_names
+# combine data
+pollu_2011 <- rbind(CO_2011, NO_2011, NO2_2011, O3_2011, SO2_2011)
+# change column names
+colnames(pollu_2011) <- my_column_names
+# select variables
+pollu_2011 <- pollu_2011 %>% select("Pollutant", "City", "Date", "H1",
+                                    "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12",
+                                    "H13", "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
+                                    "H24")
+# filter for Toronto
+pollu_2011 <- pollu_2011 %>% filter(City == "Toronto")
+# remove duplicate dates
+unique_rows <- !duplicated(pollu_2011[c("Pollutant", "Date")])
+pollu_2011 <- pollu_2011[unique_rows,]
+# replace -999 missing values with NA
+pollu_2011 <- pollu_2011 %>% replace_with_na(replace = list(H1 = -999, H2 = -999, H3 = -999, 
+                                                            H4 = -999, H5 = -999, H6 = -999,
[…]

### 2022-06-23 scripts/plots.R
Update plots.R

@@ -154,5 +154,12 @@ pollution_2011_2020 %>%
 
 
 
-
+pollution %>%
+  select(Pollutant, City, Date, Mean, Year) %>%
+  slice(c(100, 1000, 1200, 1500, 1800, 2000, 2500, 2800, 3600, 5000, 6000, 7000, 10000, 11000, 15000)) %>%
+  kable(
+    caption = "Extracting rows from the Pollution data from 2010 - 2020",
+    col.names=c("Pollutant", "City", "Date", "Mean", "Year"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
 

### 2022-06-24 scripts/plots.R
Adding 2009 data

@@ -137,7 +137,35 @@ ggplot(pollution_2011_2020, aes(Year, January, color=Pollutant)) + geom_point()
 
 ggplot(pollution, aes(Year, Mean, color=Pollutant)) + geom_point()
 
+pollution %>% filter(Pollutant != "NO") %>% ggplot(aes(x = Year, 
+                                                       y = Mean, 
+                                                       color = Pollutant)) + geom_point()
+pollution %>% filter(Pollutant != "NO" & Pollutant != "O3") %>% ggplot(aes(x = Year, 
+                                                       y = Mean, 
+                                                       color = Pollutant)) + geom_point()
 
+pollution %>% filter(Pollutant != "NO" & Pollutant != "O3" & Pollutant != "NO2") %>% ggplot(
+  aes(x = Year, y = Mean, color = Pollutant)) + geom_point()
+# CO unit ppm
+pollution %>% filter(Pollutant == "CO") %>% ggplot(
+  aes(x = Year, y = Mean, color = Pollutant)) + geom_point()
+pollution %>% filter(Pollutant == "CO") %>% ggplot(
+  aes(x = Year, y = Mean*1000, color = Pollutant)) + geom_point()
+# NO unit ppb
+pollution %>% filter(Pollutant == "NO") %>% ggplot(
+  aes(x = Year, y = Mean, color = Pollutant)) + geom_point()
+# O3 unit ppb
+pollution %>% filter(Pollutant == "O3") %>% ggplot(
+  aes(x = Year, y = Mean, color = Pollutant)) + geom_point()
+# NO2 unit ppb
+pollution %>% filter(Pollutant == "NO2") %>% ggplot(
+  aes(x = Year, y = Mean, color = Pollutant)) + geom_point()
+# SO2 unit ppb
+pollution %>% filter(Pollutant == "SO2") %>% ggplot(
+  aes(x = Year, y = Mean, color = Pollutant)) + geom_point()
+## different units for CO for the above so different axis
+pollution %>% filter(Pollutant != "CO") %>% ggplot(
+  aes(x = Year, y = Mean, color = Pollutant)) + geom_point()
 
 ## tables ----
 pollution_2011_2020 %>%

### 2022-06-24 scripts/pollution.R
Update pollution.R

@@ -559,6 +559,56 @@ pollu_2010 <- pollu_2010 %>% mutate(Year = "2010")
 # Remove hourly data columns
 pollu_2010 <- pollu_2010 %>% select("Pollutant", "City", "Date", "Mean", "Year")
 
+## 2009 Data
+
+# read data
+CO_2009 <- read_csv("inputs/data/raw_pollution_data/2009/CO_2009.csv", 
+                    skip = 7)
+NO_2009 <- read_csv("inputs/data/raw_pollution_data/2009/NO_2009.csv", 
+                    skip = 7)
+NO2_2009 <- read_csv("inputs/data/raw_pollution_data/2009/NO2_2009.csv", 
+                     skip = 7)
+O3_2009 <- read_csv("inputs/data/raw_pollution_data/2009/O3_2009.csv", 
+                    skip = 7)
+SO2_2009 <- read_csv("inputs/data/raw_pollution_data/2009/SO2_2009.csv", 
+                     skip = 7)
+colnames(CO_2009) <- my_column_names
+colnames(NO_2009) <- my_column_names
+colnames(NO2_2009) <- my_column_names
+colnames(O3_2009) <- my_column_names
+colnames(SO2_2009) <- my_column_names
+# combine data
+pollu_2009 <- rbind(CO_2009, NO_2009, NO2_2009, O3_2009, SO2_2009)
+# change column names
+colnames(pollu_2009) <- my_column_names
+# select variables
+pollu_2009 <- pollu_2009 %>% select("Pollutant", "City", "Date", "H1",
+                                    "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12",
+                                    "H13", "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
+                                    "H24")
+# filter for Toronto
+pollu_2009 <- pollu_2009 %>% filter(City == "Toronto")
+# remove duplicate dates
+unique_rows <- !duplicated(pollu_2009[c("Pollutant", "Date")])
+pollu_2009 <- pollu_2009[unique_rows,]
+# replace -999 missing values with NA
+pollu_2009 <- pollu_2009 %>% replace_with_na(replace = list(H1 = -999, H2 = -999, H3 = -999, 
+                                                            H4 = -999, H5 = -999, H6 = -999,
+                                                            H7 = -999, H8 = -999, H9 = -999, 
[…]

### 2022-06-24 scripts/pollution.R
Cleaning 2008 pollution data

@@ -609,6 +609,56 @@ pollu_2009 <- pollu_2009 %>% mutate(Year = "2009")
 # Remove hourly data columns
 pollu_2009 <- pollu_2009 %>% select("Pollutant", "City", "Date", "Mean", "Year")
 
+## 2008 Data
+
+# read data
+CO_2008 <- read_csv("inputs/data/raw_pollution_data/2008/CO_2008.csv", 
+                    skip = 7)
+NO_2008 <- read_csv("inputs/data/raw_pollution_data/2008/NO_2008.csv", 
+                    skip = 7)
+NO2_2008 <- read_csv("inputs/data/raw_pollution_data/2008/NO2_2008.csv", 
+                     skip = 7)
+O3_2008 <- read_csv("inputs/data/raw_pollution_data/2008/O3_2008.csv", 
+                    skip = 7)
+SO2_2008 <- read_csv("inputs/data/raw_pollution_data/2008/SO2_2008.csv", 
+                     skip = 7)
+colnames(CO_2008) <- my_column_names
+colnames(NO_2008) <- my_column_names
+colnames(NO2_2008) <- my_column_names
+colnames(O3_2008) <- my_column_names
+colnames(SO2_2008) <- my_column_names
+# combine data
+pollu_2008 <- rbind(CO_2008, NO_2008, NO2_2008, O3_2008, SO2_2008)
+# change column names
+colnames(pollu_2008) <- my_column_names
+# select variables
+pollu_2008 <- pollu_2008 %>% select("Pollutant", "City", "Date", "H1",
+                                    "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12",
+                                    "H13", "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
+                                    "H24")
+# filter for Toronto
+pollu_2008 <- pollu_2008 %>% filter(City == "Toronto")
+# remove duplicate dates
+unique_rows <- !duplicated(pollu_2008[c("Pollutant", "Date")])
+pollu_2008 <- pollu_2008[unique_rows,]
+# replace -999 missing values with NA
+pollu_2008 <- pollu_2008 %>% replace_with_na(replace = list(H1 = -999, H2 = -999, H3 = -999, 
+                                                            H4 = -999, H5 = -999, H6 = -999,
+                                                            H7 = -999, H8 = -999, H9 = -999, 
[…]

### 2022-06-24 scripts/pollution.R
Cleaning 2007 pollution data

@@ -659,6 +659,56 @@ pollu_2008 <- pollu_2008 %>% mutate(Year = "2008")
 # Remove hourly data columns
 pollu_2008 <- pollu_2008 %>% select("Pollutant", "City", "Date", "Mean", "Year")
 
+## 2007 Data
+
+# read data
+CO_2007 <- read_csv("inputs/data/raw_pollution_data/2007/CO_2007.csv", 
+                    skip = 7)
+NO_2007 <- read_csv("inputs/data/raw_pollution_data/2007/NO_2007.csv", 
+                    skip = 7)
+NO2_2007 <- read_csv("inputs/data/raw_pollution_data/2007/NO2_2007.csv", 
+                     skip = 7)
+O3_2007 <- read_csv("inputs/data/raw_pollution_data/2007/O3_2007.csv", 
+                    skip = 7)
+SO2_2007 <- read_csv("inputs/data/raw_pollution_data/2007/SO2_2007.csv", 
+                     skip = 7)
+colnames(CO_2007) <- my_column_names
+colnames(NO_2007) <- my_column_names
+colnames(NO2_2007) <- my_column_names
+colnames(O3_2007) <- my_column_names
+colnames(SO2_2007) <- my_column_names
+# combine data
+pollu_2007 <- rbind(CO_2007, NO_2007, NO2_2007, O3_2007, SO2_2007)
+# change column names
+colnames(pollu_2007) <- my_column_names
+# select variables
+pollu_2007 <- pollu_2007 %>% select("Pollutant", "City", "Date", "H1",
+                                    "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12",
+                                    "H13", "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
+                                    "H24")
+# filter for Toronto
+pollu_2007 <- pollu_2007 %>% filter(City == "Toronto")
+# remove duplicate dates
+unique_rows <- !duplicated(pollu_2007[c("Pollutant", "Date")])
+pollu_2007 <- pollu_2007[unique_rows,]
+# replace -999 missing values with NA
+pollu_2007 <- pollu_2007 %>% replace_with_na(replace = list(H1 = -999, H2 = -999, H3 = -999, 
+                                                            H4 = -999, H5 = -999, H6 = -999,
+                                                            H7 = -999, H8 = -999, H9 = -999, 
[…]

### 2022-06-24 scripts/pollution.R
Cleaning 2006 pollution data

@@ -709,6 +709,57 @@ pollu_2007 <- pollu_2007 %>% mutate(Year = "2007")
 # Remove hourly data columns
 pollu_2007 <- pollu_2007 %>% select("Pollutant", "City", "Date", "Mean", "Year")
 
+
+## 2006 Data
+
+# read data
+CO_2006 <- read_csv("inputs/data/raw_pollution_data/2006/CO_2006.csv", 
+                    skip = 7)
+NO_2006 <- read_csv("inputs/data/raw_pollution_data/2006/NO_2006.csv", 
+                    skip = 7)
+NO2_2006 <- read_csv("inputs/data/raw_pollution_data/2006/NO2_2006.csv", 
+                     skip = 7)
+O3_2006 <- read_csv("inputs/data/raw_pollution_data/2006/O3_2006.csv", 
+                    skip = 7)
+SO2_2006 <- read_csv("inputs/data/raw_pollution_data/2006/SO2_2006.csv", 
+                     skip = 7)
+colnames(CO_2006) <- my_column_names
+colnames(NO_2006) <- my_column_names
+colnames(NO2_2006) <- my_column_names
+colnames(O3_2006) <- my_column_names
+colnames(SO2_2006) <- my_column_names
+# combine data
+pollu_2006 <- rbind(CO_2006, NO_2006, NO2_2006, O3_2006, SO2_2006)
+# change column names
+colnames(pollu_2006) <- my_column_names
+# select variables
+pollu_2006 <- pollu_2006 %>% select("Pollutant", "City", "Date", "H1",
+                                    "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12",
+                                    "H13", "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
+                                    "H24")
+# filter for Toronto
+pollu_2006 <- pollu_2006 %>% filter(City == "Toronto")
+# remove duplicate dates
+unique_rows <- !duplicated(pollu_2006[c("Pollutant", "Date")])
+pollu_2006 <- pollu_2006[unique_rows,]
+# replace -999 missing values with NA
+pollu_2006 <- pollu_2006 %>% replace_with_na(replace = list(H1 = -999, H2 = -999, H3 = -999, 
+                                                            H4 = -999, H5 = -999, H6 = -999,
+                                                            H7 = -999, H8 = -999, H9 = -999, 
+                                                            H10 = -999, H11 = -999, H12 = -999, 
[…]

### 2022-06-24 scripts/pollution.R
Cleaning 2005 pollution data

@@ -760,6 +760,56 @@ pollu_2006 <- pollu_2006 %>% mutate(Year = "2006")
 # Remove hourly data columns
 pollu_2006 <- pollu_2006 %>% select("Pollutant", "City", "Date", "Mean", "Year")
 
+## 2005 Data
+
+# read data
+CO_2005 <- read_csv("inputs/data/raw_pollution_data/2005/CO_2005.csv", 
+                    skip = 7)
+NO_2005 <- read_csv("inputs/data/raw_pollution_data/2005/NO_2005.csv", 
+                    skip = 7)
+NO2_2005 <- read_csv("inputs/data/raw_pollution_data/2005/NO2_2005.csv", 
+                     skip = 7)
+O3_2005 <- read_csv("inputs/data/raw_pollution_data/2005/O3_2005.csv", 
+                    skip = 7)
+SO2_2005 <- read_csv("inputs/data/raw_pollution_data/2005/SO2_2005.csv", 
+                     skip = 7)
+colnames(CO_2005) <- my_column_names
+colnames(NO_2005) <- my_column_names
+colnames(NO2_2005) <- my_column_names
+colnames(O3_2005) <- my_column_names
+colnames(SO2_2005) <- my_column_names
+# combine data
+pollu_2005 <- rbind(CO_2005, NO_2005, NO2_2005, O3_2005, SO2_2005)
+# change column names
+colnames(pollu_2005) <- my_column_names
+# select variables
+pollu_2005 <- pollu_2005 %>% select("Pollutant", "City", "Date", "H1",
+                                    "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12",
+                                    "H13", "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
+                                    "H24")
+# filter for Toronto
+pollu_2005 <- pollu_2005 %>% filter(City == "Toronto")
+# remove duplicate dates
+unique_rows <- !duplicated(pollu_2005[c("Pollutant", "Date")])
+pollu_2005 <- pollu_2005[unique_rows,]
+# replace -999 missing values with NA
+pollu_2005 <- pollu_2005 %>% replace_with_na(replace = list(H1 = -999, H2 = -999, H3 = -999, 
+                                                            H4 = -999, H5 = -999, H6 = -999,
+                                                            H7 = -999, H8 = -999, H9 = -999, 
[…]

### 2022-06-24 scripts/pollution.R
Cleaning 2004 pollution data

@@ -810,6 +810,57 @@ pollu_2005 <- pollu_2005 %>% mutate(Year = "2005")
 # Remove hourly data columns
 pollu_2005 <- pollu_2005 %>% select("Pollutant", "City", "Date", "Mean", "Year")
 
+## 2004 Data
+
+# read data
+CO_2004 <- read_csv("inputs/data/raw_pollution_data/2004/CO_2004.csv", 
+                    skip = 7)
+NO_2004 <- read_csv("inputs/data/raw_pollution_data/2004/NO_2004.csv", 
+                    skip = 7)
+NO2_2004 <- read_csv("inputs/data/raw_pollution_data/2004/NO2_2004.csv", 
+                     skip = 7)
+O3_2004 <- read_csv("inputs/data/raw_pollution_data/2004/O3_2004.csv", 
+                    skip = 7)
+SO2_2004 <- read_csv("inputs/data/raw_pollution_data/2004/SO2_2004.csv", 
+                     skip = 7)
+colnames(CO_2004) <- my_column_names
+colnames(NO_2004) <- my_column_names
+colnames(NO2_2004) <- my_column_names
+colnames(O3_2004) <- my_column_names
+colnames(SO2_2004) <- my_column_names
+# combine data
+pollu_2004 <- rbind(CO_2004, NO_2004, NO2_2004, O3_2004, SO2_2004)
+# change column names
+colnames(pollu_2004) <- my_column_names
+# select variables
+pollu_2004 <- pollu_2004 %>% select("Pollutant", "City", "Date", "H1",
+                                    "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12",
+                                    "H13", "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
+                                    "H24")
+# filter for Toronto
+pollu_2004 <- pollu_2004 %>% filter(City == "Toronto")
+# remove duplicate dates
+unique_rows <- !duplicated(pollu_2004[c("Pollutant", "Date")])
+pollu_2004 <- pollu_2004[unique_rows,]
+# replace -999 missing values with NA
+pollu_2004 <- pollu_2004 %>% replace_with_na(replace = list(H1 = -999, H2 = -999, H3 = -999, 
+                                                            H4 = -999, H5 = -999, H6 = -999,
+                                                            H7 = -999, H8 = -999, H9 = -999, 
[…]

### 2022-06-24 scripts/pollution.R
Clean 2003 pollution data

@@ -861,6 +861,56 @@ pollu_2004 <- pollu_2004 %>% mutate(Year = "2004")
 pollu_2004 <- pollu_2004 %>% select("Pollutant", "City", "Date", "Mean", "Year")
 
 
+## 2003 Data
+
+# read data
+CO_2003 <- read_csv("inputs/data/raw_pollution_data/2003/CO_2003.csv", 
+                    skip = 7)
+NO_2003 <- read_csv("inputs/data/raw_pollution_data/2003/NO_2003.csv", 
+                    skip = 7)
+NO2_2003 <- read_csv("inputs/data/raw_pollution_data/2003/NO2_2003.csv", 
+                     skip = 7)
+O3_2003 <- read_csv("inputs/data/raw_pollution_data/2003/O3_2003.csv", 
+                    skip = 7)
+SO2_2003 <- read_csv("inputs/data/raw_pollution_data/2003/SO2_2003.csv", 
+                     skip = 7)
+colnames(CO_2003) <- my_column_names
+colnames(NO_2003) <- my_column_names
+colnames(NO2_2003) <- my_column_names
+colnames(O3_2003) <- my_column_names
+colnames(SO2_2003) <- my_column_names
+# combine data
+pollu_2003 <- rbind(CO_2003, NO_2003, NO2_2003, O3_2003, SO2_2003)
+# change column names
+colnames(pollu_2003) <- my_column_names
+# select variables
+pollu_2003 <- pollu_2003 %>% select("Pollutant", "City", "Date", "H1",
+                                    "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12",
+                                    "H13", "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
+                                    "H24")
+# filter for Toronto
+pollu_2003 <- pollu_2003 %>% filter(City == "Toronto")
+# remove duplicate dates
+unique_rows <- !duplicated(pollu_2003[c("Pollutant", "Date")])
+pollu_2003 <- pollu_2003[unique_rows,]
+# replace -999 missing values with NA
+pollu_2003 <- pollu_2003 %>% replace_with_na(replace = list(H1 = -999, H2 = -999, H3 = -999, 
+                                                            H4 = -999, H5 = -999, H6 = -999,
+                                                            H7 = -999, H8 = -999, H9 = -999, 
+                                                            H10 = -999, H11 = -999, H12 = -999, 
[…]

### 2022-06-24 scripts/01-data_cleaning.R
Cleaning pollution data

@@ -105,102 +105,969 @@ temp_2002_2022 <- temp_2002_2022 %>% drop_na(mean_temp)
 
 #### Cleaning Pollution Data ----
 
-### Attach pollution data
-attach(pollution_2020)
-attach(pollution_2019)
-attach(pollution_2018)
-attach(pollution_2017)
-attach(pollution_2016)
-attach(pollution_2015)
-attach(pollution_2014)
-attach(pollution_2013)
-attach(pollution_2012)
-attach(pollution_2011)
+## 2020 Data
 
-### Filter for Toronto and reduce to variables of interest and add year
+# read data
+CO_2020 <- read_csv("inputs/data/raw_pollution_data/2020/CO_2020.csv", 
+                    skip = 7)
+NO_2020 <- read_csv("inputs/data/raw_pollution_data/2020/NO_2020.csv", 
+                    skip = 7)
+NO2_2020 <- read_csv("inputs/data/raw_pollution_data/2020/NO2_2020.csv", 
+                     skip = 7)
+O3_2020 <- read_csv("inputs/data/raw_pollution_data/2020/O3_2020.csv", 
+                    skip = 7)
+SO2_2020 <- read_csv("inputs/data/raw_pollution_data/2020/SO2_2020.csv", 
+                     skip = 7)
 
-pollution_2020 <- pollution_2020 %>%
-  filter(City == "Toronto") %>% select(City, Pollutant, January, February, March, April, May, 
-                                       June, July, August, September, October, November, December, 
-                                       Average, StdDeviation) %>%
-  mutate(Year = "2020")
+# change column names
+my_column_names <- c("Pollutant", "NAPS ID", "City", "Province", "Latitude", "Longitude", "Date", "H1",
+                     "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12", "H13",
+                     "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
+                     "H24")
+colnames(CO_2020) <- my_column_names
+colnames(NO_2020) <- my_column_names
+colnames(NO2_2020) <- my_column_names
+colnames(O3_2020) <- my_column_names
+colnames(SO2_2020) <- my_column_names
+# combine data
+pollu_2020 <- rbind(CO_2020, NO_2020, NO2_2020, O3_2020, SO2_2020)
+colnames(pollu_2020) <- my_column_names
+# select variables
[…]

### 2022-06-24 scripts/pollution.R

@@ -911,6 +911,59 @@ pollu_2003 <- pollu_2003 %>% mutate(Year = "2003")
 # Remove hourly data columns
 pollu_2003 <- pollu_2003 %>% select("Pollutant", "City", "Date", "Mean", "Year")
 
+
+## 2002 Data
+
+# read data
+CO_2002 <- read_csv("inputs/data/raw_pollution_data/2002/CO_2002.csv", 
+                    skip = 7)
+NO_2002 <- read_csv("inputs/data/raw_pollution_data/2002/NO_2002.csv", 
+                    skip = 7)
+NO2_2002 <- read_csv("inputs/data/raw_pollution_data/2002/NO2_2002.csv", 
+                     skip = 7)
+O3_2002 <- read_csv("inputs/data/raw_pollution_data/2002/O3_2002.csv", 
+                    skip = 7)
+SO2_2002 <- read_csv("inputs/data/raw_pollution_data/2002/SO2_2002.csv", 
+                     skip = 7)
+colnames(CO_2002) <- my_column_names
+colnames(NO_2002) <- my_column_names
+colnames(NO2_2002) <- my_column_names
+colnames(O3_2002) <- my_column_names
+colnames(SO2_2002) <- my_column_names
+# combine data
+pollu_2002 <- rbind(CO_2002, NO_2002, NO2_2002, O3_2002, SO2_2002)
+# change column names
+colnames(pollu_2002) <- my_column_names
+# select variables
+pollu_20032 <- pollu_2002 %>% select("Pollutant", "City", "Date", "H1",
+                                    "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12",
+                                    "H13", "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
+                                    "H24")
+# filter for Toronto
+pollu_2002 <- pollu_2002 %>% filter(City == "Toronto")
+# remove duplicate dates
+unique_rows <- !duplicated(pollu_2002[c("Pollutant", "Date")])
+pollu_2002 <- pollu_2002[unique_rows,]
+# replace -999 missing values with NA
+pollu_2002 <- pollu_2002 %>% replace_with_na(replace = list(H1 = -999, H2 = -999, H3 = -999, 
+                                                            H4 = -999, H5 = -999, H6 = -999,
+                                                            H7 = -999, H8 = -999, H9 = -999, 
[…]

### 2022-06-24 scripts/pollution.R
Deleting pollution script

@@ -1,969 +1,0 @@
-library(tidyverse)
-library(dplyr)
-library(ggplot2)
-library(janitor)
-library(naniar)
-
-## 2020 Data
-
-# read data
-CO_2020 <- read_csv("inputs/data/raw_pollution_data/2020/CO_2020.csv", 
-                    skip = 7)
-NO_2020 <- read_csv("inputs/data/raw_pollution_data/2020/NO_2020.csv", 
-                    skip = 7)
-NO2_2020 <- read_csv("inputs/data/raw_pollution_data/2020/NO2_2020.csv", 
-                     skip = 7)
-O3_2020 <- read_csv("inputs/data/raw_pollution_data/2020/O3_2020.csv", 
-                                                  skip = 7)
-SO2_2020 <- read_csv("inputs/data/raw_pollution_data/2020/SO2_2020.csv", 
-                     skip = 7)
-
-# change column names
-my_column_names <- c("Pollutant", "NAPS ID", "City", "Province", "Latitude", "Longitude", "Date", "H1",
-                     "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12", "H13",
-                     "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
-                     "H24")
-colnames(CO_2020) <- my_column_names
-colnames(NO_2020) <- my_column_names
-colnames(NO2_2020) <- my_column_names
-colnames(O3_2020) <- my_column_names
-colnames(SO2_2020) <- my_column_names
-# combine data
-pollu_2020 <- rbind(CO_2020, NO_2020, NO2_2020, O3_2020, SO2_2020)
-colnames(pollu_2020) <- my_column_names
-# select variables
-pollu_2020 <- pollu_2020 %>% select("Pollutant", "City", "Date", "H1",
-                                    "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12",
-                                    "H13", "H14", "H15", "H16", "H17", "H18", "H19", "H20", "H21", "H22", "H23", 
-                                    "H24")
-# filter for Toronto
-pollu_2020 <- pollu_2020 %>% filter(City == "Toronto")
-# remove duplicate dates
-unique_rows <- !duplicated(pollu_2020[c("Pollutant", "Date")])
-pollu_2020 <- pollu_2020[unique_rows,]
[…]

### 2022-06-27 outputs/paper/paper.qmd
Updating paper.qmd

@@
 As a response to this increase, the Government of Canada developed a national climate change plan called *The Pan-Canadian Framework on Clean Growth and Climate Change*.
-This framework notes that human activities are driving severe changes to the Earth's climate which poses substantial
-risks to human health.
+This framework notes that human activities are driving severe changes to the Earth's climate which poses substantial risks to human health.
 
@@
 I will be using continuous data for each year.
-The continuous data contains annual summaries and hourly data as well.
-I will be looking at the means for each year in the annual summaries data.
-These summaries include the city (which is Toronto in our case), the pollutant (these include: NO, NO2, NOX, O3, CO, PM2.5, and SO2), as well as the levels of each pollutant for each month of the year.
+The continuous data contains annual summaries and hourly data.
+I will be looking at the hourly data for CO (Carbon Monoxide), NO (Nitric Oxide), NO2 (Nitrogen Dioxide) , O3 (Ozone), and SO2 (Sulfur Dioxide).
+This data includes the city (which is Toronto in our case), the levels of each pollutant for each hour of every day.
 
@@
 
--   Average
-
--   StdDeviation
+-   Mean level of pollution each day
 

### 2022-06-29 scripts/plots.R
Updates to analysis

@@ -163,6 +163,7 @@ pollution %>% filter(Pollutant == "NO2") %>% ggplot(
 # SO2 unit ppb
 pollution %>% filter(Pollutant == "SO2") %>% ggplot(
   aes(x = Year, y = Mean, color = Pollutant)) + geom_point()
+
 ## different units for CO for the above so different axis
 pollution %>% filter(Pollutant != "CO") %>% ggplot(
   aes(x = Year, y = Mean, color = Pollutant)) + geom_point()

### 2022-06-29 scripts/02-data_analysis.R
Update 02-data_analysis.R

@@ -11,9 +11,41 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(knitr)
+library(kableExtra)
 library(ggplot2)
+library(dplyr)
+
+# Load temperature data
+temperature_2002_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/
+                                  clean_temp_data/temperature_2002_2022.csv")
+# Load pollution data
+poll_2011_2020 <- read.csv("~/Desktop/SUDS2022/inputs/data/
+                           clean_pollution_data/pollution_2011_2020.csv")
 
 ### Plots ----
+
+## Temperature ----
+
+# Visualizing Temperature by Months
+dec_mar <- temperature_2002_2022 %>%
+  filter(month_name == "Dec" | month_name == "Jan" | month_name == "Feb" | month_name == "Mar")
+
+ggplot(dec_mar, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
+  labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
+
+apr_jul <- temperature_2002_2022 %>%
+  filter(month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul")
+
+ggplot(apr_jul, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
+  labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
+
+aug_nov <- temperature_2002_2022 %>%
+  filter(month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")
+
+ggplot(aug_nov, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
+  labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
+
 
 ### Models ----
 

### 2022-06-29 scripts/00-download_data.R
Cleaning scripts

@@ -15,8 +15,9 @@ library(tidyverse)
 library(dplyr)
 library(readr)
 
-
 #### Load temperature datasets ####
+
+# Use below lines to load data sets if running for the first time
 temp_jan_may_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_climate_daily_ON_6158355_2022_P1D.csv")
 temp_jan_dec_2021 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_climate_daily_ON_6158355_2021_P1D.csv")
 temp_jan_dec_2020 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_climate_daily_ON_6158355_2020_P1D.csv")
@@ -62,32 +63,256 @@ file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2004_P1D
 file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2003_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_dec_2003.csv")
 file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2002_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_dec_2002.csv")
 
-
-#
-
+# Use below lines if running after the first time
+temp_jan_may_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/temp_jan_dec_2022.csv")
+temp_jan_dec_2021 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/temp_jan_dec_2021.csv")
+temp_jan_dec_2020 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/temp_jan_dec_2020.csv")
+temp_jan_dec_2019 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/temp_jan_dec_2019.csv")
+temp_jan_dec_2018 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/temp_jan_dec_2018.csv")
+temp_jan_dec_2017 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/temp_jan_dec_2017.csv")
[…]

### 2022-06-29 scripts/01-data_cleaning.R

@@ -17,30 +17,14 @@ library(janitor)
 
 #### Cleaning Temperature Data ----
 
-### Attach temperature data
-attach(temp_jan_dec_2002)
-attach(temp_jan_dec_2003)
-attach(temp_jan_dec_2004)
-attach(temp_jan_dec_2005)
-attach(temp_jan_dec_2006)
-attach(temp_jan_dec_2007)
-attach(temp_jan_dec_2008)
-attach(temp_jan_dec_2009)
-attach(temp_jan_dec_2010)
-attach(temp_jan_dec_2011)
-attach(temp_jan_dec_2012)
-attach(temp_jan_dec_2013)
-attach(temp_jan_dec_2014)
-attach(temp_jan_dec_2015)
-attach(temp_jan_dec_2016)
-attach(temp_jan_dec_2017)
-attach(temp_jan_dec_2018)
-attach(temp_jan_dec_2019)
-attach(temp_jan_dec_2020)
-attach(temp_jan_dec_2021)
-attach(temp_jan_may_2022)
+# Change column names 
+my_column_names <- c("station.name", "date", "year", "month", "day", "max.temp", "min.temp", 
+                     "mean.temp")
+
+colnames(temp_2002_2022) <- my_column_names
 
 # Combine 2002 - 2022 data
+
 temp_2002_2022 <- rbind(temp_jan_dec_2002, temp_jan_dec_2003, temp_jan_dec_2004, 
                         temp_jan_dec_2005, temp_jan_dec_2006, temp_jan_dec_2007,
                         temp_jan_dec_2008, temp_jan_dec_2009, temp_jan_dec_2010,
@@ -49,36 +33,10 @@ temp_2002_2022 <- rbind(temp_jan_dec_2002, temp_jan_dec_2003, temp_jan_dec_2004,
                         temp_jan_dec_2017, temp_jan_dec_2018, temp_jan_dec_2019,
                         temp_jan_dec_2020, temp_jan_dec_2021, temp_jan_may_2022)
 
-temp_2012_2022 <- rbind(temp_jan_dec_2012, temp_jan_dec_2013, 
-                        temp_jan_dec_2014, temp_jan_dec_2015, temp_jan_dec_2016,
-                        temp_jan_dec_2017, temp_jan_dec_2018, temp_jan_dec_2019,
-                        temp_jan_dec_2020, temp_jan_dec_2021, temp_jan_may_2022)
-
-
[…]

### 2022-06-29 scripts/02-data_analysis.R

@@ -46,6 +46,11 @@ aug_nov <- temperature_2002_2022 %>%
 ggplot(aug_nov, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
   labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
 
+## Pollution ----
+
+# Visualizing Pollution by Months
+
+
 
 ### Models ----
 

### 2022-06-29 scripts/01-data_cleaning.R
Update 01-data_cleaning.R

@@ -18,10 +18,34 @@ library(janitor)
 #### Cleaning Temperature Data ----
 
 # Change column names 
-my_column_names <- c("station.name", "date", "year", "month", "day", "max.temp", "min.temp", 
-                     "mean.temp")
+my_column_names <- c("Longitude (x)", "Latitude (y)", "Station Name", "Climate ID", 
+                     "Date", "Year", "Month", "Day", "Data Quality", "Max Temp", 
+                     "Max Temp Flag", "Min Temp", "Min Temp Flag", "Mean Temp", 
+                     "Mean Temp Flag", "Heat Deg Days", "Heat Deg Days Flag", 
+                     "Cool Deg Days", "Cool Deg Days Flag", "Total Rain", 
+                     "Total Rain Flag", "Total Snow", "Total Snow Flag", "Total Precip", 
+                     "Total Precip Flag", "Snow on Grnd", "Snow on Grnd Flag", 
+                     "Dir of Max Gust", "Dir of Max Gust Flag", "Spd of Max Gust", 
+                     "Spd of Max Gust Flag")
 
+colnames(temp_jan_may_2022) <- my_column_names
+colnames(temp_jan_dec_2021) <- my_column_names
 colnames(temp_2002_2022) <- my_column_names
+colnames(temp_2002_2022) <- my_column_names
+colnames(temp_2002_2022) <- my_column_names
+colnames(temp_2002_2022) <- my_column_names
+colnames(temp_2002_2022) <- my_column_names
+colnames(temp_2002_2022) <- my_column_names
+colnames(temp_2002_2022) <- my_column_names
+colnames(temp_2002_2022) <- my_column_names
+colnames(temp_2002_2022) <- my_column_names
+colnames(temp_2002_2022) <- my_column_names
+colnames(temp_2002_2022) <- my_column_names
+colnames(temp_2002_2022) <- my_column_names
+colnames(temp_2002_2022) <- my_column_names
+colnames(temp_2002_2022) <- my_column_names
+colnames(temp_2002_2022) <- my_column_names
+
 
 # Combine 2002 - 2022 data
 

### 2022-06-29 scripts/01-data_cleaning.R
Update 01-data_cleaning.R

@@ -30,25 +30,27 @@ my_column_names <- c("Longitude (x)", "Latitude (y)", "Station Name", "Climate I
 
 colnames(temp_jan_may_2022) <- my_column_names
 colnames(temp_jan_dec_2021) <- my_column_names
-colnames(temp_2002_2022) <- my_column_names
-colnames(temp_2002_2022) <- my_column_names
-colnames(temp_2002_2022) <- my_column_names
-colnames(temp_2002_2022) <- my_column_names
-colnames(temp_2002_2022) <- my_column_names
-colnames(temp_2002_2022) <- my_column_names
-colnames(temp_2002_2022) <- my_column_names
-colnames(temp_2002_2022) <- my_column_names
-colnames(temp_2002_2022) <- my_column_names
-colnames(temp_2002_2022) <- my_column_names
-colnames(temp_2002_2022) <- my_column_names
-colnames(temp_2002_2022) <- my_column_names
-colnames(temp_2002_2022) <- my_column_names
-colnames(temp_2002_2022) <- my_column_names
-colnames(temp_2002_2022) <- my_column_names
-
+colnames(temp_jan_dec_2020) <- my_column_names
+colnames(temp_jan_dec_2019) <- my_column_names
+colnames(temp_jan_dec_2018) <- my_column_names
+colnames(temp_jan_dec_2017) <- my_column_names
+colnames(temp_jan_dec_2016) <- my_column_names
+colnames(temp_jan_dec_2015) <- my_column_names
+colnames(temp_jan_dec_2014) <- my_column_names
+colnames(temp_jan_dec_2013) <- my_column_names
+colnames(temp_jan_dec_2012) <- my_column_names
+colnames(temp_jan_dec_2011) <- my_column_names
+colnames(temp_jan_dec_2010) <- my_column_names
+colnames(temp_jan_dec_2009) <- my_column_names
+colnames(temp_jan_dec_2008) <- my_column_names
+colnames(temp_jan_dec_2007) <- my_column_names
+colnames(temp_jan_dec_2006) <- my_column_names
+colnames(temp_jan_dec_2005) <- my_column_names
+colnames(temp_jan_dec_2004) <- my_column_names
+colnames(temp_jan_dec_2003) <- my_column_names
+colnames(temp_jan_dec_2002) <- my_column_names
 
 # Combine 2002 - 2022 data
-
[…]

### 2022-06-29 scripts/00-download_data.R
Update 00-download_data.R

@@ -28,7 +28,7 @@ temp_jan_dec_2016 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_c
 temp_jan_dec_2015 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_climate_daily_ON_6158355_2015_P1D.csv")
 temp_jan_dec_2014 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_climate_daily_ON_6158355_2014_P1D.csv")
 temp_jan_dec_2013 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_climate_daily_ON_6158355_2013_P1D.csv")
-temp_jan_may_2012 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_climate_daily_ON_6158355_2012_P1D.csv")
+temp_jan_dec_2012 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_climate_daily_ON_6158355_2012_P1D.csv")
 temp_jan_dec_2011 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_climate_daily_ON_6158355_2011_P1D.csv")
 temp_jan_dec_2010 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_climate_daily_ON_6158355_2010_P1D.csv")
 temp_jan_dec_2009 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_climate_daily_ON_6158355_2009_P1D.csv")
@@ -41,27 +41,27 @@ temp_jan_dec_2003 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_c
 temp_jan_dec_2002 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_climate_daily_ON_6158355_2002_P1D.csv")
 
 # Rename files to be more readable
-file.rename("~/Desktop/SUDS2022/inputs/data/en_climate_daily_ON_6158355_2022_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_may_2022.csv")
[…]

### 2022-06-29 scripts/00-download_data.R
Update 00-download_data.R

@@ -64,7 +64,7 @@ file.rename("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_climate_daily_ON_61
 file.rename("~/Desktop/SUDS2022/inputs/data/raw_temp_data/en_climate_daily_ON_6158355_2002_P1D.csv", "~/Desktop/SUDS2022/inputs/data/temp_jan_dec_2002.csv")
 
 # Use below lines if running after the first time
-temp_jan_may_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/temp_jan_dec_2022.csv")
+temp_jan_may_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/temp_jan_may_2022.csv")
 temp_jan_dec_2021 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/temp_jan_dec_2021.csv")
 temp_jan_dec_2020 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/temp_jan_dec_2020.csv")
 temp_jan_dec_2019 <- read.csv("~/Desktop/SUDS2022/inputs/data/raw_temp_data/temp_jan_dec_2019.csv")

### 2022-06-29 scripts/01-data_cleaning.R
Cleaning pollution data

@@ -14,6 +14,7 @@ library(tidyverse)
 library(dplyr)
 library(ggplot2)
 library(janitor)
+library(naniar)
 
 #### Cleaning Temperature Data ----
 
@@ -59,9 +60,8 @@ temp_2002_2022 <- rbind(temp_jan_dec_2002, temp_jan_dec_2003, temp_jan_dec_2004,
                         temp_jan_dec_2017, temp_jan_dec_2018, temp_jan_dec_2019,
                         temp_jan_dec_2020, temp_jan_dec_2021, temp_jan_may_2022)
 
-temp_2002_2022 <- temp_2002_2022 %>% select()
-
-colnames(temp_2002_2022) <- my_column_names
+temp_2002_2022 <- temp_2002_2022 %>% select("Station Name", "Date", "Year", "Month", "Day", 
+                                            "Max Temp", "Min Temp", "Mean Temp")
 temp_2002_2022 <- clean_names(temp_2002_2022)
 
 ### Add column for Month Names
@@ -108,7 +108,6 @@ colnames(O3_2020) <- my_column_names
 colnames(SO2_2020) <- my_column_names
 # combine data
 pollu_2020 <- rbind(CO_2020, NO_2020, NO2_2020, O3_2020, SO2_2020)
-colnames(pollu_2020) <- my_column_names
 # select variables
 pollu_2020 <- pollu_2020 %>% select("Pollutant", "City", "Date", "H1",
                                     "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12",
@@ -137,6 +136,7 @@ pollu_2020 <- pollu_2020 %>% mutate(Year = "2020")
 # Remove hourly data columns
 pollu_2020 <- pollu_2020 %>% select("Pollutant", "City", "Date", "Mean", "Year")
 
+
 ## 2019 Data
 
 colnames(CO_2019) <- my_column_names
@@ -146,8 +146,6 @@ colnames(O3_2019) <- my_column_names
 colnames(SO2_2019) <- my_column_names
 # combine data
 pollu_2019 <- rbind(CO_2019, NO_2019, NO2_2019, O3_2019, SO2_2019)
-# change column names
-colnames(pollu_2019) <- my_column_names
 # select variables
 pollu_2019 <- pollu_2019 %>% select("Pollutant", "City", "Date", "H1",
                                     "H2", "H3", "H4", "H5", "H6", "H7", "H8", "H9", "H10", "H11", "H12",
[…]

### 2022-06-29 scripts/01-data_cleaning.R
Update 01-data_cleaning.R

@@ -810,9 +810,10 @@ pollution <- rbind(pollu_2002, pollu_2003, pollu_2004, pollu_2005, pollu_2006, p
 # Save pollution data 2002-2020
 write_csv(
   x = pollution,
-  file = "~/Desktop/SUDS2022/inputs/data/clean_temp_data/pollution_2002_2020.csv"
+  file = "~/Desktop/SUDS2022/inputs/data/clean_pollution_data/pollution_2002_2020.csv"
 )
 
 
 #### Combine Temperature and Pollution Data ----
+pollu_temp <- rbind(temp_2002_2022, pollution)
 

### 2022-06-29 scripts/01-data_cleaning.R
Cleaning temperature and pollution data and combining the datasets

@@ -19,7 +19,7 @@ library(naniar)
 #### Cleaning Temperature Data ----
 
 # Change column names 
-my_column_names <- c("Longitude (x)", "Latitude (y)", "Station Name", "Climate ID", 
+my_column_names <- c("Longitude (x)", "Latitude (y)", "City", "Climate ID", 
                      "Date", "Year", "Month", "Day", "Data Quality", "Max Temp", 
                      "Max Temp Flag", "Min Temp", "Min Temp Flag", "Mean Temp", 
                      "Mean Temp Flag", "Heat Deg Days", "Heat Deg Days Flag", 
@@ -60,25 +60,31 @@ temp_2002_2022 <- rbind(temp_jan_dec_2002, temp_jan_dec_2003, temp_jan_dec_2004,
                         temp_jan_dec_2017, temp_jan_dec_2018, temp_jan_dec_2019,
                         temp_jan_dec_2020, temp_jan_dec_2021, temp_jan_may_2022)
 
-temp_2002_2022 <- temp_2002_2022 %>% select("Station Name", "Date", "Year", "Month", "Day", 
+temp_2002_2022 <- temp_2002_2022 %>% select("City", "Date", "Year", "Month", "Day", 
                                             "Max Temp", "Min Temp", "Mean Temp")
 temp_2002_2022 <- clean_names(temp_2002_2022)
 
 ### Add column for Month Names
 temp_2002_2022 <- temp_2002_2022 %>%
   mutate(month_name = case_when(
-    month == 1 ~ "Jan",
-    month == 2 ~ "Feb", 
-    month == 3 ~ "Mar", 
-    month == 4 ~ "Apr",
-    month == 5 ~ "May", 
-    month == 6 ~ "Jun",
-    month == 7 ~ "Jul",
-    month == 8 ~ "Aug", 
-    month == 9 ~ "Sep",
-    month == 10 ~ "Oct",
-    month == 11 ~ "Nov", 
-    month == 12 ~ "Dec"
+    month == "01" ~ "Jan", 
+    month == "02" ~ "Feb", 
+    month == "03" ~ "Mar", 
+    month == "04" ~ "Apr", 
+    month == "05" ~ "May", 
+    month == "06" ~ "Jun", 
[…]

### 2022-06-29 scripts/01-data_cleaning.R
Update 01-data_cleaning.R

@@ -905,7 +905,12 @@ write_csv(
 #### Combine Temperature and Pollution Data ----
 pollu_temp <- dplyr::full_join(temp_2002_2022, pollution)
 
+pollu_temp <- pollu_temp %>% drop_na(pollutant)
+pollu_temp <- pollu_temp %>% select(city, pollutant, mean, max_temp, min_temp, mean_temp, 
+                                    day, month, year, month_name)
+
 write_csv(
   x = pollution,
   file = "~/Desktop/SUDS2022/inputs/data/clean_temperature_pollution.csv"
 )
+

### 2022-06-29 scripts/01-data_cleaning.R
Cleaning combined pollution and temperature data

@@ -910,7 +910,7 @@ pollu_temp <- pollu_temp %>% select(city, pollutant, mean, max_temp, min_temp, m
                                     day, month, year, month_name)
 
 write_csv(
-  x = pollution,
+  x = pollu_temp,
   file = "~/Desktop/SUDS2022/inputs/data/clean_temperature_pollution.csv"
 )
 

### 2022-06-29 scripts/02-data_analysis.R

@@ -46,9 +46,13 @@ aug_nov <- temperature_2002_2022 %>%
 ggplot(aug_nov, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
   labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
 
+""
+
+
 ## Pollution ----
 
 # Visualizing Pollution by Months
+
 
 
 

### 2022-06-29 scripts/plots.R

@@ -1,4 +1,4 @@
-#### ggplots ####
+#### temperature ggplots ####
 library(ggplot2)
 
 ggplot(temp_2002_2022, aes(x=year, y=max_temp, color=month)) + geom_point()
@@ -20,7 +20,7 @@ ggplot(temp_2002_2022, aes(x=month, y=mean_temp, color=year)) + geom_point()
 ggplot(temp_2002_2022, aes(x=month_name, y=mean_temp, col=year)) + geom_col()
 
 
-#### tables ####
+#### temperature tables ####
 library(knitr)
 install.packages("kableExtra")
 library(kableExtra)
@@ -75,7 +75,7 @@ temp_2002_2022 %>%
   kable_styling(latex_options = "HOLD_position", font_size = 10)
 
 
-# summarising data ----
+# temperature summarising data ----
 summary_max_data <- temp_2002_2022 %>%
   group_by(month_name) %>%
   summarise(mean_max_temp = mean(max_temp))
@@ -192,3 +192,12 @@ pollution %>%
     linesep = "", digits = 2, booktabs=TRUE) %>%
   kable_styling(latex_options = "HOLD_position", font_size = 10)
 
+
+
+# temperature and pollution data ----
+
+## ggplot ----
+
+
+
+## tables ----

### 2022-06-29 scripts/plots.R
Update plots.R

@@ -197,7 +197,25 @@ pollution %>%
 # temperature and pollution data ----
 
 ## ggplot ----
+pollu_temp %>% filter(pollutant != "CO") %>% ggplot(aes(max_temp, 
+                      mean, color=pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Maximum Temperature") + ylab("Mean Pollution")
+pollu_temp %>% filter(pollutant == "CO") %>% ggplot(aes(max_temp, 
+                      mean, color=pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Maximum Temperature") + ylab("Mean Pollution")
 
+
+pollu_temp %>% filter(pollutant != "CO") %>% ggplot(aes(min_temp, 
+                      mean, color=pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Minimum Temperature") + ylab("Mean Pollution")
+pollu_temp %>% filter(pollutant == "CO") %>% ggplot(aes(min_temp, 
+                      mean, color=pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Minimum Temperature") + ylab("Mean Pollution")
+
+pollu_temp %>% filter(month_name == "Dec" | month_name == "Jan" |
+  month_name == "Feb" | month_name == "Mar") %>% ggplot(aes(max_temp, 
+  mean, color=month_name)) + geom_point() + labs(color="Month") + 
+  xlab("Maximum Temperature") + ylab("Mean Pollution")
 
 
 ## tables ----

### 2022-06-30 scripts/01-data_cleaning.R
Updates to analysis

@@ -78,7 +78,16 @@ temp_2002_2022 <- temp_2002_2022 %>%
     month == "09" ~ "Sep", 
     month == "10" ~ "Oct", 
     month == "11" ~ "Nov", 
-    month == "12" ~ "Dec"
+    month == "12" ~ "Dec", 
+    month == "1" ~ "Jan", 
+    month == "2" ~ "Feb", 
+    month == "3" ~ "Mar", 
+    month == "4" ~ "Apr", 
+    month == "5" ~ "May", 
+    month == "6" ~ "Jun", 
+    month == "7" ~ "Jul", 
+    month == "8" ~ "Aug", 
+    month == "9" ~ "Sep", 
   ))
 
 # rename TORONTO CITY to Toronto
@@ -820,7 +829,8 @@ pollu_2002 <- pollu_2002 %>% select("Pollutant", "City", "Date", "Mean", "Year")
 pollution <- rbind(pollu_2002, pollu_2003, pollu_2004, pollu_2005, pollu_2006, pollu_2007, pollu_2008, pollu_2009, pollu_2010, pollu_2011, pollu_2012, pollu_2013, pollu_2014, pollu_2015, pollu_2016, 
                    pollu_2017, pollu_2018, pollu_2019, pollu_2020)
 
-substr("20020102", 7, 8)
+column_names <- c("pollutant", "city", "date", "mean_pollution", "year")
+colnames(pollution) <- column_names
 
 # Add column for Month
 pollution <- pollution %>%

### 2022-06-30 scripts/02-data_analysis.R

@@ -28,32 +28,45 @@ poll_2011_2020 <- read.csv("~/Desktop/SUDS2022/inputs/data/
 ## Temperature ----
 
 # Visualizing Temperature by Months
-dec_mar <- temperature_2002_2022 %>%
-  filter(month_name == "Dec" | month_name == "Jan" | month_name == "Feb" | month_name == "Mar")
+temp_2002_2022 %>% filter(month_name == "Dec" | month_name == "Jan"| 
+               month_name == "Feb" | month_name == "Mar") %>% ggplot(
+              aes(x=year, y=max_temp, color=month_name)) + 
+  geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
 
-ggplot(dec_mar, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
-  labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
+temp_2002_2022 %>% filter(month_name == "Apr" | month_name == "May" | 
+                            month_name == "Jun" | month_name == "Jul") %>% ggplot(
+                              aes(x=year, y=max_temp, color=month_name)) + 
+  geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
 
-apr_jul <- temperature_2002_2022 %>%
-  filter(month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul")
-
-ggplot(apr_jul, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
-  labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
-
-aug_nov <- temperature_2002_2022 %>%
-  filter(month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")
-
-ggplot(aug_nov, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
-  labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
+temp_2002_2022 %>% filter(month_name == "Aug" | month_name == "Sep" | 
+                            month_name == "Oct" | month_name == "Nov") %>% ggplot(
+                              aes(x=year, y=max_temp, color=month_name)) + 
+  geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
 
 ""
 
 
 ## Pollution ----
 
[…]

### 2022-06-30 scripts/plots.R

@@ -169,25 +169,11 @@ pollution %>% filter(Pollutant != "CO") %>% ggplot(
   aes(x = Year, y = Mean, color = Pollutant)) + geom_point()
 
 ## tables ----
-pollution_2011_2020 %>%
-  select(Year, Pollutant, January, February, March, April, May, June, July, August, September, October, 
-         November, December) %>%
-  slice(c(2, 34, 74, 91, 117, 147, 184, 223, 252, 308)) %>%
-  kable(
-    caption = "Extracting rows from the Pollution data from 2011 - 2020",
-    col.names=c("Year", "Pollutant", "Jan", "Feb", "Mar", "Apr", "May",
-                "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"),
-    linesep = "", digits = 2, booktabs=TRUE) %>%
-  kable_styling(latex_options = "HOLD_position", font_size = 10)
-
-
-
-
 pollution %>%
   select(Pollutant, City, Date, Mean, Year) %>%
   slice(c(100, 1000, 1200, 1500, 1800, 2000, 2500, 2800, 3600, 5000, 6000, 7000, 10000, 11000, 15000)) %>%
   kable(
-    caption = "Extracting rows from the Pollution data from 2010 - 2020",
+    caption = "Extracting rows from the Pollution data from 2002 - 2020",
     col.names=c("Pollutant", "City", "Date", "Mean", "Year"),
     linesep = "", digits = 2, booktabs=TRUE) %>%
   kable_styling(latex_options = "HOLD_position", font_size = 10)
@@ -219,3 +205,14 @@ pollu_temp %>% filter(month_name == "Dec" | month_name == "Jan" |
 
 
 ## tables ----
+
+pollu_temp %>%
+  select(city, pollutant, mean, max_temp, min_temp, mean_temp, day, month_name, year) %>%
+  slice(c(100, 2800, 3600, 5000, 6000, 7000, 10000, 11000, 15000, 18000, 20000, 
+          30000)) %>%
+  kable(
+    caption = "Extracting rows from the Temperature and Pollution data from 2002 - 2020",
+    col.names=c("City", "Pollutant", "Mean Pollution", "Max Temp", 
+                "Min Temp", "Mean Temp", "Day", "Month", "Year"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)

### 2022-06-30 scripts/02-data_analysis.R
Update 02-data_analysis.R

@@ -68,6 +68,21 @@ pollution %>% filter(pollutant == "CO") %>% ggplot(
   aes(x = month_name, y = mean_pollution, color = pollutant)) + geom_point() + 
   labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
 
+## Pollution and Temperature ----
+
+pollu_temp %>% filter(pollutant != "CO") %>% ggplot(aes(max_temp, 
+                                                        mean, color=pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Maximum Temperature") + ylab("Mean Pollution")
+pollu_temp %>% filter(pollutant == "CO") %>% ggplot(aes(max_temp, 
+                                                        mean, color=pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Maximum Temperature") + ylab("Mean Pollution")
+
+pollu_temp %>% filter(pollutant != "CO") %>% ggplot(aes(month_name, 
+                                                        max_temp, color=pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Month") + ylab("Maximum Temperature")
+pollu_temp %>% filter(pollutant == "CO") %>% ggplot(aes(month_name, 
+                                                        max_temp, color=pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Month") + ylab("Maximum Temperature")
 
 ### Models ----
 

### 2022-06-30 scripts/02-data_analysis.R
Update 02-data_analysis.R

@@ -70,6 +70,8 @@ pollution %>% filter(pollutant == "CO") %>% ggplot(
 
 ## Pollution and Temperature ----
 
+#Visualizing Mean pollution by temperature and pollutant and month
+
 pollu_temp %>% filter(pollutant != "CO") %>% ggplot(aes(max_temp, 
                                                         mean, color=pollutant)) + geom_point() + 
   labs(color="Pollutant") + xlab("Maximum Temperature") + ylab("Mean Pollution")
@@ -77,12 +79,48 @@ pollu_temp %>% filter(pollutant == "CO") %>% ggplot(aes(max_temp,
                                                         mean, color=pollutant)) + geom_point() + 
   labs(color="Pollutant") + xlab("Maximum Temperature") + ylab("Mean Pollution")
 
-pollu_temp %>% filter(pollutant != "CO") %>% ggplot(aes(month_name, 
-                                                        max_temp, color=pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Month") + ylab("Maximum Temperature")
-pollu_temp %>% filter(pollutant == "CO") %>% ggplot(aes(month_name, 
-                                                        max_temp, color=pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Month") + ylab("Maximum Temperature")
+pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & 
+  (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
+  month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, 
+  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
+  labs(color="Month") + xlab("Maximum Temperature") + 
+  ylab("Mean Pollution")
+
+pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & 
+  (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
+  month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, 
+  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
+  labs(color="Month") + xlab("Maximum Temperature") + 
+  ylab("Mean Pollution")
+
+pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & 
+                        (month_name == "Apr" | month_name == "May" | 
+                           month_name == "Jun" | month_name == "Jul")) %>% ggplot(aes(max_temp, mean_pollution, 
[…]

### 2022-06-30 outputs/paper/paper.Rmd
Updates to paper

--- 2022-06-27 outputs/paper/paper.qmd
@@
 This data includes the city (which is Toronto in our case), the levels of each pollutant for each hour of every day.
+I reduced the data to the mean of pollution levels by pollutant of each day.
 
@@
 
--   Date
+-   Day
 
@@
 
-temperature_2002_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/clean_temp_data/temperature_2002_2022.csv")
+temp_2002_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/clean_temp_data/temperature_2002_2022.csv")
 
@@
 
-temperature_2002_2022 %>%
+temp_2002_2022 %>%
   select(year, month, day, max_temp, min_temp, mean_temp) %>%
@@
 
-poll_2011_2020 <- read.csv("~/Desktop/SUDS2022/inputs/data/clean_pollution_data/pollution_2011_2020.csv")
+pollution <- read.csv("~/Desktop/SUDS2022/inputs/data/clean_pollution_data/pollution_2002_2020.csv")
 
@@
 
-poll_2011_2020 %>%
-  select(Year, Pollutant, January, February, March, April, May, June, July, August, September, October, 
-         November, December) %>%
-  slice(c(2, 34, 74, 91, 117, 147, 184, 223, 252, 308)) %>%
+pollution %>%
+  select(Pollutant, City, Date, Mean, Year) %>%
+  slice(c(100, 1000, 1200, 1500, 1800, 2000, 2500, 2800, 3600, 5000, 6000, 7000, 10000, 11000, 15000)) %>%
   kable(
-    caption = "Extracting rows from the Pollution data from 2011 - 2020",
-    col.names=c("Year", "Pollutant", "Jan", "Feb", "Mar", "Apr", "May",
-                "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"),
+    caption = "Extracting rows from the Pollution data from 2002 - 2020",
+    col.names=c("Pollutant", "City", "Date", "Mean", "Year"),
     linesep = "", digits = 2, booktabs=TRUE) %>%
@@
 
-#### Visualizing Temperature by Months
+#### Visualizing Temperature by Month and Year
 
@@
 
+In the above plot...
+
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2022 for April to July", fig.width=9, fig.height=3.2}
@@
 ```
+
+In the above plot...
 
@@
 
+In the above plot...
+
+## Visualizing Pollution from 2002-2020
+
+#### Visualizing Pollution by Year
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 by Year", fig.width=9, fig.height=3.2}
+
+pollution %>% filter(pollutant != "CO") %>% ggplot(
+  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)")
+```
+
+In the above plot...
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Year", fig.width=9, fig.height=3.2}
+
+pollution %>% filter(pollutant == "CO") %>% ggplot(
+  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")
+
+```
+
+In the above plot...
+
+#### Visualizing Pollution by Month
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 by Month", fig.width=9, fig.height=3.2}
+
+pollution %>% filter(pollutant != "CO") %>% ggplot(
+  aes(x = month_name, y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppb)")
+
+```
+
+In the above plot...
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Month", fig.width=9, fig.height=3.2}
+
+pollution %>% filter(pollutant == "CO") %>% ggplot(
+  aes(x = month_name, y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
+
+```
+
+In the above plot...
+
+## Visualizing Pollution and Temperature from 2002-2020
+
+#### Visualizing Mean Pollution by Temperature, Pollutant and Month
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for December to March", fig.width=9, fig.height=3.2}
+
+pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & 
+  (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
+  month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, 
+  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
+  labs(color="Month") + xlab("Maximum Temperature") + 
+  ylab("Mean Pollution")
+```
+
+In the above plot...
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO and SO2 by Month, Maximum Temperature, and Pollutant for December to March", fig.width=9, fig.height=3.2}
+
+pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & 
+  (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
+  month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, 
+  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
+  labs(color="Month") + xlab("Maximum Temperature") + 
+  ylab("Mean Pollution")
+
+```
+
+In the above plot...
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for April to July", fig.width=9, fig.height=3.2}
+
+pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & 
+  (month_name == "Apr" | month_name == "May" | month_name == "Jun" | 
+  month_name == "Jul")) %>% ggplot(aes(max_temp, mean_pollution, 
+  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
+  labs(color="Month") + xlab("Maximum Temperature") + 
+  ylab("Mean Pollution")
+
+```
+
+In the above plot...
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO and SO2 by Month, Maximum Temperature, and Pollutant for April to July", fig.width=9, fig.height=3.2}
+
+pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & (month_name == "Apr" |
+   month_name == "May" | month_name == "Jun" | month_name == "Jul")) %>% ggplot(aes(max_temp,
+   mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
+  labs(color="Month") + xlab("Maximum Temperature") + 
+  ylab("Mean Pollution")
+
+```
+
+In the above plot...
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for August to November", fig.width=9, fig.height=3.2}
+
+pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & (month_name == "Aug" | 
+    month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>% ggplot(aes(max_temp,
+    mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
+    labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
+
+```
+
+In the above plot...
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO and SO2 by Month, Maximum Temperature, and Pollutant for August to November", fig.width=9, fig.height=3.2}
+
+pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & 
+  (month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>%
+  ggplot(aes(max_temp, mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~
+  pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
+
+```
+
+In the above plot...
+
 ## Limitations

### 2022-06-30 scripts/02-data_analysis.R
Updates to paper

@@ -72,13 +72,6 @@ pollution %>% filter(pollutant == "CO") %>% ggplot(
 
 #Visualizing Mean pollution by temperature and pollutant and month
 
-pollu_temp %>% filter(pollutant != "CO") %>% ggplot(aes(max_temp, 
-                                                        mean, color=pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Maximum Temperature") + ylab("Mean Pollution")
-pollu_temp %>% filter(pollutant == "CO") %>% ggplot(aes(max_temp, 
-                                                        mean, color=pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Maximum Temperature") + ylab("Mean Pollution")
-
 pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & 
   (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
   month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, 

### 2022-06-30 scripts/02-data_analysis.R
Update 02-data_analysis.R

@@ -117,6 +117,8 @@ pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") &
 
 ### Models ----
 
+
+
 ### Anova Tests ----
 
 ### t-tests ----

### 2022-06-30 outputs/paper/paper.Rmd
Updates to paper

@@
 pollution %>%
-  select(Pollutant, City, Date, Mean, Year) %>%
+  select(city, pollutant, mean_pollution, day, month, year) %>%
   slice(c(100, 1000, 1200, 1500, 1800, 2000, 2500, 2800, 3600, 5000, 6000, 7000, 10000, 11000, 15000)) %>%
@@
     caption = "Extracting rows from the Pollution data from 2002 - 2020",
-    col.names=c("Pollutant", "City", "Date", "Mean", "Year"),
+    col.names=c("City", "Pollutant", "Mean Pollution", "Day", "Month", "Year"),
     linesep = "", digits = 2, booktabs=TRUE) %>%

### 2022-06-30 outputs/paper/presentation.qmd
Updates to paper

---
title: "Untitled"
format: pptx
editor: visual
---

## Quarto

Quarto enables you to weave together content and executable code into a finished presentation. To learn more about Quarto presentations see <https://quarto.org/docs/presentations/>.

## Bullets

When you click the **Render** button a document will be generated that includes:

-   Content authored with markdown
-   Output from executable code

## Code

When you click the **Render** button a presentation will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

### 2022-07-06 outputs/paper/paper.qmd
Updates

@@
 
+Urgent action is needed and the science shows that it will make a tangible difference.
+The health impacts from climate change can be prevented by improving efforts both in societies and health systems in Canada.
+Changes in measures can prevent health impacts and build health systems and facilities that are more resilient to climate change.
+
 ## Data
@@
 
+Table 1 shows the rows of the Temperature data from 2002 to 2022.
+Variable "Year" indicates the year, variable "Month" indicates the month, and variable "Day" indicates the day.
+Variable "Max Temp" represents the maximum temperature on the day with the following month and year.
+Variable "Min Temp" represents the minimum temperature on the day with the following month and year.
+Variable "Mean Temp" represents the mean temperature on the day with the following month and year.
+
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
@@
 
-poll_2011_2020 %>%
-  select(Year, Pollutant, January, February, March, April, May, June, July, August, September, October, 
-         November, December) %>%
-  slice(c(2, 34, 74, 91, 117, 147, 184, 223, 252, 308)) %>%
+pollution %>%
+  select(city, pollutant, mean_pollution, day, month, year) %>%
+  slice(c(100, 1000, 1200, 1500, 1800, 2000, 2500, 2800, 3600, 5000, 6000, 7000, 10000, 11000, 15000)) %>%
   kable(
-    caption = "Extracting rows from the Pollution data from 2011 - 2020",
-    col.names=c("Year", "Pollutant", "Jan", "Feb", "Mar", "Apr", "May",
-                "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"),
+    caption = "Extracting rows from the Pollution data from 2002 - 2020",
+    col.names=c("City", "Pollutant", "Mean Pollution", "Day", "Month", "Year"),
     linesep = "", digits = 2, booktabs=TRUE) %>%
@@
 
+Table 2 shows the rows of the Pollution data from 2002 to 2020.
+Variable "City" indicates the city from which the data was collected.
+Variable "Pollutant" indicates the pollutant and includes 5 types: CO, NO, NO2, O3, and SO2.
+Variable "Mean Pollution" represents the mean pollution levels of the following pollutant on the following day of the month and year.
+Variable "Day" indicates the day, variable "Month" indicates the month, and variable "Year" indicates the year.
+Variable "Max Temp" represents the maximum temperature on the day with the following month and year.
+
 ## Visualizing Temperature from 2002-2022
 
-#### Visualizing Temperature by Months
+#### Visualizing Temperature by Month and Year
 
@@
 
-dec_mar <- temperature_2002_2022 %>%
-  filter(month_name == "Dec" | month_name == "Jan" | month_name == "Feb" | month_name == "Mar")
-
-ggplot(dec_mar, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
-  labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
+temp_2002_2022 %>% filter(month_name == "Dec" | month_name == "Jan"| 
+               month_name == "Feb" | month_name == "Mar") %>% ggplot(
+              aes(x=year, y=max_temp, color=month_name)) + 
+  geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
 
@@
 
+Figure 1 shows an increasing pattern for the maximum temperature from year 2002 to 2022 from December to March.
+
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2022 for April to July", fig.width=9, fig.height=3.2}
 
-apr_jul <- temperature_2002_2022 %>%
-  filter(month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul")
-
-ggplot(apr_jul, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
-  labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
+temp_2002_2022 %>% filter(month_name == "Apr" | month_name == "May" | 
+                            month_name == "Jun" | month_name == "Jul") %>% ggplot(
+                              aes(x=year, y=max_temp, color=month_name)) + 
+  geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
 
@@
 
+Figure 2 shows an increasing pattern for the maximum temperature from year 2002 to 2022 from April to July.
+
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2022 for August to November", fig.width=9, fig.height=3.2}
 
-aug_nov <- temperature_2002_2022 %>%
-  filter(month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")
-
-ggplot(aug_nov, aes(x=year, y=max_temp, color=month_name)) + geom_point() + 
-  labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
+temp_2002_2022 %>% filter(month_name == "Aug" | month_name == "Sep" | 
+                            month_name == "Oct" | month_name == "Nov") %>% ggplot(
+                              aes(x=year, y=max_temp, color=month_name)) + 
+  geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
 
 ```
+
+Figure 3 shows an increasing pattern for the maximum temperature from year 2002 to 2022 from August to November.
+
+## Visualizing Pollution from 2002-2020
+
+#### Visualizing Pollution by Year
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 by Year", fig.width=9, fig.height=3.2}
+
+pollution %>% filter(pollutant != "CO") %>% ggplot(
+  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)")
+```
+
+Figure 4 shows an ... pattern for the mean pollution from year 2002 to 2020 for NO, NO2, O3, and SO2 by year.
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Year", fig.width=9, fig.height=3.2}
+
+pollution %>% filter(pollutant == "CO") %>% ggplot(
+  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")
+
+```
+
+Figure 5 shows an ... pattern for the mean pollution from year 2002 to 2020 for CO by year.
+
+#### Visualizing Pollution by Month
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 by Month", fig.width=9, fig.height=3.2}
+
+pollution %>% filter(pollutant != "CO") %>% ggplot(
+  aes(x = month_name, y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppb)")
+
+```
+
+Figure 6 shows an ... pattern for the mean pollution from year 2002 to 2020 for NO, NO2, O3, and SO2 by month.
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Month", fig.width=9, fig.height=3.2}
+
+pollution %>% filter(pollutant == "CO") %>% ggplot(
+  aes(x = month_name, y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
+
+```
+
+Figure 7 shows an ... pattern for the mean pollution from year 2002 to 2020 for CO by month.
+
+## Visualizing Pollution and Temperature from 2002-2020
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+
+# Load pollution and temperature data
+
+pollu_temp <- read.csv("~/Desktop/SUDS2022/inputs/data/clean_temperature_pollution.csv")
+```
+
+#### Visualizing Mean Pollution by Temperature, Pollutant and Month
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for December to March", fig.width=9, fig.height=3.2}
+
+pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & 
+  (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
+  month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, 
+  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
+  labs(color="Month") + xlab("Maximum Temperature") + 
+  ylab("Mean Pollution")
+```
+
+In the above plot...
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO and SO2 by Month, Maximum Temperature, and Pollutant for December to March", fig.width=9, fig.height=3.2}
+
+pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & 
+  (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
+  month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, 
+  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
+  labs(color="Month") + xlab("Maximum Temperature") + 
+  ylab("Mean Pollution")
+
+```
+
+In the above plot...
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for April to July", fig.width=9, fig.height=3.2}
+
+pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & 
+  (month_name == "Apr" | month_name == "May" | month_name == "Jun" | 
+  month_name == "Jul")) %>% ggplot(aes(max_temp, mean_pollution, 
+  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
+  labs(color="Month") + xlab("Maximum Temperature") + 
+  ylab("Mean Pollution")
+
+```
+
+In the above plot...
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO and SO2 by Month, Maximum Temperature, and Pollutant for April to July", fig.width=9, fig.height=3.2}
+
+pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & (month_name == "Apr" |
+   month_name == "May" | month_name == "Jun" | month_name == "Jul")) %>% ggplot(aes(max_temp,
+   mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
+  labs(color="Month") + xlab("Maximum Temperature") + 
+  ylab("Mean Pollution")
+
+```
+
+In the above plot...
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for August to November", fig.width=9, fig.height=3.2}
+
+pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & (month_name == "Aug" | 
+    month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>% ggplot(aes(max_temp,
+    mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
+    labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
+
+```
+
+In the above plot...
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO and SO2 by Month, Maximum Temperature, and Pollutant for August to November", fig.width=9, fig.height=3.2}
+
+pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & 
+  (month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>%
+  ggplot(aes(max_temp, mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~
+  pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
+
+```
+
+In the above plot...
 

### 2022-07-06 scripts/01-data_cleaning.R

@@ -902,7 +902,7 @@ pollution <- pollution %>%
     month == 12 ~ "Dec"
   ))
 
-pollution <- pollution %>% select(City, Pollutant, Mean, day, month, Year, month_name)
+pollution <- pollution %>% select(City, Pollutant, mean_pollution, day, month, Year, month_name)
 pollution <- clean_names(pollution)
 
 # Save pollution data 2002-2020
@@ -916,7 +916,7 @@ write_csv(
 pollu_temp <- dplyr::full_join(temp_2002_2022, pollution)
 
 pollu_temp <- pollu_temp %>% drop_na(pollutant)
-pollu_temp <- pollu_temp %>% select(city, pollutant, mean, max_temp, min_temp, mean_temp, 
+pollu_temp <- pollu_temp %>% select(city, pollutant, mean_pollution, max_temp, min_temp, mean_temp, 
                                     day, month, year, month_name)
 
 write_csv(

### 2022-07-06 scripts/01-data_cleaning.R
Update 01-data_cleaning.R

@@ -152,11 +152,11 @@ pollu_2020 <- pollu_2020 %>% replace_with_na(replace = list(H1 = -999, H2 = -999
 # replace NA with 0
 pollu_2020[is.na(pollu_2020)] <- 0
 # Add column for average 
-pollu_2020 <- pollu_2020 %>% mutate(Mean = rowMeans(pollu_2020[ ,c(4:27)], na.rm=TRUE))
+pollu_2020 <- pollu_2020 %>% mutate(Mean_pollution = rowMeans(pollu_2020[ ,c(4:27)], na.rm=TRUE))
 # Add column for Year
 pollu_2020 <- pollu_2020 %>% mutate(Year = "2020")
 # Remove hourly data columns
-pollu_2020 <- pollu_2020 %>% select("Pollutant", "City", "Date", "Mean", "Year")
+pollu_2020 <- pollu_2020 %>% select("Pollutant", "City", "Date", "Mean_pollution", "Year")
 
 
 ## 2019 Data
@@ -190,11 +190,11 @@ pollu_2019 <- pollu_2019 %>% replace_with_na(replace = list(H1 = -999, H2 = -999
 # replace NA with 0
 pollu_2019[is.na(pollu_2019)] <- 0
 # Add column for average 
-pollu_2019 <- pollu_2019 %>% mutate(Mean = rowMeans(pollu_2019[ ,c(4:27)], na.rm=TRUE))
+pollu_2019 <- pollu_2019 %>% mutate(Mean_pollution = rowMeans(pollu_2019[ ,c(4:27)], na.rm=TRUE))
 # Add column for Year
 pollu_2019 <- pollu_2019 %>% mutate(Year = "2019")
 # Remove hourly data columns
-pollu_2019 <- pollu_2019 %>% select("Pollutant", "City", "Date", "Mean", "Year")
+pollu_2019 <- pollu_2019 %>% select("Pollutant", "City", "Date", "Mean_pollution", "Year")
 
 ## 2018 Data
 colnames(CO_2018) <- my_column_names
@@ -226,11 +226,11 @@ pollu_2018 <- pollu_2018 %>% replace_with_na(replace = list(H1 = -999, H2 = -999
 # replace NA with 0
 pollu_2018[is.na(pollu_2018)] <- 0
 # Add column for average 
-pollu_2018 <- pollu_2018 %>% mutate(Mean = rowMeans(pollu_2018[ ,c(4:27)], na.rm=TRUE))
[…]

### 2022-07-06 outputs/paper/presentation.qmd
Updates to paper.qmd
Updates to presentation

@@
 ---
-title: "Untitled"
+#### Preamble ####
+title: "Disparities in Climate-Induced Health Outcomes in the Greater Toronto Area."
+author: "Sarah Mansoor"
+email: "sarah.mansoor@mail.utoronto.ca"
+thanks: "Code and data are available at: https://github.com/sarahmansoorr/SUDS2022"
+date: "`r format(Sys.time(), '%d %B %Y')`"
 format: pptx
 editor: visual
+toc: true
+# bibliography: references.bib
 ---
@@
 ```
+
+![](images/paste-01E2A963.png)

### 2022-07-06 scripts/02-data_analysis.R
Update 02-data_analysis.R

@@ -43,6 +43,12 @@ temp_2002_2022 %>% filter(month_name == "Aug" | month_name == "Sep" |
                               aes(x=year, y=max_temp, color=month_name)) + 
   geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
 
+# Lowest and highest points
+r <- temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| 
+                            month_name == "Feb" | month_name == "Mar") & year == 2015) %>% select(max_temp)
+range(r)
+"-19.2  14.4"
+
 ""
 
 

### 2022-07-06 scripts/02-data_analysis.R
Update 02-data_analysis.R

@@ -49,6 +49,11 @@ r <- temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"|
 range(r)
 "-19.2  14.4"
 
+r <- temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| 
+                                  month_name == "Feb" | month_name == "Mar") & year == 2012) %>% select(max_temp)
+range(r)
+"-10.4  24.3"
+
 ""
 
 

### 2022-07-06 outputs/paper/paper.qmd
Update paper.qmd

@@
 Figure 1 shows an increasing pattern for the maximum temperature from year 2002 to 2022 from December to March.
+Temperatures in March continue to be increasingly higher than any of the other months in this plot.
+It is noticeable that the temperatures for all months are the highest in 2012.
+According to NASA, the average temperatures in 2012 was warmer than the mid-20th century baseline and is considered to be an outlying of seasonal extremes that are warmer.
+Even the lowest maximum temperature in January is higher than other years in 2012.
+The plot shows that the average maximum temperature from December to March has been between -19.2 degrees Celsius in February of 2015 and 24.3 degrees Celsius in March of 2012.
 

### 2022-07-06 outputs/paper/paper.qmd
Updates to paper

@@
 Figure 2 shows an increasing pattern for the maximum temperature from year 2002 to 2022 from April to July.
+Temperatures in July are consistently higher throughout the years than any of the other months in the plot and temperatures in April are consistently lower throughout the years than any of the other months.
+There are some years where June has the highest temperature.
+The plot shows that the average maximum temperature from April to May has been between -2.4 degrees Celsius in April of 2003 and 38.2 degrees Celsius in July of 2011.
 

### 2022-07-06 scripts/02-data_analysis.R

@@ -54,6 +54,16 @@ r <- temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"|
 range(r)
 "-10.4  24.3"
 
+r <- temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | 
+                                  month_name == "Jun" | month_name == "Jul") & year == 2003) %>% select(max_temp)
+range(r)
+"-2.4 33.6"
+
+r <- temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | 
+                                  month_name == "Jun" | month_name == "Jul") & year == 2011) %>% select(max_temp)
+range(r)
+"3.3 38.2"
+
 ""
 
 

### 2022-07-06 outputs/paper/paper.qmd
Update paper.qmd
Updates to paper

@@
 
+I'm interested in finding the trends of the maximum temperatures from 2002 to 2022 by the Month and Year.
+
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2022 for December to March", fig.width=9, fig.height=3.2}
@@
 Figure 3 shows an increasing pattern for the maximum temperature from year 2002 to 2022 from August to November.
+Temperatures in both August and September are the highest throughout the years compared to November and October.
+Temperatures in November are consistently the lowest throughout the years compared to the other months.
+There are a few high temperatures in October throughout the years as well and this is typically seen when the August and September have some of the highest temperatures compared to other years.
+The plot shows that the average maximum temperature from April to May has been between -5.1 degrees Celsius in November of 2018 and 36.8 degrees Celsius in August of 2006.
 

### 2022-07-06 scripts/02-data_analysis.R

@@ -64,6 +64,16 @@ r <- temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" |
 range(r)
 "3.3 38.2"
 
+r <- temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | 
+                                  month_name == "Oct" | month_name == "Nov") & year == 2018) %>% select(max_temp)
+range(r)
+"-5.1 32.3"
+
+r <- temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | 
+                                  month_name == "Oct" | month_name == "Nov") & year == 2006) %>% select(max_temp)
+range(r)
+"2.0 36.8"
+
 ""
 
 

### 2022-07-06 outputs/paper/paper.qmd
Update paper.qmd

@@
 
-I'm interested in finding the trends of the maximum temperatures from 2002 to 2022 by the Month and Year.
+I'm interested in finding the trends of the maximum temperatures from 2002 to 2022 by the month and year.
 
@@
 
+I'm interested in finding the trends of the mean pollution levels from 2002 to 2020 by year.
+The units of NO, NO2, O3, and SO2 are the same (ppb) whereas the units for CO are different (ppm).
+For this reason, these plots have been made with regards to those units.
+
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 by Year", fig.width=9, fig.height=3.2}
@@
 
-Figure 4 shows an ... pattern for the mean pollution from year 2002 to 2020 for NO, NO2, O3, and SO2 by year.
+Figure 4 shows an decreasing pattern for the mean pollution from year 2002 to 2020 for NO by year. We can notice that O3 and SO2 levels remain mostly the same throughout the years.
 
@@
 
-Figure 5 shows an ... pattern for the mean pollution from year 2002 to 2020 for CO by year.
+Figure 5 shows a decreasing pattern for the mean pollution from year 2002 to 2020 for CO by year.
 

### 2022-07-07 outputs/paper/paper.qmd
Update paper.qmd

@@
 
+-   City
+
+-   Maximum Temperature
+
+-   Minimum Temperature
+
+-   Mean Temperature
+
 -   Day
-
--   Maximum temperature
-
--   Minimum temperature
-
--   Mean temperature.
-
-Pollution variables of interest:
-
--   Pollutant
 
@@
 
--   Mean level of pollution each day
+Pollution variables of interest:
+
+-   City
+
+-   Pollutant
+
+-   Mean pollution for pollutant each day
+
+-   Day
+
+-   Month
+
+-   Year
 
 #### Data Processing
+
+*Temperature data*
+
+The data for temperature was processed using R studio and R.
+I started off by renaming the column names for each temperature data set from 2002 to 2022.
+Once this was done I combined the 2002 to 2022 temperature data together.
+The data set initially contained 31 variables.
+From these I selected the city, date, year, month, day, maximum temperature, minimum temperature, and mean temperature using the dplyr package.
+I then cleaned the names of the data set using the janitor package.
+I added a column for the names of the month by mutating the data set and changed the city rows to represent only Toronto data.
+From here, I dropped any NA values from the maximum, minimum, and mean temperature columns.
+Lastly, I saved the cleaned data into the repository as a csv file.
+
+*Pollution Data*
 

### 2022-07-07 outputs/paper/paper.qmd
Update paper.qmd

@@
 *Pollution Data*
+
+The pollution data was also processed using R studio and R.
+I first loaded the different data sets for each pollutant for each year.
+Each year had 5 data sets for each pollutant (NO, NO2, CO, O3, SO2).
+I started off by also renaming the column names for the data sets from 2002 to 2020.
+Once this was done I combined the pollutant data sets for each year into one data frame for the year.
+The data initially contained 27 columns.
+The data sets initially contained pollution levels for each hour of every day in the respective year.
+I replaced these 24 columns (for each hour of the day) with one column that contained the mean pollution level for that day and pollutant.
 

### 2022-07-07 outputs/paper/paper.qmd
Update paper.qmd

@@
 I replaced these 24 columns (for each hour of the day) with one column that contained the mean pollution level for that day and pollutant.
+I filtered for Toronto from the list of cities in the data set using the dplyr package and removed the duplicate rows where the pollutant and date repeated.
+The missing values in this data set were labeled as "-999", so I replaced these values with NA and then with 0 for consistency.
+This process was repeated for all years from 2002 to 2020.
+Once this was completed, the data was combined for all the years into one data frame and the following columns were selected: city, pollutant, mean pollution, day, month, year.
+This cleaned data was then saved into the repository as a csv file.
 
-#### Population, Frame or Sample
+#### Sample
+
+The sample includes data from Toronto of daily temperature maximum, minimum and means from 2002 to 2022 and daily pollution levels for NO, NO2, CO, O3, SO2 from 2002 to 2020.
 

### 2022-07-07 outputs/paper/paper.qmd
Updates to paper
Update paper.qmd

@@
 
-## Results
+In this limitations section, I will focus on the data used and look into why it is not easy to take the implications mentioned above directly.
+
+### Missing Pollution Data from 2021-2022
+
+## Results and Discussion
 
@@
 
-## Conclusion
+Following on from the previous section, we can make good suggestions on ways to improve the data so that the findings will be less ambiguous and more ethical to use with confidence.
+First off, we can start by collecting pollution data from 2021 to 2022.
+It is possible that this data is available but not published yet on the website.
+In doing this, the data scientist who gets a hold of the data will have the liberty to conduct their analysis with the full data sets for temperature and pollution for 20 years.
+This will not only ease the job of the data scientist but will also improve the reproducibility of their workflow and finding.

### 2022-07-07 scripts/02-data_analysis.R
Updates to analysis

@@ -99,6 +99,23 @@ pollution %>% filter(pollutant == "CO") %>% ggplot(
   aes(x = month_name, y = mean_pollution, color = pollutant)) + geom_point() + 
   labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
 
+# Lowest and highest points
+r <- pollution %>% filter((pollutant == "NO")) %>% select(mean_pollution)
+range(r)
+#  0.0000 150.2917
+
+r <- pollution %>% filter((pollutant == "NO2")) %>% select(mean_pollution)
+range(r)
+#  0 65
+
+r <- pollution %>% filter((pollutant == "O3")) %>% select(mean_pollution)
+range(r)
+# 0.00000 65.95833
+
+r <- pollution %>% filter((pollutant == "SO2")) %>% select(mean_pollution)
+range(r)
+# -0.02500 22.79167
+
 ## Pollution and Temperature ----
 
 #Visualizing Mean pollution by temperature and pollutant and month

### 2022-07-07 outputs/paper/paper.qmd
Update paper.qmd

@@
 
-Figure 4 shows an decreasing pattern for the mean pollution from year 2002 to 2020 for NO by year. We can notice that O3 and SO2 levels remain mostly the same throughout the years.
+Figure 4 shows a decreasing pattern for the mean pollution from year 2002 to 2020 for NO by year.
+We can notice that O3 and SO2 levels remain mostly the same throughout the years.
+The highest levels of pollution are as follows: 150.3 ppb for NO in 2002, 65 ppb for NO2 in 2005, 65.95833 ppb for O3 in 2003, and 22.79167 ppb for SO2 in 2002.
 
@@
 Figure 5 shows a decreasing pattern for the mean pollution from year 2002 to 2020 for CO by year.
+The highest level of pollution for CO is 1.7625 ppm in 2002.
+We can notice that there is a drop in the pollution levels from 2003 to 2005 but it rises again in 2006.
+Looking back at figure 3, this was the same year some of the highest temperatures were recorded from August to November.
 
@@
 
+### Temperature
+
+### Pollution
+
+### Health
+
+### Temperature and Pollution
+
+### Temperature and Health
+
+### Pollution and Health
+
+### Temperature, Pollution and Health
+
 ## Next Steps

### 2022-07-07 scripts/02-data_analysis.R
Update 02-data_analysis.R

@@ -99,6 +99,13 @@ pollution %>% filter(pollutant == "CO") %>% ggplot(
   aes(x = month_name, y = mean_pollution, color = pollutant)) + geom_point() + 
   labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
 
+pollution1 <- pollution %>%
+  mutate(month_name = factor(month_name, levels = c("January", "February","March", 
+                                                    "April", "May", "June", "July", 
+                                                    "August","September", "October", 
+                                                    "November",  "December")))
+
+
 # Lowest and highest points
 r <- pollution %>% filter((pollutant == "NO")) %>% select(mean_pollution)
 range(r)
@@ -115,6 +122,10 @@ range(r)
 r <- pollution %>% filter((pollutant == "SO2")) %>% select(mean_pollution)
 range(r)
 # -0.02500 22.79167
+
+r <- pollution %>% filter((pollutant == "CO")) %>% select(mean_pollution)
+range(r)
+0.0000 1.7625
 
 ## Pollution and Temperature ----
 

### 2022-07-07 scripts/01-data_cleaning.R
Updating paper and cleaning

@@ -905,6 +905,9 @@ pollution <- pollution %>%
 pollution <- pollution %>% select(city, pollutant, mean_pollution, day, month, year, month_name)
 pollution <- clean_names(pollution)
 
+# drop NA
+pollution <- pollution %>% drop_na(month_name)
+
 # Save pollution data 2002-2020
 write_csv(
   x = pollution,

### 2022-07-07 scripts/01-data_cleaning.R
Updates to cleaning
Ordering the months in the right order

@@ -908,6 +908,12 @@ pollution <- clean_names(pollution)
 # drop NA
 pollution <- pollution %>% drop_na(month_name)
 
+pollution$month_name <- factor(pollution$month_name, levels=c("January", "February","March", 
+                                                              "April", "May", "June", "July", 
+                                                              "August","September", "October", 
+                                                              "November",  "December"))
+
+
 # Save pollution data 2002-2020
 write_csv(
   x = pollution,

### 2022-07-07 scripts/01-data_cleaning.R
Updates to month names

@@ -908,10 +908,10 @@ pollution <- clean_names(pollution)
 # drop NA
 pollution <- pollution %>% drop_na(month_name)
 
-pollution$month_name <- factor(pollution$month_name, levels=c("January", "February","March", 
-                                                              "April", "May", "June", "July", 
-                                                              "August","September", "October", 
-                                                              "November",  "December"))
+pollution$month_name <- factor(pollution$month_name, levels=c("Jan", "Feb","Mar", 
+                                                              "Apr", "May", "Jun", "Jul", 
+                                                              "Aug","Sep", "Oct", 
+                                                              "Nov",  "Dec"))
 
 
 # Save pollution data 2002-2020

### 2022-07-07 scripts/02-data_analysis.R

@@ -100,10 +100,10 @@ pollution %>% filter(pollutant == "CO") %>% ggplot(
   labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
 
 pollution1 <- pollution %>%
-  mutate(month_name = factor(month_name, levels = c("January", "February","March", 
-                                                    "April", "May", "June", "July", 
-                                                    "August","September", "October", 
-                                                    "November",  "December")))
+  mutate(month_name = factor(month_name, levels = c("Jan", "Feb","Mar", 
+                                                    "Apr", "May", "Jun", "Jul", 
+                                                    "Aug","Sep", "Oct", 
+                                                    "Nov",  "Dec")))
 
 
 # Lowest and highest points

### 2022-07-07 outputs/paper/paper.qmd
Updates to paper

@@
 
+level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
+
 pollution %>% filter(pollutant != "CO") %>% ggplot(
-  aes(x = month_name, y = mean_pollution, color = pollutant)) + geom_point() + 
+  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
   labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppb)")

### 2022-07-07 outputs/paper/paper.qmd
Updates to paper

@@
 
+level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
+
 pollution %>% filter(pollutant == "CO") %>% ggplot(
-  aes(x = month_name, y = mean_pollution, color = pollutant)) + geom_point() + 
+  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
   labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")

### 2022-07-07 outputs/paper/paper.qmd
Updates to paper

@@
 
-temp_2002_2022 %>% filter(month_name == "Dec" | month_name == "Jan"| 
-               month_name == "Feb" | month_name == "Mar") %>% ggplot(
-              aes(x=year, y=max_temp, color=month_name)) + 
-  geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
+level_order <- c("Dec", "Jan", "Feb","Mar")
+
+temp_2002_2022 %>% filter(month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
 
@@
 
-temp_2002_2022 %>% filter(month_name == "Apr" | month_name == "May" | 
-                            month_name == "Jun" | month_name == "Jul") %>% ggplot(
-                              aes(x=year, y=max_temp, color=month_name)) + 
-  geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
+level_order <- c("Apr", "May", "Jun", "Jul")
+
+temp_2002_2022 %>% filter(month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
 
@@
 
-temp_2002_2022 %>% filter(month_name == "Aug" | month_name == "Sep" | 
-                            month_name == "Oct" | month_name == "Nov") %>% ggplot(
-                              aes(x=year, y=max_temp, color=month_name)) + 
-  geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
+level_order <- c("Aug","Sep", "Oct","Nov")
+
+temp_2002_2022 %>% filter(month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
 

### 2022-07-07 outputs/paper/paper.qmd
Update paper.qmd

@@
 
-pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & 
-  (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
-  month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, 
-  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
-  labs(color="Month") + xlab("Maximum Temperature") + 
-  ylab("Mean Pollution")
+level_order <- c("Dec", "Jan", "Feb","Mar")
+
+pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
+~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") +  ylab("Mean Pollution")
 ```
@@
 
-pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & 
-  (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
-  month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, 
-  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
-  labs(color="Month") + xlab("Maximum Temperature") + 
-  ylab("Mean Pollution")
+level_order <- c("Dec", "Jan", "Feb","Mar")
+
+pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
+~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") +  ylab("Mean Pollution")
 
@@
 
-pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & 
-  (month_name == "Apr" | month_name == "May" | month_name == "Jun" | 
-  month_name == "Jul")) %>% ggplot(aes(max_temp, mean_pollution, 
-  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
-  labs(color="Month") + xlab("Maximum Temperature") + 
-  ylab("Mean Pollution")
+level_order <- c("Apr", "May", "Jun", "Jul")
+
+pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & (month_name == "Apr" | month_name == "May" | month_name == "Jun" |  month_name == "Jul")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
+~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
 
@@
 
-pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & (month_name == "Apr" |
-   month_name == "May" | month_name == "Jun" | month_name == "Jul")) %>% ggplot(aes(max_temp,
-   mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
-  labs(color="Month") + xlab("Maximum Temperature") + 
-  ylab("Mean Pollution")
+level_order <- c("Apr", "May", "Jun", "Jul")
+
+pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & (month_name == "Apr" |month_name == "May" | month_name == "Jun" | month_name == "Jul")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
+~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") +  ylab("Mean Pollution")
 
@@
 
-pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & (month_name == "Aug" | 
-    month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>% ggplot(aes(max_temp,
-    mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
-    labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
+level_order <- c("Aug","Sep", "Oct","Nov")
+
+pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & (month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
+~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
 
@@
 
-pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & 
-  (month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>%
-  ggplot(aes(max_temp, mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~
-  pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
+level_order <- c("Aug","Sep", "Oct","Nov")
+
+pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & (month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
+~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
 

### 2022-07-07 scripts/01-data_cleaning.R
Update 01-data_cleaning.R

@@ -908,12 +908,6 @@ pollution <- clean_names(pollution)
 # drop NA
 pollution <- pollution %>% drop_na(month_name)
 
-pollution$month_name <- factor(pollution$month_name, levels=c("Jan", "Feb","Mar", 
-                                                              "Apr", "May", "Jun", "Jul", 
-                                                              "Aug","Sep", "Oct", 
-                                                              "Nov",  "Dec"))
-
-
 # Save pollution data 2002-2020
 write_csv(
   x = pollution,

### 2022-07-07 scripts/01-data_cleaning.R
Update 01-data_cleaning.R

@@ -908,6 +908,12 @@ pollution <- clean_names(pollution)
 # drop NA
 pollution <- pollution %>% drop_na(month_name)
 
+# Change the units to be the same
+pollution1 <- pollution %>% mutate(mean_pollution = case_when(
+  pollutant == "CO" ~ mean_pollution*1000, 
+  pollutant != "CO" ~ mean_pollution
+))
+
 # Save pollution data 2002-2020
 write_csv(
   x = pollution,

### 2022-07-07 scripts/01-data_cleaning.R
Update 01-data_cleaning.R

@@ -909,7 +909,7 @@ pollution <- clean_names(pollution)
 pollution <- pollution %>% drop_na(month_name)
 
 # Change the units to be the same
-pollution1 <- pollution %>% mutate(mean_pollution = case_when(
+pollution <- pollution %>% mutate(mean_pollution = case_when(
   pollutant == "CO" ~ mean_pollution*1000, 
   pollutant != "CO" ~ mean_pollution
 ))

### 2022-07-07 outputs/paper/paper.qmd
Updates to paper

--- 2022-07-07 outputs/paper/paper.qmd
@@
 
-#### Visualizing Pollution by Year
+### Visualizing Pollution by Year
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 by Year", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 and CO by Year", fig.width=9, fig.height=3.2}
 
-pollution %>% filter(pollutant != "CO") %>% ggplot(
+pollution %>% ggplot(
   aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
@@
 The highest levels of pollution are as follows: 150.3 ppb for NO in 2002, 65 ppb for NO2 in 2005, 65.95833 ppb for O3 in 2003, and 22.79167 ppb for SO2 in 2002.
+There is also a decreasing pattern for the mean pollution from year 2002 to 2020 for CO by year.
+The highest level of pollution for CO is 1.7625 ppm in 2002.
+We can notice that there is a drop in the pollution levels from 2003 to 2005 but it rises again in 2006.
+Looking back at figure 3, this was the same year some of the highest temperatures were recorded from August to November.
+
+#### Visualizing NO (Nitric Oxide) by Year
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO by Year", fig.width=9, fig.height=3.2}
+
+pollution %>% filter(pollutant == "NO") %>% ggplot(
+  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)")
+```
+
+#### Visualizing NO2 (Nitrogen Dioxide) by Year
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO2 by Year", fig.width=9, fig.height=3.2}
+
+pollution %>% filter(pollutant == "NO2") %>% ggplot(
+  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")
+
+```
+
+#### Visualizing O3 (Ozone) by Year
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for O3 by Year", fig.width=9, fig.height=3.2}
+
+pollution %>% filter(pollutant == "O3") %>% ggplot(
+  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")
+```
+
+#### Visualizing CO (Carbon Monoxide) by Year
 
@@
   labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")
-
 ```
 
-Figure 5 shows a decreasing pattern for the mean pollution from year 2002 to 2020 for CO by year.
-The highest level of pollution for CO is 1.7625 ppm in 2002.
-We can notice that there is a drop in the pollution levels from 2003 to 2005 but it rises again in 2006.
-Looking back at figure 3, this was the same year some of the highest temperatures were recorded from August to November.
+#### Visualizing SO2 (Sulfur Dioxide) by Year
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for SO2 by Year", fig.width=9, fig.height=3.2}
+
+pollution %>% filter(pollutant == "SO2") %>% ggplot(
+  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")
+```
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 by Month", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 and CO by Month", fig.width=9, fig.height=3.2}
 
@@
 
-pollution %>% filter(pollutant != "CO") %>% ggplot(
+pollution %>% ggplot(
   aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
@@
 
-Figure 6 shows an ... pattern for the mean pollution from year 2002 to 2020 for NO, NO2, O3, and SO2 by month.
+#### Visualizing NO (Nitric Oxide) by Month
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO by Month", fig.width=9, fig.height=3.2}
+
+level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
+
+pollution %>% filter(pollutant == "NO") %>% ggplot(
+  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
+
+```
+
+#### Visualizing NO2 (Nitrogen Dioxide) by Month
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO2 by Month", fig.width=9, fig.height=3.2}
+
+level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
+
+pollution %>% filter(pollutant == "NO2") %>% ggplot(
+  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
+
+```
+
+#### Visualizing O3 (Ozone) by Month
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for O3 by Month", fig.width=9, fig.height=3.2}
+
+level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
+
+pollution %>% filter(pollutant == "O3") %>% ggplot(
+  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
+
+```
+
+#### Visualizing CO (Carbon Monoxide) by Month
 
@@
 
-Figure 7 shows an ... pattern for the mean pollution from year 2002 to 2020 for CO by month.
+#### Visualizing SO2 (Sulfur Dioxide) by Month
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO by Month", fig.width=9, fig.height=3.2}
+
+level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
+
+pollution %>% filter(pollutant == "SO2") %>% ggplot(
+  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
+
+```
 
@@
 
-pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & 
-  (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
-  month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, 
-  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
-  labs(color="Month") + xlab("Maximum Temperature") + 
-  ylab("Mean Pollution")
-```
+level_order <- c("Dec", "Jan", "Feb","Mar")
 
-In the above plot...
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO and SO2 by Month, Maximum Temperature, and Pollutant for December to March", fig.width=9, fig.height=3.2}
-
-pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & 
-  (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
-  month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, 
-  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
-  labs(color="Month") + xlab("Maximum Temperature") + 
-  ylab("Mean Pollution")
-
+pollu_temp %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
+~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") +  ylab("Mean Pollution")
 ```
@@
 
-pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & 
-  (month_name == "Apr" | month_name == "May" | month_name == "Jun" | 
-  month_name == "Jul")) %>% ggplot(aes(max_temp, mean_pollution, 
-  color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
-  labs(color="Month") + xlab("Maximum Temperature") + 
-  ylab("Mean Pollution")
+level_order <- c("Apr", "May", "Jun", "Jul")
 
-```
-
-In the above plot...
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO and SO2 by Month, Maximum Temperature, and Pollutant for April to July", fig.width=9, fig.height=3.2}
-
-pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & (month_name == "Apr" |
-   month_name == "May" | month_name == "Jun" | month_name == "Jul")) %>% ggplot(aes(max_temp,
-   mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
-  labs(color="Month") + xlab("Maximum Temperature") + 
-  ylab("Mean Pollution")
+pollu_temp %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" |  month_name == "Jul")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
+~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
 
@@
 
-pollu_temp %>% filter(pollutant != "CO" & pollutant != "SO2" & (month_name == "Aug" | 
-    month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>% ggplot(aes(max_temp,
-    mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~ pollutant) + 
-    labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
+level_order <- c("Aug","Sep", "Oct","Nov")
 
-```
-
-In the above plot...
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO and SO2 by Month, Maximum Temperature, and Pollutant for August to November", fig.width=9, fig.height=3.2}
-
-pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") & 
-  (month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>%
-  ggplot(aes(max_temp, mean_pollution, color=month_name)) + geom_point() + facet_grid(. ~
-  pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
+pollu_temp %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
+~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
 

### 2022-07-07 outputs/paper/paper.qmd
Update paper.qmd

@@
 
-## Visualizing Pollution from 2002-2020
-
-### Visualizing Pollution by Year
+## Visualizing Pollution from 2002-202
 
@@
 
-#### Visualizing NO (Nitric Oxide) by Year
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 and CO by Month", fig.width=9, fig.height=3.2}
+
+level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
+
+pollution %>% ggplot(
+  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppb)")
+```
+
+#### Visualizing NO (Nitric Oxide)
 
@@
 
-#### Visualizing NO2 (Nitrogen Dioxide) by Year
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO by Month", fig.width=9, fig.height=3.2}
+
+level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
+
+pollution %>% filter(pollutant == "NO") %>% ggplot(
+  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
+```
+
+#### Visualizing NO2 (Nitrogen Dioxide)
 
@@
 
-#### Visualizing O3 (Ozone) by Year
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO2 by Month", fig.width=9, fig.height=3.2}
+
+level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
+
+pollution %>% filter(pollutant == "NO2") %>% ggplot(
+  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
+```
+
+#### Visualizing O3 (Ozone)
 
@@
 ```
-
-#### Visualizing CO (Carbon Monoxide) by Year
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Year", fig.width=9, fig.height=3.2}
-
-pollution %>% filter(pollutant == "CO") %>% ggplot(
-  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")
-```
-
-#### Visualizing SO2 (Sulfur Dioxide) by Year
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for SO2 by Year", fig.width=9, fig.height=3.2}
-
-pollution %>% filter(pollutant == "SO2") %>% ggplot(
-  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")
-```
-
-#### Visualizing Pollution by Month
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 and CO by Month", fig.width=9, fig.height=3.2}
-
-level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
-
-pollution %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppb)")
-
-```
-
-#### Visualizing NO (Nitric Oxide) by Month
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO by Month", fig.width=9, fig.height=3.2}
-
-level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
-
-pollution %>% filter(pollutant == "NO") %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
-
-```
-
-#### Visualizing NO2 (Nitrogen Dioxide) by Month
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO2 by Month", fig.width=9, fig.height=3.2}
-
-level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
-
-pollution %>% filter(pollutant == "NO2") %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
-
-```
-
-#### Visualizing O3 (Ozone) by Month
 
@@
 
-#### Visualizing CO (Carbon Monoxide) by Month
+#### Visualizing CO (Carbon Monoxide)
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Year", fig.width=9, fig.height=3.2}
+
+pollution %>% filter(pollutant == "CO") %>% ggplot(
+  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")
+```
 
@@
   labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
-
 ```
 
-#### Visualizing SO2 (Sulfur Dioxide) by Month
+#### Visualizing SO2 (Sulfur Dioxide)
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for SO2 by Year", fig.width=9, fig.height=3.2}
+
+pollution %>% filter(pollutant == "SO2") %>% ggplot(
+  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")
+```
 
@@
   labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
-
 ```

### 2022-07-07 outputs/paper/paper.qmd
Updates to paper

@@
 
+level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
+
 pollution %>% filter(pollutant == "NO") %>% ggplot(
-  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)")
+  aes(x = year, y = mean_pollution, factor(month_name, level = level_order))) + geom_point() + 
+  labs(color="Month") + xlab("Year") + ylab("Mean Pollution (ppb)")
 ```
@@
 
-#### Visualizing NO2 (Nitrogen Dioxide)
+#### Visualizing NO2 (Nitrogen Dioxide) and O3 (Ozone)
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO2 by Year", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO2 and O3 by Year", fig.width=9, fig.height=3.2}
 
-pollution %>% filter(pollutant == "NO2") %>% ggplot(
+pollution %>% filter(pollutant == "NO2" | pollutant == "O3") %>% ggplot(
   aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO2 by Month", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO2 and O3 by Month", fig.width=9, fig.height=3.2}
 
@@
 
-pollution %>% filter(pollutant == "NO2") %>% ggplot(
+pollution %>% filter(pollutant == "NO2" | pollutant == "O3") %>% ggplot(
   aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
   labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
-```
-
-#### Visualizing O3 (Ozone)
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for O3 by Year", fig.width=9, fig.height=3.2}
-
-pollution %>% filter(pollutant == "O3") %>% ggplot(
-  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for O3 by Month", fig.width=9, fig.height=3.2}
-
-level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
-
-pollution %>% filter(pollutant == "O3") %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
-
 ```

### 2022-07-07 scripts/01-data_cleaning.R

@@ -906,7 +906,7 @@ pollution <- pollution %>% select(city, pollutant, mean_pollution, day, month, y
 pollution <- clean_names(pollution)
 
 # drop NA
-pollution <- pollution %>% drop_na(month_name)
+# pollution <- pollution %>% drop_na(month_name)
 
 # Change the units to be the same
 pollution <- pollution %>% mutate(mean_pollution = case_when(

### 2022-07-11 outputs/paper/paper.qmd
Updates to paper

@@
 This cleaned data was then saved into the repository as a csv file.
+The units for CO were ppm while the rest were ppb.
+To keep the data consistent, I converted those values to ppb as well by multiplying by 1,000.
 
@@
 I'm interested in finding the trends of the mean pollution levels from 2002 to 2020 by year.
-The units of NO, NO2, O3, and SO2 are the same (ppb) whereas the units for CO are different (ppm).
-For this reason, these plots have been made with regards to those units.
 
@@
 There is also a decreasing pattern for the mean pollution from year 2002 to 2020 for CO by year.
-The highest level of pollution for CO is 1.7625 ppm in 2002.
+The highest level of pollution for CO is 1762.5 ppb in 2002.
 We can notice that there is a drop in the pollution levels from 2003 to 2005 but it rises again in 2006.
@@
 
+Figure 5 suggests that the highest levels of CO are during winter and the colder months.
+This might be due to the fact that heaters are used the most in this time and do emit carbon monoxide.
+This plot does not tell us much else about the other pollutants due so from here we will take a look at each pollutant and how it has behaved from 2002 to 2020 by year and month.
+
 #### Visualizing NO (Nitric Oxide)
@@
 pollution %>% filter(pollutant == "NO") %>% ggplot(
-  aes(x = year, y = mean_pollution, factor(month_name, level = level_order))) + geom_point() + 
-  labs(color="Month") + xlab("Year") + ylab("Mean Pollution (ppb)")
+  aes(x = year, y = mean_pollution, factor(month_name, level = level_order))) +
+  geom_point() + labs(color="Month") + xlab("Year") + ylab("Mean Pollution (ppb)")
 ```
@@
 pollution %>% filter(pollutant == "NO") %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
+  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point() +
+  xlab("Month") + ylab("Mean Pollution (ppb)")
 ```
+
+Figure 6 shows a decreasing trend from 2002 to 2020 of the level of nitric oxide in the air.
+The highest level of pollution was in 2002 where the level was 150.2917 ppb.
+Figure 7 shows that in the colder months the levels of NO in the air are much higher than in the warmer months.
+The highest level in 2020 was around 49 ppb, a drastic drop from 2002.
+The highest level that was mentioned previously was seen in the month of December.
+However, it looks like on average the level of pollution for winter months is around 80 ppb.
+There are only some points that show high levels for the winter months.
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO2 and O3 by Year", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO2 by Year", fig.width=9, fig.height=3.2}
 
-pollution %>% filter(pollutant == "NO2" | pollutant == "O3") %>% ggplot(
-  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")
-
+pollution %>% filter(pollutant == "NO2" ) %>% ggplot(
+  aes(x = year, y = mean_pollution)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)")
 ```
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO2 and O3 by Month", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO2 by Month", fig.width=9, fig.height=3.2}
 
@@
 
-pollution %>% filter(pollutant == "NO2" | pollutant == "O3") %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
+pollution %>% filter(pollutant == "NO2") %>% ggplot(
+  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point() +
+  ylab("Mean Pollution (ppb)")
 ```
+
+Figure 8 shows the mean pollution for NO2 has been consistently decreasing from 2002 to 2020.
+The highest value was seen in 2005 where the level was 65 ppb.The highest level in 2020 was around 31 ppb.
+There are some years between 2002 and 2020 where the level increases again in 2011 and 2014.
+The mean levels of pollution by month in figure 9 appear to have higher levels in the winter months than in the summer months, but not by a lot.
+The average levels are around 40 ppb for most months.
+
+#### Visualizing O3 (Ozone)
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for O3 by Year", fig.width=9, fig.height=3.2}
+
+pollution %>% filter(pollutant == "O3") %>% ggplot(
+  aes(x = year, y = mean_pollution)) + geom_point() + 
+  xlab("Year") + ylab("Mean Pollution (ppb)")
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for O3 by Month", fig.width=9, fig.height=3.2}
+
+level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
+
+pollution %>% filter(pollutant == "O3") %>% ggplot(
+  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point() +
+  xlab("Month") + ylab("Mean Pollution (ppb)")
+```
+
+In figure 10, we can see that the mean pollution levels for ozone do not change much from 2002 to 2020.
+There is a slight decrease from the earlier years to the more recent years but not much.
+The highest level of CO was in 2003 with 65.95 ppb.
+In 2006, 2008, and 2012 the levels were also higher than 60 ppb.
+The average high levels are around 50 pb and that does not change from 2002 to 2020.
+This suggests that the level of ozone in our air has not been getting better and continues to affect our air and health.
+In figure 11, the hotter months for O3 have higher levels of pollution than the winter months.
+This suggests that ozone has a relationship with the high levels of temperature and that may be why the levels of pollution have not decreased since 2002, as our average global temperatures also keep rising.
+The highest value of 65.95 ppb was seen in June.
 
@@
 pollution %>% filter(pollutant == "CO") %>% ggplot(
-  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")
+  aes(x = year, y = mean_pollution)) + geom_point() + 
+  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)")
 ```
@@
 pollution %>% filter(pollutant == "CO") %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
+  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point() +
+  xlab("Month") + ylab("Mean Pollution (ppb)")
 ```
+
+In figure 12 and 13 we see a similar pattern as NO and NO2 where the level of pollution is decreasing throughout the years from 2002 to 2020 and the highest levels are in the winter months.
+The highest level of pollution for CO was seen in 2002 with 1762.5 ppb.
+This level is much higher than any of the other pollutants.
+This suggests that carbon monoxide levels are much more widespread than any of the other pollutants we have analyzed and due to this can have a affect on our health.
+Even with the decreasing levels, the affects will still be dangerous.
+The highest level in 2020 was around 550 ppb which is almost 5 times higher than the levels of NO, NO2, and O3 in 2020.
 
@@
 pollution %>% filter(pollutant == "SO2") %>% ggplot(
-  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppm)")
+  aes(x = year, y = mean_pollution)) + geom_point() + 
+  xlab("Year") + ylab("Mean Pollution (ppb)")
 ```
@@
 pollution %>% filter(pollutant == "SO2") %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppm)")
+  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point() +
+  xlab("Month") + ylab("Mean Pollution (ppb)")
 ```
+
+In figure 14, there is a decreasing pattern for the level of SO2 in the air from 2002 to 2020.
+The highest level in 2002 was 22.79 ppb and the highest level in 2020 seems to be less than 3 ppb.
+This is a drastic drop and suggests that the level of SO2 in the air is not affected by the increasing temperatures.
+In figure 15 we also see this.
+The levels of SO2 in the air remain consistent throughout the months and there is no difference between the winter and summer months.
 
@@
 
-In the above plot...
+In figure 16,
 
@@
 
-In the above plot...
-
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for August to November", fig.width=9, fig.height=3.2}
@@
 ```
-
-In the above plot...
 

### 2022-07-12 outputs/paper/paper.qmd
Updates to paper

@@
 
-#### Visualizing NO2 (Nitrogen Dioxide) and O3 (Ozone)
+#### Visualizing NO2 (Nitrogen Dioxide)
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for December to March", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 and SO2 by Month, Maximum Temperature, and Pollutant for December to March", fig.width=9, fig.height=3.2}
 
@@
 
-pollu_temp %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
+pollu_temp %>% filter((pollutant != "CO") & (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
 ~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") +  ylab("Mean Pollution")
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for April to July", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Month, Maximum Temperature, and Pollutant for December to March", fig.width=9, fig.height=3.2}
+
+level_order <- c("Dec", "Jan", "Feb","Mar")
+
+pollu_temp %>% filter((pollutant == "CO") & (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
+~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") +  ylab("Mean Pollution")
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 and SO2 by Month, Maximum Temperature, and Pollutant for April to July", fig.width=9, fig.height=3.2}
 
@@
 
-pollu_temp %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" |  month_name == "Jul")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
+pollu_temp %>% filter((pollutant != "CO") & (month_name == "Apr" | month_name == "May" | month_name == "Jun" |  month_name == "Jul")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
 ~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 by Month, Maximum Temperature, and Pollutant for August to November", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Month, Maximum Temperature, and Pollutant for April to July", fig.width=9, fig.height=3.2}
+
+level_order <- c("Apr", "May", "Jun", "Jul")
+
+pollu_temp %>% filter((pollutant == "CO") & (month_name == "Apr" | month_name == "May" | month_name == "Jun" |  month_name == "Jul")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
+~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 and SO2 by Month, Maximum Temperature, and Pollutant for August to November", fig.width=9, fig.height=3.2}
 
@@
 
-pollu_temp %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
+pollu_temp %>% filter((pollutant != "CO") & (month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
+~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Month, Maximum Temperature, and Pollutant for August to November", fig.width=9, fig.height=3.2}
+
+level_order <- c("Aug","Sep", "Oct","Nov")
+
+pollu_temp %>% filter((pollutant == "CO") & (month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
 ~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
@@
 
-### Temperature
+*Temperature*
 
-### Pollution
+*Pollution*
 
-### Health
+Air pollution continues to affect Canadians.
+Smog in the air, which is a mixture of smoke and fog, contains O3, SO2, and NO.
+Air pollutants like these are influenced by many factors including air temperature.
+Canada's population and economy continues to grow thus increasing the demand for production and delivery.
+This means more energy will be used to meet these demands and most of this energy will come from fossil fuels which will impact the quality of the air.
 
-### Temperature and Pollution
+From the analysis of NO, NO2, CO and SO2 we saw that there was a decline in the mean pollution levels of those pollutants.
+This is due to implementation of regulations and technological improvements for transportation vehicles and industrial processes.
+Canadians using more environmentally sustainable practices like public transport and carpooling have contributed to the decrease of lower pollution levels.
 
-### Temperature and Health
+*Health*
 
-### Pollution and Health
+*Temperature and Pollution*
 
-### Temperature, Pollution and Health
+*Temperature and Health*
+
+*Pollution and Health*
+
+Exposure to air pollution is known as a major cause to illnesses and death.
+
+Exposure to NO (Nitric Oxide) can irritate the lungs, reduce lung function and increase vulnerability to allergens for those with asthma.
+
+O3 (Ozone) is the main component of smog and has been associated with eye, nose, and throat irritations.
+Ozone also causes shortness of breath, aggravation of respiratory conditions and allergies, chronic obstructive pulmonary disease and asthma.
+Increased risk of cardiovascular disease is a result of exposure to O3 as well and all of these symptoms affect the young and the elderly that live near cities and/or with acute illnesses.
+
+*Temperature, Pollution and Health*
 

### 2022-07-12 outputs/paper/paper.qmd
Updates to paper

@@
     linesep = "", digits = 2, booktabs=TRUE) %>%
-  kable_styling(latex_options = "HOLD_position", font_size = 10)
+  kable_styling(latex_options = "HOLD_position", font_size = 10) + theme_bw()
 
@@
 
-temp_2002_2022 %>% filter(month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
+temp_2002_2022 %>% filter(month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw()
 
@@
 
-temp_2002_2022 %>% filter(month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
+temp_2002_2022 %>% filter(month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw()
 
@@
 
-temp_2002_2022 %>% filter(month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature")
+temp_2002_2022 %>% filter(month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw()
 
@@
   aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)")
+  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)") + theme_bw()
 ```
@@
   aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppb)")
+  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppb)") + theme_bw()
 ```
@@
   aes(x = year, y = mean_pollution, factor(month_name, level = level_order))) +
-  geom_point() + labs(color="Month") + xlab("Year") + ylab("Mean Pollution (ppb)")
+  geom_point() + labs(color="Month") + xlab("Year") + ylab("Mean Pollution (ppb)") + theme_bw()
 ```
@@
   aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point() +
-  xlab("Month") + ylab("Mean Pollution (ppb)")
+  xlab("Month") + ylab("Mean Pollution (ppb)") + theme_bw()
 ```
@@
   aes(x = year, y = mean_pollution)) + geom_point() + 
-  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)")
+  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)") + theme_bw()
 ```
@@
   aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point() +
-  ylab("Mean Pollution (ppb)")
+  ylab("Mean Pollution (ppb)") + theme_bw()
 ```
@@
   aes(x = year, y = mean_pollution)) + geom_point() + 
-  xlab("Year") + ylab("Mean Pollution (ppb)")
+  xlab("Year") + ylab("Mean Pollution (ppb)") + theme_bw()
 ```
@@
   aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point() +
-  xlab("Month") + ylab("Mean Pollution (ppb)")
+  xlab("Month") + ylab("Mean Pollution (ppb)") + theme_bw()
 ```
 
+Carbon monoxide is a product of incomplete combustion of hydrocarbon-based fuels.
 In figure 12 and 13 we see a similar pattern as NO and NO2 where the level of pollution is decreasing throughout the years from 2002 to 2020 and the highest levels are in the winter months.
@@
 
-#### Visualizing Mean Pollution by Temperature, Pollutant and Month
+#### Visualizing Mean Pollution by Temperature and Pollutant
 
@@
 
-level_order <- c("Dec", "Jan", "Feb","Mar")
-
-pollu_temp %>% filter((pollutant != "CO") & (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
-~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") +  ylab("Mean Pollution")
+pollu_temp %>% filter((pollutant != "CO" & pollutant != "SO2")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point() + facet_grid(.
+~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + theme_bw()
 ```
@@
 
-level_order <- c("Dec", "Jan", "Feb","Mar")
-
-pollu_temp %>% filter((pollutant == "CO") & (month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
-~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") +  ylab("Mean Pollution")
+pollu_temp %>% filter((pollutant == "CO")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point() + facet_grid(. ~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + theme_bw()
 ```
@@
 
-level_order <- c("Apr", "May", "Jun", "Jul")
-
-pollu_temp %>% filter((pollutant != "CO") & (month_name == "Apr" | month_name == "May" | month_name == "Jun" |  month_name == "Jul")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
-~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
-
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Month, Maximum Temperature, and Pollutant for April to July", fig.width=9, fig.height=3.2}
-
-level_order <- c("Apr", "May", "Jun", "Jul")
-
-pollu_temp %>% filter((pollutant == "CO") & (month_name == "Apr" | month_name == "May" | month_name == "Jun" |  month_name == "Jul")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
-~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
-
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 and SO2 by Month, Maximum Temperature, and Pollutant for August to November", fig.width=9, fig.height=3.2}
-
-level_order <- c("Aug","Sep", "Oct","Nov")
-
-pollu_temp %>% filter((pollutant != "CO") & (month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
-~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
-
-```
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Month, Maximum Temperature, and Pollutant for August to November", fig.width=9, fig.height=3.2}
-
-level_order <- c("Aug","Sep", "Oct","Nov")
-
-pollu_temp %>% filter((pollutant == "CO") & (month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov")) %>% ggplot(aes(max_temp, mean_pollution, color=factor(month_name, level=level_order))) + geom_point() + facet_grid(.
-~ pollutant) + labs(color="Month") + xlab("Maximum Temperature") + ylab("Mean Pollution")
+pollu_temp %>% filter((pollutant == "SO2")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point() + facet_grid(. ~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + theme_bw()
 
@@
 
+Carbon monoxide can significantly impact human health by entering the blood stream through the lungs and therefore inhibiting the blood's capacity to carry oxygen to organs and tissues.
+CO is harmful to those with heart disease and those with respiratory conditions.
+It also affects healthy individuals by reducing their exercise capacity, visual perception, manual dexterity, learning functions, and ability to perform complex tasks.
+
 *Temperature, Pollution and Health*

### 2022-07-13 outputs/paper/paper.qmd
Updating paper and analysis

@@
     linesep = "", digits = 2, booktabs=TRUE) %>%
-  kable_styling(latex_options = "HOLD_position", font_size = 10) + theme_bw()
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
 
@@
   aes(x = year, y = mean_pollution, factor(month_name, level = level_order))) +
-  geom_point() + labs(color="Month") + xlab("Year") + ylab("Mean Pollution (ppb)") + theme_bw()
+  geom_point(color = "gray") + labs(color="Month") + xlab("Year") + ylab("Mean Pollution (ppb)") + theme_bw()
 ```
@@
 pollution %>% filter(pollutant == "NO") %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point() +
+  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point(color = "gray") +
   xlab("Month") + ylab("Mean Pollution (ppb)") + theme_bw()
@@
 pollution %>% filter(pollutant == "NO2" ) %>% ggplot(
-  aes(x = year, y = mean_pollution)) + geom_point() + 
+  aes(x = year, y = mean_pollution)) + geom_point(color = "light blue") + 
   labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)") + theme_bw()
@@
 pollution %>% filter(pollutant == "NO2") %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point() +
+  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point(color = "light blue") +
   ylab("Mean Pollution (ppb)") + theme_bw()
@@
 pollution %>% filter(pollutant == "O3") %>% ggplot(
-  aes(x = year, y = mean_pollution)) + geom_point() + 
+  aes(x = year, y = mean_pollution)) + geom_point(color = "light pink") + 
   xlab("Year") + ylab("Mean Pollution (ppb)") + theme_bw()
@@
 pollution %>% filter(pollutant == "O3") %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point() +
+  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point(color = "light pink") +
   xlab("Month") + ylab("Mean Pollution (ppb)") + theme_bw()
@@
 pollution %>% filter(pollutant == "CO") %>% ggplot(
-  aes(x = year, y = mean_pollution)) + geom_point() + 
+  aes(x = year, y = mean_pollution)) + geom_point(color = "#9AD67F") + 
   labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)") + theme_bw()
@@
 pollution %>% filter(pollutant == "CO") %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point() +
+  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point(color = "#9AD67F") +
   xlab("Month") + ylab("Mean Pollution (ppb)") + theme_bw()
@@
 pollution %>% filter(pollutant == "SO2") %>% ggplot(
-  aes(x = year, y = mean_pollution)) + geom_point() + 
+  aes(x = year, y = mean_pollution)) + geom_point(color = "#847FD6") + 
   xlab("Year") + ylab("Mean Pollution (ppb)") + theme_bw()
@@
 pollution %>% filter(pollutant == "SO2") %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point() +
+  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point(color = "#847FD6") +
   xlab("Month") + ylab("Mean Pollution (ppb)") + theme_bw()
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 and SO2 by Month, Maximum Temperature, and Pollutant for December to March", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 and SO2 by Maximum Temperature and Pollutant for December to March", fig.width=9, fig.height=3.2}
 
-pollu_temp %>% filter((pollutant != "CO" & pollutant != "SO2")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point() + facet_grid(.
+pollu_temp %>% filter((pollutant != "CO" & pollutant != "SO2")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD") + facet_grid(.
 ~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + theme_bw()
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Month, Maximum Temperature, and Pollutant for December to March", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for CO by Maximum Temperature and Pollutant for December to March", fig.width=9, fig.height=3.2}
 
-pollu_temp %>% filter((pollutant == "CO")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point() + facet_grid(. ~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + theme_bw()
+pollu_temp %>% filter((pollutant == "CO")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD") + facet_grid(.
+~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + theme_bw()
 ```
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 and SO2 by Month, Maximum Temperature, and Pollutant for April to July", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3 and SO2 by Maximum Temperature and Pollutant for April to July", fig.width=9, fig.height=3.2}
 
-pollu_temp %>% filter((pollutant == "SO2")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point() + facet_grid(. ~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + theme_bw()
+pollu_temp %>% filter((pollutant == "SO2")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD") + facet_grid(.
+~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + theme_bw()
 
@@
 
-*Health*
+NO2 is usually formed through the burning of fossil fuels, transportation services, energy production and industrial processes.
+Nitrogen dioxide at higher concentration levels has a strong and harsh odor and can be seen over large cities as a brownish haze.
+NO2 can combine with H2O and form nitric acid and nitrous acid.
+This falls to earth through precipitation like rain, snow, and fog, and contributed to acidification and eutrophication of ecosystems.
+
+SO2 emissions come from natural sources like volcanoes and forest fires but these emissions are small compared to those from industrial sources.
+Environment Canada accounted for 1 million tonnes of man-made SO2 emissions compared to 100 tonnes of natural SO2 emissions in 2022.
+SO2 is seen in acid rain and due to this early recognition policies were put in place that have reduced SO2 pollution in recent decades as seen in figure 14.
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=9, fig.height=3.2}
+
+r <- pollution %>% filter((pollutant == "NO")) %>% select(mean_pollution)
+rr <- range(r)
+# 0.0000 150.2917
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 13.60186
+
+r <- pollution %>% filter((pollutant == "NO2")) %>% select(mean_pollution)
+rr <- range(r)
+#  0 65
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 14.97535
+
+r <- pollution %>% filter((pollutant == "O3")) %>% select(mean_pollution)
+rr <- range(r)
+# 0.00000 65.95833
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 22.98821
+
+r <- pollution %>% filter((pollutant == "SO2")) %>% select(mean_pollution)
+rr <- range(r)
+# -0.02500 22.79167
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 1.233524
+
+r <- pollution %>% filter((pollutant == "CO")) %>% select(mean_pollution)
+rr <- range(r)
+# 0.0 1762.5
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 254.4278
+
+pollutants <- c("NO", "NO2", "O3", "SO2", "CO")
+means_pollutants <- c(13.60, 14.97, 22.99, 1.23, 254.43)
+
+df <- data.frame(pollutants, means_pollutants)
+
+df %>%
+  select(pollutants, means_pollutants) %>%
+  kable(
+    caption = "Mean Pollution Levels of Each Pollutant",
+    col.names=c("Pollutant", "Mean Pollution Level"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=9, fig.height=3.2}
+
+# NO
+
+r <- pollution %>% filter((pollutant == "NO") & (year == 2002 | year == 2003 |
+                                                 year == 2004 | year == 2005 | 
+                                                   year == 2006)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 15.04657
+
+r <- pollution %>% filter((pollutant == "NO") & (year == 2007 | year == 2008 |
+                                                   year == 2009 | year == 2010 | 
+                                                   year == 2011)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 8.573613
+
+r <- pollution %>% filter((pollutant == "NO") & (year == 2012 | year == 2013 |
+                                                   year == 2014 | year == 2015 | 
+                                                   year == 2016)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 5.852171
+
+r <- pollution %>% filter((pollutant == "NO") & (year == 2017 | year == 2018 |
+                                                   year == 2019 | year == 2020)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 3.250599
+
+# NO2
+
+r <- pollution %>% filter((pollutant == "NO2") & (year == 2002 | year == 2003 |
+                                                   year == 2004 | year == 2005 | 
+                                                   year == 2006)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 19.87938
+
+r <- pollution %>% filter((pollutant == "NO2") & (year == 2007 | year == 2008 |
+                                                   year == 2009 | year == 2010 | 
+                                                   year == 2011)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 15.56782
+
+r <- pollution %>% filter((pollutant == "NO2") & (year == 2012 | year == 2013 |
+                                                   year == 2014 | year == 2015 | 
+                                                   year == 2016)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 13.39803
+
+r <- pollution %>% filter((pollutant == "NO2") & (year == 2017 | year == 2018 |
+                                                   year == 2019 | year == 2020)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 10.07811
+
+# O3
+
+r <- pollution %>% filter((pollutant == "O3") & (year == 2002 | year == 2003 |
+                                                    year == 2004 | year == 2005 | 
+                                                    year == 2006)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 20.94733
+
+r <- pollution %>% filter((pollutant == "O3") & (year == 2007 | year == 2008 |
+                                                    year == 2009 | year == 2010 | 
+                                                    year == 2011)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 22.5484
+
+r <- pollution %>% filter((pollutant == "O3") & (year == 2012 | year == 2013 |
+                                                    year == 2014 | year == 2015 | 
+                                                    year == 2016)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 23.97225
+
+r <- pollution %>% filter((pollutant == "O3") & (year == 2017 | year == 2018 |
+                                                    year == 2019 | year == 2020)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 24.85809
+
+# SO2
+
+r <- pollution %>% filter((pollutant == "SO2") & (year == 2002 | year == 2003 |
+                                                   year == 2004 | year == 2005 | 
+                                                   year == 2006)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 2.476657
+
+r <- pollution %>% filter((pollutant == "SO2") & (year == 2007 | year == 2008 |
+                                                   year == 2009 | year == 2010 | 
+                                                   year == 2011)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 1.276447
+
+r <- pollution %>% filter((pollutant == "SO2") & (year == 2012 | year == 2013 |
+                                                   year == 2014 | year == 2015 | 
+                                                   year == 2016)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 0.6760719
+
+r <- pollution %>% filter((pollutant == "SO2") & (year == 2017 | year == 2018 |
+                                                   year == 2019 | year == 2020)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 0.3232774
+
+# CO
+
+r <- pollution %>% filter((pollutant == "CO") & (year == 2002 | year == 2003 |
+                                                    year == 2004 | year == 2005 | 
+                                                    year == 2006)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 314.0882
+
+r <- pollution %>% filter((pollutant == "CO") & (year == 2007 | year == 2008 |
+                                                    year == 2009 | year == 2010 | 
+                                                    year == 2011)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 223.6469
+
+r <- pollution %>% filter((pollutant == "CO") & (year == 2012 | year == 2013 |
+                                                    year == 2014 | year == 2015 | 
+                                                    year == 2016)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 244.8328
+
+r <- pollution %>% filter((pollutant == "CO") & (year == 2017 | year == 2018 |
+                                                    year == 2019 | year == 2020)) %>% select(mean_pollution)
+rm <- mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 230.332
+
+pollutants <- c("NO", "NO", "NO", "NO", 
+                "NO2", "NO2", "NO2", "NO2",
+                "O3", "O3", "O3", "O3", 
+                "SO2", "SO2", "SO2", "SO2", 
+                "CO", "CO", "CO", "CO")
+year <- c("2002 to 2006", "2007 to 2011", "2012 to 2016", "2017 to 2020", 
+          "2002 to 2006", "2007 to 2011", "2012 to 2016", "2017 to 2020",
+          "2002 to 2006", "2007 to 2011", "2012 to 2016", "2017 to 2020", 
+          "2002 to 2006", "2007 to 2011", "2012 to 2016", "2017 to 2020", 
+          "2002 to 2006", "2007 to 2011", "2012 to 2016", "2017 to 2020")
+means_pollutants <- c(15.05, 8.57, 5.85, 3.25,
+                      19.88, 15.57, 13.40, 10.1,
+                      20.95, 22.55, 23.97, 24.86, 
+                      2.48, 1.28, 0.68, 0.32, 
+                      314.1, 223.65, 244.83, 230.33)
+
+df <- data.frame(pollutants, means_pollutants, year)
+df %>%
+  select(pollutants, means_pollutants, year) %>%
+  kable(
+    caption = "Mean Pollution Levels of Each Pollutant by 5-Year Intervals",
+    col.names=c("Pollutant", "Mean Pollution Level", "5-Year Interval"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
+```
 
@@
 Exposure to NO (Nitric Oxide) can irritate the lungs, reduce lung function and increase vulnerability to allergens for those with asthma.
+Nitric Oxide is seen in smog and can react with other substances to harm human health.
 
@@
 It also affects healthy individuals by reducing their exercise capacity, visual perception, manual dexterity, learning functions, and ability to perform complex tasks.
+
+SO2 (Sulfur Dioxide) can harm human respiratory systems by triggering respiratory illness and worsening existing respiratory and cardiovascular diseases.
+Sulfur dioxide is seen in smog and can react with other substances like nitric oxide to harm human health.
+When SO2 transforms into sulfate particles and combines with other particles in the atmosphere it can contribute to the formation of particulate matter (PM2.5) which harms human health as well as the environment.
+Sensitive populations such as children, the elderly, those with asthma are more susceptible to the affects of SO2.
+
+NO2 (Nitrogen Dioxide) has consequences on human health including emergency hospitalizations for respiratory conditions for adults that live in lower income communities in Toronto.
+Children are more vulnerable to these effects of respiratory damage.
+Children that have been exposed to NO2 have an increased risk of asthma, wheezing, infections in the ear, nose and throat, influenza, serious colds, and reduced lung function.
 

### 2022-07-13 scripts/01-data_cleaning.R

@@ -908,6 +908,7 @@ pollution <- clean_names(pollution)
 # drop NA
 # pollution <- pollution %>% drop_na(month_name)
 
+
 # Change the units to be the same
 pollution <- pollution %>% mutate(mean_pollution = case_when(
   pollutant == "CO" ~ mean_pollution*1000, 
@@ -925,6 +926,7 @@ write_csv(
 pollu_temp <- dplyr::full_join(temp_2002_2022, pollution)
 
 pollu_temp <- pollu_temp %>% drop_na(pollutant)
+pollu_temp <- pollu_temp %>% drop_na(month_name)
 pollu_temp <- pollu_temp %>% select(city, pollutant, mean_pollution, max_temp, min_temp, mean_temp, 
                                     day, month, year, month_name)
 

### 2022-07-13 scripts/02-data_analysis.R

@@ -16,10 +16,10 @@ library(kableExtra)
 library(ggplot2)
 library(dplyr)
 
-# Load temperature data
+# Load temperature data ----
 temperature_2002_2022 <- read.csv("~/Desktop/SUDS2022/inputs/data/
                                   clean_temp_data/temperature_2002_2022.csv")
-# Load pollution data
+# Load pollution data ----
 poll_2011_2020 <- read.csv("~/Desktop/SUDS2022/inputs/data/
                            clean_pollution_data/pollution_2011_2020.csv")
 
@@ -79,7 +79,7 @@ range(r)
 
 ## Pollution ----
 
-# Visualizing Pollution by Year 
+# Visualizing Pollution by Year ----
 
 pollution %>% filter(pollutant != "CO") %>% ggplot(
   aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
@@ -106,26 +106,196 @@ pollution1 <- pollution %>%
                                                     "Nov",  "Dec")))
 
 
-# Lowest and highest points
+# Lowest and highest points and means ----
+
 r <- pollution %>% filter((pollutant == "NO")) %>% select(mean_pollution)
 range(r)
 #  0.0000 150.2917
+mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 13.60186
 
 r <- pollution %>% filter((pollutant == "NO2")) %>% select(mean_pollution)
 range(r)
 #  0 65
+mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 14.97535
 
 r <- pollution %>% filter((pollutant == "O3")) %>% select(mean_pollution)
 range(r)
 # 0.00000 65.95833
+mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 22.98821
 
 r <- pollution %>% filter((pollutant == "SO2")) %>% select(mean_pollution)
 range(r)
 # -0.02500 22.79167
+mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 1.233524
 
 r <- pollution %>% filter((pollutant == "CO")) %>% select(mean_pollution)
 range(r)
-0.0000 1.7625
+# 0.0 1762.5
+mean(as.numeric(r$mean_pollution),na.rm=TRUE)
+# 254.4278
+
+
+# Means of each pollutant in 5 year intervals ----
+
+# NO
+
+r <- pollution %>% filter((pollutant == "NO") & (year == 2002 | year == 2003 |
[…]

### 2022-07-13 scripts/plots.R

@@ -216,3 +216,19 @@ pollu_temp %>%
                 "Min Temp", "Mean Temp", "Day", "Month", "Year"),
     linesep = "", digits = 2, booktabs=TRUE) %>%
   kable_styling(latex_options = "HOLD_position", font_size = 10)
+
+
+
+
+
+pollutants <- c("NO", "NO2", "O3", "SO2", "CO")
+means_pollutants <- c(13.60, 14.97, 22.99, 1.23, 254.43) 
+
+df <- data.frame(pollutants, means_pollutants)
+
+df %>%
+  select(pollutants, means_pollutants) %>%
+  kable(
+    col.names=c("Pollutant", "Mean Pollution Level"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)

### 2022-07-18 scripts/plots.R
Update plots.R

@@ -216,19 +216,3 @@ pollu_temp %>%
                 "Min Temp", "Mean Temp", "Day", "Month", "Year"),
     linesep = "", digits = 2, booktabs=TRUE) %>%
   kable_styling(latex_options = "HOLD_position", font_size = 10)
-
-
-
-
-
-pollutants <- c("NO", "NO2", "O3", "SO2", "CO")
-means_pollutants <- c(13.60, 14.97, 22.99, 1.23, 254.43) 
-
-df <- data.frame(pollutants, means_pollutants)
-
-df %>%
-  select(pollutants, means_pollutants) %>%
-  kable(
-    col.names=c("Pollutant", "Mean Pollution Level"),
-    linesep = "", digits = 2, booktabs=TRUE) %>%
-  kable_styling(latex_options = "HOLD_position", font_size = 10)

### 2022-07-18 outputs/paper/paper.qmd
Updates to paper

@@
 toc: true
-# bibliography: references.bib
+bibliography: references.bib
 ---
@@
 
-As climate change worsens, the danger it presents to human health worsens as well.
+As climate change worsens, the danger it presents to human health worsens as well [@citenasaclimatewarn].
 The health impacts due to climate change are already being demonstrated in parts of Canada.
-Canada is warming at almost twice the global rate and if this continues, emissions will exceed 2 degrees Celsius worldwide.
+Canada is warming at almost twice the global rate and if this continues, emissions will exceed 2 degrees Celsius worldwide [@citehealthcanada].
 
-The Paris Agreement within the United Nations Framework Convention on Climate Change includes 175 countries that have agreed to limit the average global temperature rise to below 2 degress Celcius.
-Canada had signed to the Paris Agreement as well and committed to cutting emission 30% below the 2005 levels by 2030.
+The Paris Agreement within the United Nations Framework Convention on Climate Change includes 175 countries that have agreed to limit the average global temperature rise to below 2 degress Celcius [@citehealthcanada].
+Canada had signed to the Paris Agreement as well and committed to cutting emission 30% below the 2005 levels by 2030 [@citehealthcanada].
 However, Canada's emission increased in 2016 compared to 1990.
-As a response to this increase, the Government of Canada developed a national climate change plan called *The Pan-Canadian Framework on Clean Growth and Climate Change*.
+As a response to this increase, the Government of Canada developed a national climate change plan called *The Pan-Canadian Framework on Clean Growth and Climate Change* [@citehealthcanada].
 This framework notes that human activities are driving severe changes to the Earth's climate which poses substantial risks to human health.
@@
 
-The data for temperature was processed using R studio and R.
+The data for temperature was processed using R studio and R [@citeR].
 I started off by renaming the column names for each temperature data set from 2002 to 2022.
@@
 The data set initially contained 31 variables.
-From these I selected the city, date, year, month, day, maximum temperature, minimum temperature, and mean temperature using the dplyr package.
-I then cleaned the names of the data set using the janitor package.
+From these I selected the city, date, year, month, day, maximum temperature, minimum temperature, and mean temperature using the dplyr package [@citedplyr].
+I then cleaned the names of the data set using the janitor package [@citejanitor].
 I added a column for the names of the month by mutating the data set and changed the city rows to represent only Toronto data.
@@
 I replaced these 24 columns (for each hour of the day) with one column that contained the mean pollution level for that day and pollutant.
-I filtered for Toronto from the list of cities in the data set using the dplyr package and removed the duplicate rows where the pollutant and date repeated.
+I filtered for Toronto from the list of cities in the data set using the dplyr package [@citedplyr] and removed the duplicate rows where the pollutant and date repeated.
 The missing values in this data set were labeled as "-999", so I replaced these values with NA and then with 0 for consistency.
@@
 
-temp_2002_2022 %>% filter(month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw()
+temp_2002_2022 %>% filter(month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw()
 
@@
 
-temp_2002_2022 %>% filter(month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw()
+temp_2002_2022 %>% filter(month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw()
 
@@
 
-temp_2002_2022 %>% filter(month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point() + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw()
+temp_2002_2022 %>% filter(month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw()
 
@@
 pollution %>% ggplot(
-  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point() + 
+  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point(alpha = 0.4) + 
   labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)") + theme_bw()
@@
 pollution %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point() + 
+  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point(alpha = 0.4) + 
   labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppb)") + theme_bw()
@@
 
-pollu_temp %>% filter((pollutant != "CO" & pollutant != "SO2")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD") + facet_grid(.
+pollu_temp %>% filter((pollutant != "CO" & pollutant != "SO2")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD", alpha = 0.4) + facet_grid(.
 ~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + theme_bw()
@@
 
-pollu_temp %>% filter((pollutant == "CO")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD") + facet_grid(.
+pollu_temp %>% filter((pollutant == "CO")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD", alpha = 0.4) + facet_grid(.
 ~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + theme_bw()
@@
 
-pollu_temp %>% filter((pollutant == "SO2")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD") + facet_grid(.
+pollu_temp %>% filter((pollutant == "SO2")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD", alpha = 0.4) + facet_grid(.
 ~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + theme_bw()
@@
 
-NO2 is usually formed through the burning of fossil fuels, transportation services, energy production and industrial processes.
-Nitrogen dioxide at higher concentration levels has a strong and harsh odor and can be seen over large cities as a brownish haze.
+NO2 is usually formed through the burning of fossil fuels, transportation services, energy production and industrial processes [@citeairquality].
+Nitrogen dioxide at higher concentration levels has a strong and harsh odor and can be seen over large cities as a brownish haze [@citeairquality].
 NO2 can combine with H2O and form nitric acid and nitrous acid.
-This falls to earth through precipitation like rain, snow, and fog, and contributed to acidification and eutrophication of ecosystems.
+This falls to earth through precipitation like rain, snow, and fog, and contributed to acidification and eutrophication of ecosystems [@citeairquality].
 
-SO2 emissions come from natural sources like volcanoes and forest fires but these emissions are small compared to those from industrial sources.
+SO2 emissions come from natural sources like volcanoes and forest fires but these emissions are small compared to those from industrial sources [@citeairquality].
 Environment Canada accounted for 1 million tonnes of man-made SO2 emissions compared to 100 tonnes of natural SO2 emissions in 2022.
@@
 
-NO2 (Nitrogen Dioxide) has consequences on human health including emergency hospitalizations for respiratory conditions for adults that live in lower income communities in Toronto.
-Children are more vulnerable to these effects of respiratory damage.
-Children that have been exposed to NO2 have an increased risk of asthma, wheezing, infections in the ear, nose and throat, influenza, serious colds, and reduced lung function.
+NO2 (Nitrogen Dioxide) has consequences on human health including emergency hospitalizations for respiratory conditions for adults that live in lower income communities in Toronto \[\@citestatscanNO2\].
+Children are more vulnerable to these effects of respiratory damage \[\@citestatscanNO2\].
+Children that have been exposed to NO2 have an increased risk of asthma, wheezing, infections in the ear, nose and throat, influenza, serious colds, and reduced lung function \[\@citestatscanNO2\].
 

### 2022-07-18 scripts/02-data_analysis.R

@@ -346,6 +346,15 @@ pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") &
 
 ### Models ----
 
+model1 <- lm(max_temp ~ mean_pollution , pollu_temp)
+
+summary(model1)
+anova(model1)
+
+model2 <- lm(mean_pollution ~ max_temp , pollu_temp)
+
+summary(model2)
+anova(model2)
 
 
 ### Anova Tests ----

### 2022-07-18 outputs/paper/paper.qmd
Updates to paper

@@
 
-Exposure to NO (Nitric Oxide) can irritate the lungs, reduce lung function and increase vulnerability to allergens for those with asthma.
+Exposure to NO (Nitric Oxide) can irritate the lungs, reduce lung function and increase vulnerability to allergens for those with asthma [@citeairquality].
 Nitric Oxide is seen in smog and can react with other substances to harm human health.
 
-O3 (Ozone) is the main component of smog and has been associated with eye, nose, and throat irritations.
-Ozone also causes shortness of breath, aggravation of respiratory conditions and allergies, chronic obstructive pulmonary disease and asthma.
-Increased risk of cardiovascular disease is a result of exposure to O3 as well and all of these symptoms affect the young and the elderly that live near cities and/or with acute illnesses.
+O3 (Ozone) is the main component of smog and has been associated with eye, nose, and throat irritations [@citeairquality].
+Ozone also causes shortness of breath, aggravation of respiratory conditions and allergies, chronic obstructive pulmonary disease and asthma [@citeairquality].
+Increased risk of cardiovascular disease is a result of exposure to O3 as well and all of these symptoms affect the young and the elderly that live near cities and/or with acute illnesses [@citeairquality].
 
-Carbon monoxide can significantly impact human health by entering the blood stream through the lungs and therefore inhibiting the blood's capacity to carry oxygen to organs and tissues.
+Carbon monoxide can significantly impact human health by entering the blood stream through the lungs and therefore inhibiting the blood's capacity to carry oxygen to organs and tissues [@citegovCanAirPoll].
 CO is harmful to those with heart disease and those with respiratory conditions.
-It also affects healthy individuals by reducing their exercise capacity, visual perception, manual dexterity, learning functions, and ability to perform complex tasks.
+It also affects healthy individuals by reducing their exercise capacity, visual perception, manual dexterity, learning functions, and ability to perform complex tasks [@citegovCanAirPoll].
 
-SO2 (Sulfur Dioxide) can harm human respiratory systems by triggering respiratory illness and worsening existing respiratory and cardiovascular diseases.
-Sulfur dioxide is seen in smog and can react with other substances like nitric oxide to harm human health.
-When SO2 transforms into sulfate particles and combines with other particles in the atmosphere it can contribute to the formation of particulate matter (PM2.5) which harms human health as well as the environment.
-Sensitive populations such as children, the elderly, those with asthma are more susceptible to the affects of SO2.
+SO2 (Sulfur Dioxide) can harm human respiratory systems by triggering respiratory illness and worsening existing respiratory and cardiovascular diseases [@citepembinainstitute].
+Sulfur dioxide is seen in smog and can react with other substances like nitric oxide to harm human health [@citeconferenceboard].
+When SO2 transforms into sulfate particles and combines with other particles in the atmosphere it can contribute to the formation of particulate matter (PM2.5) which harms human health as well as the environment [@citeconferenceboard].
+Sensitive populations such as children, the elderly, those with asthma are more susceptible to the affects of SO2 [@citepembinainstitute].
 
-NO2 (Nitrogen Dioxide) has consequences on human health including emergency hospitalizations for respiratory conditions for adults that live in lower income communities in Toronto \[\@citestatscanNO2\].
-Children are more vulnerable to these effects of respiratory damage \[\@citestatscanNO2\].
-Children that have been exposed to NO2 have an increased risk of asthma, wheezing, infections in the ear, nose and throat, influenza, serious colds, and reduced lung function \[\@citestatscanNO2\].
+NO2 (Nitrogen Dioxide) has consequences on human health including emergency hospitalizations for respiratory conditions for adults that live in lower income communities in Toronto \[\@citestatscan\].
+Children are more vulnerable to these effects of respiratory damage \[\@citestatscan\].
+Children that have been exposed to NO2 have an increased risk of asthma, wheezing, infections in the ear, nose and throat, influenza, serious colds, and reduced lung function \[\@citestatscan\].
 
@@
 This will not only ease the job of the data scientist but will also improve the reproducibility of their workflow and finding.
+
+\newpage
+
+## References

### 2022-07-18 scripts/02-data_analysis.R

@@ -347,15 +347,20 @@ pollu_temp %>% filter((pollutant == "CO" | pollutant == "SO2") &
 ### Models ----
 
 model1 <- lm(max_temp ~ mean_pollution , pollu_temp)
-
 summary(model1)
 anova(model1)
 
 model2 <- lm(mean_pollution ~ max_temp , pollu_temp)
-
 summary(model2)
 anova(model2)
+plot(model2$residuals, pch = 10, col = "red")
 
+model3 <- lm(mean_pollution ~ max_temp + min_temp + mean_temp , pollu_temp)
+summary(model3)
+anova(model3)
+plot(model3$residuals, pch = 10, col = "red")
+
+plot(pollu_temp$max_temp, pollu_temp$mean_pollution)
 
 ### Anova Tests ----
 

### 2022-07-19 outputs/paper/paper.qmd
Updates to paper

@@
 Changes in measures can prevent health impacts and build health systems and facilities that are more resilient to climate change.
+
+I collected data from the Government of Canada's website for both temperature and pollution.
+Once the data was collected for the available years between 2002 and 2022, it was cleaned using multiple packages in R which I will describe in the next section.
+Then the analysis begun and I found the following: temperatures have continued to increase since 2002, winters are less cold and summers are hotter, pollution levels for NO, NO2, CO, and SO2 have gone down since 2002 due to laws and regulations put in place and these pollutants typically have higher pollution levels in winter months, O3 pollution levels have remained mostly the same since 2002 and these levels are higher in summer months, temperature does seem to have a normal relationship with NO, CO, and SO2 levels and a positive relationship with NO2 and O3 levels.
+The presence of these pollutants have affects on health including lung irritation, increase in allergens for those with asthma, shortness of breath, respiratory illnesses, heart disease, cardiovascular diseases, infections, and reduced lung function.
+
+This paper outlines the data, how it was collected, variables of interest and how the data was processed and cleaned.
+Next I will look at the trends in temperature from 2002 to 2022 followed by the trends in pollution levels for each pollutant from 2002 to 2020.
+From here, the relationship between pollution levels and temperature will be analyzed followed by limitations and results and lastly next steps to take.
 
@@
 
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=9, fig.height=3.2}
+
+model2 <- lm(mean_pollution ~ max_temp , pollu_temp)
+summary(model2)
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=9, fig.height=3.2}
+
+model3 <- lm(mean_pollution ~ max_temp + min_temp + mean_temp , pollu_temp)
+summary(model3)
+```
+
 ## Limitations
@@
 *Temperature*
+
+Toronto will continue to experience warmer climate in the long-term due to climate change [@citefutureclimate].
+Toronto will have more extreme hot days where the temperature will be above 30 degrees celcius and less extreme cold days where the temperature will be below -20 degrees celcius [@citefutureclimate].
+By 2100, Toronto is expected to have warmed by 5 degrees celcius [@citefutureclimate].
+These warmer temperatures have significant impacts on the health and well-being of Toronto residents.
+If carbon emissions continue and no changes are put in place now, the maximum average annual temperatures are expected to be: 12-14 degrees celcius by 2040, 14-16 degrees celcius by 2070, and 16-18 degrees celcius by 2100 [@citefutureclimate].
+With more extreme heat days, Toronto will experience a greater demand for cooling which will increase the use of air conditioning causing a strain on electrical supply.
+These increases can be seen in figures 1, 2 and 3.
+In these we saw an increase in average temperatures for the Fall, Winter and Summer months.
 

### 2022-07-19 scripts/02-data_analysis.R

@@ -357,7 +357,7 @@ plot(model2$residuals, pch = 10, col = "red")
 
 model3 <- lm(mean_pollution ~ max_temp + min_temp + mean_temp , pollu_temp)
 summary(model3)
-anova(model3)
+anova(model3) 
 plot(model3$residuals, pch = 10, col = "red")
 
 plot(pollu_temp$max_temp, pollu_temp$mean_pollution)

### 2022-07-25 outputs/paper/paper.qmd
Updates to paper

@@
 abstract: "The growing impact of climate change in Canada demonstrates a clear and current threat to human health.
-These threats are becoming revealed in Canada as the country continues to face the consequences of warming almost at twice the global rate.
-The focus for this paper is on tempreture increases, lower air quality and more pollution and a decline in health for Toronto.
-Once the data for this research was collected, the data cleaning process begun which entailed of using packages in R like tidyverse, janitor, dplyr, ….
-Duplicate and irrelevant observations were removed, as well as unwanted outliers.
-The process was complete once the data was correct, consistent and usable.
-"
+These threats are prevalent in Canada as the country continues to face the consequences of warming almost at twice the global rate (Canada Warming Twice as Fast as the Rest of the World, BBC).
+The focus for this paper is on analyzing temperature increases, pollution levels and the health of effects of these in Toronto.
+Once the data for this research was collected, the data was analyzed using exploratory analysis, data visualization, and regression analysis.
+After analysis, trends and patterns show that temperatures have continued to increase since 2002, winters are less cold and summers are hotter, pollution levels for NO, NO2, CO, and SO2 have gone down since 2002 due to laws and regulations put in place and these pollutants typically have higher pollution levels in winter months, O3 pollution levels have remained mostly the same since 2002 and these levels are higher in summer months, temperature does seem to have a normal relationship with NO, CO, and SO2 levels and a positive relationship with NO2 and O3 levels.
+The presence of these pollutants has effects on health including lung irritation, increase in allergens for those with asthma, shortness of breath, respiratory illnesses, heart disease, cardiovascular diseases, infections, and reduced lung function."
 toc: true
@@
 
-I collected data from the Government of Canada's website for both temperature and pollution.
+Data was collected from the Government of Canada's website for both temperature and pollution.
 Once the data was collected for the available years between 2002 and 2022, it was cleaned using multiple packages in R which I will describe in the next section.
-Then the analysis begun and I found the following: temperatures have continued to increase since 2002, winters are less cold and summers are hotter, pollution levels for NO, NO2, CO, and SO2 have gone down since 2002 due to laws and regulations put in place and these pollutants typically have higher pollution levels in winter months, O3 pollution levels have remained mostly the same since 2002 and these levels are higher in summer months, temperature does seem to have a normal relationship with NO, CO, and SO2 levels and a positive relationship with NO2 and O3 levels.
+Then the analysis begun and the following was found: temperatures have continued to increase since 2002, winters are less cold and summers are hotter, pollution levels for NO, NO2, CO, and SO2 have gone down since 2002 due to laws and regulations put in place and these pollutants typically have higher pollution levels in winter months, O3 pollution levels have remained mostly the same since 2002 and these levels are higher in summer months, temperature does seem to have a normal relationship with NO, CO, and SO2 levels and a positive relationship with NO2 and O3 levels.
 The presence of these pollutants have affects on health including lung irritation, increase in allergens for those with asthma, shortness of breath, respiratory illnesses, heart disease, cardiovascular diseases, infections, and reduced lung function.
@@
 This paper outlines the data, how it was collected, variables of interest and how the data was processed and cleaned.
-Next I will look at the trends in temperature from 2002 to 2022 followed by the trends in pollution levels for each pollutant from 2002 to 2020.
+Following this, the trends in temperature from 2002 to 2022 followed by the trends in pollution levels for each pollutant from 2002 to 2020.
 From here, the relationship between pollution levels and temperature will be analyzed followed by limitations and results and lastly next steps to take.
@@
 This data was collected from January 2002 to May 2022.
-The original data contains daily data on maximum temperature, minimum temperature, mean temperature, heat degree days, cool degree days, total rain (mm), total snow (cm), total precipitation (mm), snow on ground (cm), direction of maximum gust, and speed of maximum gust (km/h).
-For this research, I will focus on only maximum temperature, minimum temperature, and mean temperature.
+The original data contains daily data on maximum temperature, minimum temperature, and mean temperature among others.
+This paper focuses on only maximum temperature, minimum temperature, and mean temperature.
 
@@
 Each year contains a file with continuous data and integrated data.
-I will be using continuous data for each year.
+Continuous data for each year will be used.
 The continuous data contains annual summaries and hourly data.
-I will be looking at the hourly data for CO (Carbon Monoxide), NO (Nitric Oxide), NO2 (Nitrogen Dioxide) , O3 (Ozone), and SO2 (Sulfur Dioxide).
+The analysis will use the hourly data for CO (Carbon Monoxide), NO (Nitric Oxide), NO2 (Nitrogen Dioxide) , O3 (Ozone), and SO2 (Sulfur Dioxide).
 This data includes the city (which is Toronto in our case), the levels of each pollutant for each hour of every day.
-I reduced the data to the mean of pollution levels by pollutant of each day.
+The data was reduced to the mean of pollution levels by pollutant of each day.
 
 #### Variables of Interest
-
-Temperature variables of interest:
-
--   City
-
--   Maximum Temperature
-
--   Minimum Temperature
-
--   Mean Temperature
-
--   Day
-
--   Month
-
--   Year
-
-Pollution variables of interest:
-
--   City
-
--   Pollutant
-
--   Mean pollution for pollutant each day
-
--   Day
-
--   Month
-
--   Year
-
-#### Data Processing
-
-*Temperature data*
-
-The data for temperature was processed using R studio and R [@citeR].
-I started off by renaming the column names for each temperature data set from 2002 to 2022.
-Once this was done I combined the 2002 to 2022 temperature data together.
-The data set initially contained 31 variables.
-From these I selected the city, date, year, month, day, maximum temperature, minimum temperature, and mean temperature using the dplyr package [@citedplyr].
-I then cleaned the names of the data set using the janitor package [@citejanitor].
-I added a column for the names of the month by mutating the data set and changed the city rows to represent only Toronto data.
-From here, I dropped any NA values from the maximum, minimum, and mean temperature columns.
-Lastly, I saved the cleaned data into the repository as a csv file.
-
-*Pollution Data*
-
-The pollution data was also processed using R studio and R.
-I first loaded the different data sets for each pollutant for each year.
-Each year had 5 data sets for each pollutant (NO, NO2, CO, O3, SO2).
-I started off by also renaming the column names for the data sets from 2002 to 2020.
-Once this was done I combined the pollutant data sets for each year into one data frame for the year.
-The data initially contained 27 columns.
-The data sets initially contained pollution levels for each hour of every day in the respective year.
-I replaced these 24 columns (for each hour of the day) with one column that contained the mean pollution level for that day and pollutant.
-I filtered for Toronto from the list of cities in the data set using the dplyr package [@citedplyr] and removed the duplicate rows where the pollutant and date repeated.
-The missing values in this data set were labeled as "-999", so I replaced these values with NA and then with 0 for consistency.
-This process was repeated for all years from 2002 to 2020.
-Once this was completed, the data was combined for all the years into one data frame and the following columns were selected: city, pollutant, mean pollution, day, month, year.
-This cleaned data was then saved into the repository as a csv file.
-The units for CO were ppm while the rest were ppb.
-To keep the data consistent, I converted those values to ppb as well by multiplying by 1,000.
-
-#### Sample
-
-The sample includes data from Toronto of daily temperature maximum, minimum and means from 2002 to 2022 and daily pollution levels for NO, NO2, CO, O3, SO2 from 2002 to 2020.
-
-#### The Data
 
@@
 library(dplyr)
+```
+
+Temperature variables of interest:
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="", fig.width=9, fig.height=3.2}
+
+temp_var <- c("City", "Maximum Temperature", "Minimum Temperature", "Mean Temperature", "Day", "Month", "Year")
+
+df <- data.frame(temp_var)
+
+df %>% select(temp_var) %>%
+  kable(
+    col.names=c("Variables"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
+
+```
+
+Pollution variables of interest:
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="", fig.width=9, fig.height=3.2}
+
+pollu_var <- c("City", "Pollutant", "Mean Pollution Level", "Day", "Month", "Year")
+
+df <- data.frame(pollu_var)
+
+df %>% select(pollu_var) %>%
+  kable(
+    col.names=c("Variables"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
+
+```
+
+#### Data Processing
+
+*Temperature data*
+
+The data for temperature was processed using R studio and R [@citeR].
+The data cleaning started with renaming the column names for each temperature data set from 2002 to 2022.
+Once this was done, the 2002 to 2022 temperature data was combined together.
+The data set initially contained 31 variables.
+From these the city, date, year, month, day, maximum temperature, minimum temperature, and mean temperature were selected using the dplyr package [@citedplyr].
+Then the names of the data set were cleaned using the janitor package [@citejanitor].
+A column was added for the names of the month by mutating the data set and changed the city rows to represent only Toronto data.
+From here, any NA values from the maximum, minimum, and mean temperature columns were dropped.
+Lastly, the cleaned data was saved into the repository as a csv file.
+
+*Pollution Data*
+
+The pollution data was also processed using R studio and R.
+The different data sets for each pollutant for each year were first loaded.
+Each year had 5 data sets for each pollutant (NO, NO2, CO, O3, SO2).
+The column names for the data sets from 2002 to 2020 were renamed.
+Once this was done, the pollutant data sets for each year were combined into one data frame for the year.
+The data initially contained 27 columns.
+The data sets initially contained pollution levels for each hour of every day in the respective year.
+These 24 columns (for each hour of the day) were replaced with one column that contained the mean pollution level for that day and pollutant.
+Toronto was filtered from the list of cities in the data set using the dplyr package [@citedplyr] and the duplicate rows were removed where the pollutant and date repeated.
+This process was repeated for all years from 2002 to 2020.
+Once this was completed, the data was combined for all the years into one data frame and the following columns were selected: city, pollutant, mean pollution, day, month, year.
+This cleaned data was then saved into the repository as a csv file.
+The units for CO were ppm while the rest were ppb.
+To keep the data consistent, those values were converted to ppb as well by multiplying by 1,000.
+
+#### Sample
+
+The sample includes data from Toronto of daily temperature maximum, minimum and means from 2002 to 2022 and daily pollution levels for NO, NO2, CO, O3, SO2 from 2002 to 2020.
+
+#### The Data
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2022 for December to March", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2012 for December to March", fig.width=9, fig.height=3.2}
 
@@
 
-temp_2002_2022 %>% filter(month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw()
+temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
+                            month_name == "Mar") & (year >= 2002 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, 
+                            color=factor(month_name, level = level_order))) + 
+  geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + 
+  theme_bw() + facet_grid(. ~ month_name)
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2022 for April to July", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2013 to 2022 for December to March", fig.width=9, fig.height=3.2}
+
+temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
+                             month_name == "Mar") & (year > 2012 & year <= 2022)) %>% ggplot(aes(x=year, y=max_temp, 
+                                                                                                  color=factor(month_name, level = level_order))) + 
+  geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + 
+  theme_bw() + facet_grid(. ~ month_name)
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2012 for April to July", fig.width=9, fig.height=3.2}
 
@@
 
-temp_2002_2022 %>% filter(month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw()
+temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") & (year >= 2002 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
+~ month_name)
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2022 for August to November", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2013 to 2022 for April to July", fig.width=9, fig.height=3.2}
+
+level_order <- c("Apr", "May", "Jun", "Jul")
+
+temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") & (year > 2012 & year <= 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
+~ month_name)
+
+```
+
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2012 for August to November", fig.width=9, fig.height=3.2}
 
@@
 
-temp_2002_2022 %>% filter(month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw()
+temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") & (year >= 2002 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
+~ month_name)
 
@@
 
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2013 to 2022 for August to November", fig.width=9, fig.height=3.2}
+
+level_order <- c("Aug","Sep", "Oct","Nov")
+
+temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") & (year > 2012 & year <= 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
+~ month_name)
+
+```
+
 ## Visualizing Pollution from 2002-202
@@
 I'm interested in finding the trends of the mean pollution levels from 2002 to 2020 by year.
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 and CO by Year", fig.width=9, fig.height=3.2}
-
-pollution %>% ggplot(
-  aes(x = year, y = mean_pollution, color = pollutant)) + geom_point(alpha = 0.4) + 
-  labs(color="Pollutant") + xlab("Year") + ylab("Mean Pollution (ppb)") + theme_bw()
-```
 
@@
 model2 <- lm(mean_pollution ~ max_temp , pollu_temp)
-summary(model2)
+# summary(model2)
+
+formula_1 <- c("formula = mean_pollution ~ max_temp", " ")
+coefficients_1 <- c("Intercept", "Max Temp")
+estimate <- c(82.0493, -0.2965)
+error <- c(1.6927, 0.0984)
+t_value <- c(48.473, -3.014)
+p_value <- c("< 2e-16", "0.00259")
+
+df <- data.frame(formula_1, coefficients_1, estimate, error, t_value, p_value)
+df %>% select(formula_1, coefficients_1, estimate, error, t_value, p_value) %>%
+  kable(
+    col.names=c("Formula", "Coefficients", "Estimate", "Error", "t-value", "p-value"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
 ```
@@
 model3 <- lm(mean_pollution ~ max_temp + min_temp + mean_temp , pollu_temp)
-summary(model3)
+# summary(model3)
+
+formula_1 <- c("formula = mean_pollution ~ max_temp + ", "min_temp + mean_temp", " ", " ")
+coefficients_1 <- c("Intercept", "Max Temp", "Min Temp", "Mean Temp")
+estimate <- c(78.654, -45.867, -46.606, 91.978)
+error <- c(2.903, 18.927, 19.010, 37.870)
+t_value <- c(27.090, -2.423, -2.452, 2.429)
+p_value <- c("< 2e-16", "0.0154", "0.0142", "0.0152")
+
+df <- data.frame(formula_1, coefficients_1, estimate, error, t_value, p_value)
+df %>% select(formula_1, coefficients_1, estimate, error, t_value, p_value) %>%
+  kable(
+    col.names=c("Formula", "Coefficients", "Estimate", "Error", "t-value", "p-value"),
+    linesep = "", digits = 2, booktabs=TRUE) %>%
+  kable_styling(latex_options = "HOLD_position", font_size = 10)
 ```

### 2022-07-25 scripts/plots.R

@@ -216,3 +216,5 @@ pollu_temp %>%
                 "Min Temp", "Mean Temp", "Day", "Month", "Year"),
     linesep = "", digits = 2, booktabs=TRUE) %>%
   kable_styling(latex_options = "HOLD_position", font_size = 10)
+
+

### 2022-07-25 outputs/paper/paper.qmd
Update paper.qmd

@@
 
-pollu_temp %>% filter((pollutant != "CO" & pollutant != "SO2")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD", alpha = 0.4) + facet_grid(.
+pollu_temp %>% filter((pollutant != "CO" & pollutant != "SO2") & (year > 2012 & year <= 2020)) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD", alpha = 0.4) + facet_grid(.
 ~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + theme_bw()
@@
 
-pollu_temp %>% filter((pollutant == "CO")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD", alpha = 0.4) + facet_grid(.
+pollu_temp %>% filter((pollutant == "CO")  & (year > 2012 & year <= 2020)) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD", alpha = 0.4) + facet_grid(.
 ~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + theme_bw()
@@
 
-pollu_temp %>% filter((pollutant == "SO2")) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD", alpha = 0.4) + facet_grid(.
+pollu_temp %>% filter((pollutant == "SO2")  & (year > 2012 & year <= 2020)) %>% ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD", alpha = 0.4) + facet_grid(.
 ~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + theme_bw()

### 2022-07-27 outputs/paper/paper.qmd
Updates to paper

--- 2022-07-25 outputs/paper/paper.qmd
@@
 These threats are prevalent in Canada as the country continues to face the consequences of warming almost at twice the global rate (Canada Warming Twice as Fast as the Rest of the World, BBC).
-The focus for this paper is on analyzing temperature increases, pollution levels and the health of effects of these in Toronto.
+The focus for this paper is on analyzing trends in temperature increases, pollution levels and the health effects of these in Toronto for the past 20 years.
+Impactful change can be made with this research in hopes of reversing some effects of climate change.
 Once the data for this research was collected, the data was analyzed using exploratory analysis, data visualization, and regression analysis.
-After analysis, trends and patterns show that temperatures have continued to increase since 2002, winters are less cold and summers are hotter, pollution levels for NO, NO2, CO, and SO2 have gone down since 2002 due to laws and regulations put in place and these pollutants typically have higher pollution levels in winter months, O3 pollution levels have remained mostly the same since 2002 and these levels are higher in summer months, temperature does seem to have a normal relationship with NO, CO, and SO2 levels and a positive relationship with NO2 and O3 levels.
-The presence of these pollutants has effects on health including lung irritation, increase in allergens for those with asthma, shortness of breath, respiratory illnesses, heart disease, cardiovascular diseases, infections, and reduced lung function."
+After analysis using these methods, trends and patterns show that temperatures have continued to increase since 2002, winters are less cold and summers are hotter, pollution levels for NO (Nitric Oxide), NO2 (Nitrogen Dioxide), CO (Carbon Monoxide), and SO2 (Sulfur Dioxide) have gone down since 2002 due to laws and regulations put in place and these pollutants typically have higher pollution levels in winter months which may correlate with the use of more electricity and heating during colder months.
+O3 (Ozone) pollution levels have remained mostly the same since 2002 and these levels are higher in summer months which may be due to the hotter temperatures in warmer months.
+Temperature seems to have a normal relationship with NO, CO, and SO2 levels and a positive relationship with NO2 and O3 levels.
+The presence of these pollutants has effects on health for the Toronto population including lung irritation, increase in allergens for those with asthma, shortness of breath, respiratory illnesses, heart disease, cardiovascular diseases, infections, and reduced lung function."
 toc: true
@@
 
-As climate change worsens, the danger it presents to human health worsens as well [@citenasaclimatewarn].
+As climate change worsens, the danger it presents to human health becomes more and more dangerous [@citenasaclimatewarn].
 The health impacts due to climate change are already being demonstrated in parts of Canada.
@@
 The Paris Agreement within the United Nations Framework Convention on Climate Change includes 175 countries that have agreed to limit the average global temperature rise to below 2 degress Celcius [@citehealthcanada].
-Canada had signed to the Paris Agreement as well and committed to cutting emission 30% below the 2005 levels by 2030 [@citehealthcanada].
-However, Canada's emission increased in 2016 compared to 1990.
-As a response to this increase, the Government of Canada developed a national climate change plan called *The Pan-Canadian Framework on Clean Growth and Climate Change* [@citehealthcanada].
+Canada had signed to the Paris Agreement as well and committed to cutting emissions 30% below the 2005 levels by 2030 [@citehealthcanada].
+However, Canada's emissions increased in 2016 compared to 1990.
+In response to this increase, the Government of Canada developed a national climate change plan called *The Pan-Canadian Framework on Clean Growth and Climate Change* [@citehealthcanada].
 This framework notes that human activities are driving severe changes to the Earth's climate which poses substantial risks to human health.
 
-Urgent action is needed and the science shows that it will make a tangible difference.
-The health impacts from climate change can be prevented by improving efforts both in societies and health systems in Canada.
-Changes in measures can prevent health impacts and build health systems and facilities that are more resilient to climate change.
+Urgent action is needed and the science shows that it will make a tangible difference [@].
+The health impacts from climate change can be mitigated by improving efforts both in societies and health systems in Canada [@].
+Changes in measures can prevent health impacts and build health systems and facilities that are more resilient to climate change [@].
 
 Data was collected from the Government of Canada's website for both temperature and pollution.
-Once the data was collected for the available years between 2002 and 2022, it was cleaned using multiple packages in R which I will describe in the next section.
-Then the analysis begun and the following was found: temperatures have continued to increase since 2002, winters are less cold and summers are hotter, pollution levels for NO, NO2, CO, and SO2 have gone down since 2002 due to laws and regulations put in place and these pollutants typically have higher pollution levels in winter months, O3 pollution levels have remained mostly the same since 2002 and these levels are higher in summer months, temperature does seem to have a normal relationship with NO, CO, and SO2 levels and a positive relationship with NO2 and O3 levels.
-The presence of these pollutants have affects on health including lung irritation, increase in allergens for those with asthma, shortness of breath, respiratory illnesses, heart disease, cardiovascular diseases, infections, and reduced lung function.
+Once the data was collected for the available years between 2002 and 2022, it was cleaned using multiple packages in R which will be described in the next section.
 
 This paper outlines the data, how it was collected, variables of interest and how the data was processed and cleaned.
-Following this, the trends in temperature from 2002 to 2022 followed by the trends in pollution levels for each pollutant from 2002 to 2020.
+Following this, the trends in temperature from 2002 to 2022 will be analyzed followed by the trends in pollution levels for each pollutant from 2002 to 2020.
 From here, the relationship between pollution levels and temperature will be analyzed followed by limitations and results and lastly next steps to take.
@@
 The data for temperature was processed using R studio and R [@citeR].
-The data cleaning started with renaming the column names for each temperature data set from 2002 to 2022.
-Once this was done, the 2002 to 2022 temperature data was combined together.
+The 2002 to 2022 temperature data was combined together.
 The data set initially contained 31 variables.
 From these the city, date, year, month, day, maximum temperature, minimum temperature, and mean temperature were selected using the dplyr package [@citedplyr].
-Then the names of the data set were cleaned using the janitor package [@citejanitor].
-A column was added for the names of the month by mutating the data set and changed the city rows to represent only Toronto data.
-From here, any NA values from the maximum, minimum, and mean temperature columns were dropped.
-Lastly, the cleaned data was saved into the repository as a csv file.
+Then any NA values from the maximum, minimum, and mean temperature columns were dropped.
 
@@
 Each year had 5 data sets for each pollutant (NO, NO2, CO, O3, SO2).
-The column names for the data sets from 2002 to 2020 were renamed.
-Once this was done, the pollutant data sets for each year were combined into one data frame for the year.
-The data initially contained 27 columns.
-The data sets initially contained pollution levels for each hour of every day in the respective year.
+The pollutant data sets for each year were combined into one data frame for the year.
+The data initially contained 27 columns including pollution levels for each hour of every day in the respective year.
 These 24 columns (for each hour of the day) were replaced with one column that contained the mean pollution level for that day and pollutant.
@@
 Once this was completed, the data was combined for all the years into one data frame and the following columns were selected: city, pollutant, mean pollution, day, month, year.
-This cleaned data was then saved into the repository as a csv file.
-The units for CO were ppm while the rest were ppb.
+The units for CO were ppm (part per million) while the rest were ppb (part per billion).
+These units are used to describe the concentration of gases in atmospheric chemistry [@citeppmppb].
 To keep the data consistent, those values were converted to ppb as well by multiplying by 1,000.
 
-#### Sample
-
-The sample includes data from Toronto of daily temperature maximum, minimum and means from 2002 to 2022 and daily pollution levels for NO, NO2, CO, O3, SO2 from 2002 to 2020.
+#### The Sample
 
@@
 
-Table 1 shows the rows of the Temperature data from 2002 to 2022.
-Variable "Year" indicates the year, variable "Month" indicates the month, and variable "Day" indicates the day.
-Variable "Max Temp" represents the maximum temperature on the day with the following month and year.
-Variable "Min Temp" represents the minimum temperature on the day with the following month and year.
-Variable "Mean Temp" represents the mean temperature on the day with the following month and year.
+Table 1 shows some of the rows of the Temperature data from 2002 to 2022.
 
@@
 
-Table 2 shows the rows of the Pollution data from 2002 to 2020.
-Variable "City" indicates the city from which the data was collected.
-Variable "Pollutant" indicates the pollutant and includes 5 types: CO, NO, NO2, O3, and SO2.
+Table 2 shows some of the rows of the Pollution data from 2002 to 2020.
 Variable "Mean Pollution" represents the mean pollution levels of the following pollutant on the following day of the month and year.
-Variable "Day" indicates the day, variable "Month" indicates the month, and variable "Year" indicates the year.
-Variable "Max Temp" represents the maximum temperature on the day with the following month and year.
 
@@
 
-I'm interested in finding the trends of the maximum temperatures from 2002 to 2022 by the month and year.
+This section is interested in finding the trends of the maximum temperatures from 2002 to 2022 by the month and year.
 
@@
 
-Figure 1 shows an increasing pattern for the maximum temperature from year 2002 to 2022 from December to March.
+Figure 1 shows an increasing pattern for the maximum temperature from year 2002 to 2012 from December to March.
 Temperatures in March continue to be increasingly higher than any of the other months in this plot.
 It is noticeable that the temperatures for all months are the highest in 2012.
-According to NASA, the average temperatures in 2012 was warmer than the mid-20th century baseline and is considered to be an outlying of seasonal extremes that are warmer.
-Even the lowest maximum temperature in January is higher than other years in 2012.
+According to NASA, the average temperatures in 2012 was warmer than the mid-20th century baseline and is considered to be an outlying of seasonal extremes that are warmer [@citenasaclimatewarn].
+The lowest maximum temperature in January is higher than other years in 2012.
 The plot shows that the average maximum temperature from December to March has been between -19.2 degrees Celsius in February of 2015 and 24.3 degrees Celsius in March of 2012.
@@
 
+Figure 2 shows an increasing pattern for the maximum temperature from year 2013 to 2022 from December to March.
+
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2012 for April to July", fig.width=9, fig.height=3.2}
@@
 
-Figure 2 shows an increasing pattern for the maximum temperature from year 2002 to 2022 from April to July.
+Figure 3 shows an increasing pattern for the maximum temperature from year 2002 to 2012 from April to July.
 Temperatures in July are consistently higher throughout the years than any of the other months in the plot and temperatures in April are consistently lower throughout the years than any of the other months.
@@
 
+Figure 4 shows an increasing pattern for the maximum temperature from year 2013 to 2022 from April to July.
+
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2012 for August to November", fig.width=9, fig.height=3.2}
@@
 
-Figure 3 shows an increasing pattern for the maximum temperature from year 2002 to 2022 from August to November.
+Figure 5 shows an increasing pattern for the maximum temperature from year 2002 to 2012 from August to November.
 Temperatures in both August and September are the highest throughout the years compared to November and October.
@@
 
+Figure 6 shows an increasing pattern for the maximum temperature from year 2013 to 2022 from August to November.
+
 ## Visualizing Pollution from 2002-202
 
-I'm interested in finding the trends of the mean pollution levels from 2002 to 2020 by year.
-
-Figure 4 shows a decreasing pattern for the mean pollution from year 2002 to 2020 for NO by year.
-We can notice that O3 and SO2 levels remain mostly the same throughout the years.
-The highest levels of pollution are as follows: 150.3 ppb for NO in 2002, 65 ppb for NO2 in 2005, 65.95833 ppb for O3 in 2003, and 22.79167 ppb for SO2 in 2002.
-There is also a decreasing pattern for the mean pollution from year 2002 to 2020 for CO by year.
-The highest level of pollution for CO is 1762.5 ppb in 2002.
-We can notice that there is a drop in the pollution levels from 2003 to 2005 but it rises again in 2006.
-Looking back at figure 3, this was the same year some of the highest temperatures were recorded from August to November.
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Mean Pollution 2002 to 2020 for NO, NO2, O3, SO2 and CO by Month", fig.width=9, fig.height=3.2}
-
-level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
-
-pollution %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution, color = pollutant)) + geom_point(alpha = 0.4) + 
-  labs(color="Pollutant") + xlab("Month") + ylab("Mean Pollution (ppb)") + theme_bw()
-```
-
-Figure 5 suggests that the highest levels of CO are during winter and the colder months.
-This might be due to the fact that heaters are used the most in this time and do emit carbon monoxide.
-This plot does not tell us much else about the other pollutants due so from here we will take a look at each pollutant and how it has behaved from 2002 to 2020 by year and month.
+This section is interested in finding the trends of the mean pollution levels from 2002 to 2020 by year.
 
@@
 
-Figure 6 shows a decreasing trend from 2002 to 2020 of the level of nitric oxide in the air.
-The highest level of pollution was in 2002 where the level was 150.2917 ppb.
-Figure 7 shows that in the colder months the levels of NO in the air are much higher than in the warmer months.
+Figure 7 shows a decreasing pattern for the mean pollution from year 2002 to 2020 for NO by year.
+The highest pollution level for NO was 150.3 ppb in 2002.
+Figure 8 shows a decreasing trend from 2002 to 2020 of the level of nitric oxide in the air.
+This figure shows that in the colder months the levels of NO in the air are much higher than in the warmer months.
 The highest level in 2020 was around 49 ppb, a drastic drop from 2002.
@@
 
-Figure 8 shows the mean pollution for NO2 has been consistently decreasing from 2002 to 2020.
+Figure 9 shows the mean pollution for NO2 has been consistently decreasing from 2002 to 2020.
 The highest value was seen in 2005 where the level was 65 ppb.The highest level in 2020 was around 31 ppb.
@@
 The highest level in 2020 was around 550 ppb which is almost 5 times higher than the levels of NO, NO2, and O3 in 2020.
+There is also a decreasing pattern for the mean pollution from year 2002 to 2020 for CO by year.
+We can notice that there is a drop in the pollution levels from 2003 to 2005 but it rises again in 2006.
+Looking back at figure 3, this was the same year some of the highest temperatures were recorded from August to November.
+The highest levels of CO are during winter and the colder months.
+This might be due to the fact that heaters are used the most in this time and do emit carbon monoxide.
+This plot does not tell us much else about the other pollutants due so from here we will take a look at each pollutant and how it has behaved from 2002 to 2020 by year and month.
 
@@
 
-*Temperature*
-
 Toronto will continue to experience warmer climate in the long-term due to climate change [@citefutureclimate].
@@
 In these we saw an increase in average temperatures for the Fall, Winter and Summer months.
-
-*Pollution*
 
@@
 ```
+
+Table 3 shows
 
@@
 
-*Temperature and Pollution*
-
-*Temperature and Health*
-
-*Pollution and Health*
+Table 4 shows ...
 
@@
 
-*Temperature, Pollution and Health*
+When looking at temperature, pollution and health together there are trends that have been analyzed in this paper that show that with temperature increases, there are higher levels of O3 (Ozone) and CO (Carbon Monoxide) levels have remained high.
+With the high presence of these pollutants in the air of Toronto, the population faces health risks that include reducing their exercise capacity, visual perception, manual dexterity, learning functions, ability to perform complex tasks, shortness of breath, aggravation of respiratory conditions and allergies, chronic obstructive pulmonary disease and asthma.
 

### 2022-07-27 scripts/plots.R

@@ -217,4 +217,66 @@ pollu_temp %>%
     linesep = "", digits = 2, booktabs=TRUE) %>%
   kable_styling(latex_options = "HOLD_position", font_size = 10)
 
+pollu_temp %>% filter((pollutant != "CO" & pollutant != "SO2") & (year > 2012 & year <= 2020)) %>% 
+  ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD", alpha = 0.4) + 
+  facet_grid(. ~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + 
+  theme_bw()
 
+pollu_temp %>% filter((pollutant != "CO" & pollutant != "SO2") & (year > 2002 & year <= 2012)) %>% 
+  ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD", alpha = 0.4) + 
+  facet_grid(. ~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + 
+  theme_bw()
+
+pollu_temp %>% filter((pollutant != "CO" & pollutant != "SO2")) %>% 
+  ggplot(aes(max_temp, mean_pollution)) + geom_point(color = "#6DD6CD", alpha = 0.4) + 
+  facet_grid(. ~ pollutant) + xlab("Maximum Temperature") +  ylab("Mean Pollution") + 
+  theme_bw()
+
+
+level_order <- c("Dec", "Jan", "Feb","Mar")
+
+temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
+                             month_name == "Mar") & (year >= 2002 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, 
+                                                                                                  color=factor(month_name, level = level_order))) + 
+  geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + 
+  theme_bw() + facet_grid(. ~ month_name, level = level_order)
+
+
+
+
+# sample
+
+pollution %>% filter(pollutant == "CO")
+# 6940
+pollution %>% filter(pollutant == "O3")
+# 6940
+pollution %>% filter(pollutant == "NO")
+# 6940
+pollution %>% filter(pollutant == "NO2")
+# 6940
+pollution %>% filter(pollutant == "SO2")
+# 6940
+
[…]

### 2022-07-27 outputs/paper/paper.qmd
Updates to paper

@@
 
-Carbon monoxide is a product of incomplete combustion of hydrocarbon-based fuels.
+Carbon monoxide is produced through burning of coal, gasoline, natural gas, oil, propane, and wood [@citeCOpoisoning].
 In figure 12 and 13 we see a similar pattern as NO and NO2 where the level of pollution is decreasing throughout the years from 2002 to 2020 and the highest levels are in the winter months.
@@
 This falls to earth through precipitation like rain, snow, and fog, and contributed to acidification and eutrophication of ecosystems [@citeairquality].
+NO2 (Nitrogen Dioxide) has consequences on human health including emergency hospitalizations for respiratory conditions for adults that live in lower income communities in Toronto \[\@citestatscan\].
+Children are more vulnerable to these effects of respiratory damage \[\@citestatscan\].
+Children that have been exposed to NO2 have an increased risk of asthma, wheezing, infections in the ear, nose and throat, influenza, serious colds, and reduced lung function \[\@citestatscan\].
 
@@
 SO2 is seen in acid rain and due to this early recognition policies were put in place that have reduced SO2 pollution in recent decades as seen in figure 14.
+SO2 (Sulfur Dioxide) can harm human respiratory systems by triggering respiratory illness and worsening existing respiratory and cardiovascular diseases [@citepembinainstitute].
+Sulfur dioxide is seen in smog and can react with other substances like nitric oxide to harm human health [@citeconferenceboard].
+When SO2 transforms into sulfate particles and combines with other particles in the atmosphere it can contribute to the formation of particulate matter (PM2.5) which harms human health as well as the environment [@citeconferenceboard].
+Sensitive populations such as children, the elderly, those with asthma are more susceptible to the affects of SO2 [@citepembinainstitute].
+
+Exposure to NO (Nitric Oxide) can irritate the lungs, reduce lung function and increase vulnerability to allergens for those with asthma [@citeairquality].
+Nitric Oxide is seen in smog and can react with other substances to harm human health.
+
+O3 (Ozone) is the main component of smog and has been associated with eye, nose, and throat irritations [@citeairquality].
+Ozone also causes shortness of breath, aggravation of respiratory conditions and allergies, chronic obstructive pulmonary disease and asthma [@citeairquality].
+Increased risk of cardiovascular disease is a result of exposure to O3 as well and all of these symptoms affect the young and the elderly that live near cities and/or with acute illnesses [@citeairquality].
+
+Carbon monoxide can significantly impact human health by entering the blood stream through the lungs and therefore inhibiting the blood's capacity to carry oxygen to organs and tissues [@citegovCanAirPoll].
+CO is harmful to those with heart disease and those with respiratory conditions.
+It also affects healthy individuals by reducing their exercise capacity, visual perception, manual dexterity, learning functions, and ability to perform complex tasks [@citegovCanAirPoll].
 
@@
 
-Table 3 shows
+Table 3 shows the mean pollution levels of each pollutant from 2002 to 2020.
+The highest mean is from CO levels followed by O3.
+From the previous analysis, carbon monoxide is produced through burning of coal, gasoline, natural gas, oil, propane, and wood and as these activities have been continuing throughout the city of Toronto thus CO levels remain high [@citeCOpoisoning].
+O3 levels have also been increasing from 2002 to 2020 while the remaining pollutants' levels (NO, NO2, and SO2) have been decreasing in that time.
 
@@
 
-Table 4 shows ...
-
-Exposure to air pollution is known as a major cause to illnesses and death.
-
-Exposure to NO (Nitric Oxide) can irritate the lungs, reduce lung function and increase vulnerability to allergens for those with asthma [@citeairquality].
-Nitric Oxide is seen in smog and can react with other substances to harm human health.
-
-O3 (Ozone) is the main component of smog and has been associated with eye, nose, and throat irritations [@citeairquality].
-Ozone also causes shortness of breath, aggravation of respiratory conditions and allergies, chronic obstructive pulmonary disease and asthma [@citeairquality].
-Increased risk of cardiovascular disease is a result of exposure to O3 as well and all of these symptoms affect the young and the elderly that live near cities and/or with acute illnesses [@citeairquality].
-
-Carbon monoxide can significantly impact human health by entering the blood stream through the lungs and therefore inhibiting the blood's capacity to carry oxygen to organs and tissues [@citegovCanAirPoll].
-CO is harmful to those with heart disease and those with respiratory conditions.
-It also affects healthy individuals by reducing their exercise capacity, visual perception, manual dexterity, learning functions, and ability to perform complex tasks [@citegovCanAirPoll].
-
-SO2 (Sulfur Dioxide) can harm human respiratory systems by triggering respiratory illness and worsening existing respiratory and cardiovascular diseases [@citepembinainstitute].
-Sulfur dioxide is seen in smog and can react with other substances like nitric oxide to harm human health [@citeconferenceboard].
-When SO2 transforms into sulfate particles and combines with other particles in the atmosphere it can contribute to the formation of particulate matter (PM2.5) which harms human health as well as the environment [@citeconferenceboard].
-Sensitive populations such as children, the elderly, those with asthma are more susceptible to the affects of SO2 [@citepembinainstitute].
-
-NO2 (Nitrogen Dioxide) has consequences on human health including emergency hospitalizations for respiratory conditions for adults that live in lower income communities in Toronto \[\@citestatscan\].
-Children are more vulnerable to these effects of respiratory damage \[\@citestatscan\].
-Children that have been exposed to NO2 have an increased risk of asthma, wheezing, infections in the ear, nose and throat, influenza, serious colds, and reduced lung function \[\@citestatscan\].
+Table 4 shows the mean pollution levels of each pollutant by five-year intervals.
+From this table, it is evident that NO, NO2, SO2, and CO levels have been decreasing since 2002 while O3 has been increasing.
+The levels of CO still remain higher than any of the other pollutants.
+The biggest difference in pollution levels are seen for NO.
+From 2002 to 2006 the mean was 15.05 but this dropped to 3.25 from 2017 to 2020.
 

### 2022-07-27 outputs/paper/paper.qmd
Updates to pollution data and paper

@@
   select(year, month, day, max_temp, min_temp, mean_temp) %>%
-  slice(c(162, 393, 745, 1318, 1819, 2115, 3222, 4222, 5222, 6333)) %>%
+  slice(c(162, 393, 4222, 5222, 6333)) %>%
   kable(
@@
   select(city, pollutant, mean_pollution, day, month, year) %>%
-  slice(c(100, 1000, 1200, 1500, 1800, 2000, 2500, 2800, 3600, 5000, 6000, 7000, 10000, 11000, 15000)) %>%
+  slice(c(100, 1000, 5000, 11000, 15000)) %>%
   kable(
@@
 There is also a decreasing pattern for the mean pollution from year 2002 to 2020 for CO by year.
-We can notice that there is a drop in the pollution levels from 2003 to 2005 but it rises again in 2006.
+It can be noticed that there is a drop in the pollution levels from 2003 to 2005 but it rises again in 2006.
 Looking back at figure 3, this was the same year some of the highest temperatures were recorded from August to November.
@@
 This is a drastic drop and suggests that the level of SO2 in the air is not affected by the increasing temperatures.
-In figure 15 we also see this.
+This is also seen in figure 15.
 The levels of SO2 in the air remain consistent throughout the months and there is no difference between the winter and summer months.
@@
 
-In this limitations section, I will focus on the data used and look into why it is not easy to take the implications mentioned above directly.
+This limitations section will focus on the data used and look into why it is not easy to take the implications mentioned above directly.
 
@@
 
-From the analysis of NO, NO2, CO and SO2 we saw that there was a decline in the mean pollution levels of those pollutants.
+From the analysis of NO, NO2, CO and SO2 it was seen that there was a decline in the mean pollution levels of those pollutants.
 This is due to implementation of regulations and technological improvements for transportation vehicles and industrial processes.
@@
 
-Following on from the previous section, we can make good suggestions on ways to improve the data so that the findings will be less ambiguous and more ethical to use with confidence.
-First off, we can start by collecting pollution data from 2021 to 2022.
+Following on from the previous section, good suggestions can be made on ways to improve the data so that the findings will be less ambiguous and more ethical to use with confidence.
+First off, by collecting pollution data from 2021 to 2022.
 It is possible that this data is available but not published yet on the website.

### 2022-07-27 scripts/01-data_cleaning.R

@@ -906,7 +906,7 @@ pollution <- pollution %>% select(city, pollutant, mean_pollution, day, month, y
 pollution <- clean_names(pollution)
 
 # drop NA
-# pollution <- pollution %>% drop_na(month_name)
+pollution <- pollution %>% drop_na(month_name)
 
 
 # Change the units to be the same

### 2022-07-27 scripts/plots.R

@@ -280,3 +280,34 @@ pollution %>% filter((pollutant == "SO2") & (year > 2012 & year <= 2020))
 
 pollution %>% filter((pollutant == "NO") & (year >= 2002 & year <= 2012) & (month == "Dec" | 
                          month == "Jan" | month == "Feb" | month == "Mar"))
+
+
+
+
+
+
+
+
+
+
+
+
+
+# pollution NA months
+
+df <- pollution
+df <- df %>% drop_na(month_name)
+
+level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
+
+df %>% filter(pollutant == "NO") %>% ggplot(
+  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point(color = "gray") +
+  xlab("Month") + ylab("Mean Pollution (ppb)") + theme_bw()
+
+
+
+
+
+
+
+

### 2022-08-04 outputs/paper/paper.qmd
Updates to paper and presentation

@@
 editor: visual
-abstract: "The growing impact of climate change in Canada demonstrates a clear and current threat to human health.
-These threats are prevalent in Canada as the country continues to face the consequences of warming almost at twice the global rate (Canada Warming Twice as Fast as the Rest of the World, BBC).
-The focus for this paper is on analyzing trends in temperature increases, pollution levels and the health effects of these in Toronto for the past 20 years.
-Impactful change can be made with this research in hopes of reversing some effects of climate change.
-Once the data for this research was collected, the data was analyzed using exploratory analysis, data visualization, and regression analysis.
-After analysis using these methods, trends and patterns show that temperatures have continued to increase since 2002, winters are less cold and summers are hotter, pollution levels for NO (Nitric Oxide), NO2 (Nitrogen Dioxide), CO (Carbon Monoxide), and SO2 (Sulfur Dioxide) have gone down since 2002 due to laws and regulations put in place and these pollutants typically have higher pollution levels in winter months which may correlate with the use of more electricity and heating during colder months.
-O3 (Ozone) pollution levels have remained mostly the same since 2002 and these levels are higher in summer months which may be due to the hotter temperatures in warmer months.
-Temperature seems to have a normal relationship with NO, CO, and SO2 levels and a positive relationship with NO2 and O3 levels.
-The presence of these pollutants has effects on health for the Toronto population including lung irritation, increase in allergens for those with asthma, shortness of breath, respiratory illnesses, heart disease, cardiovascular diseases, infections, and reduced lung function."
+abstract: "Climate change has a growing impact in Canada.
+This paper analyses the trends in temperature and pollution levels in Toronto over the last 20 years using exploratory analysis, data visualization, and regression analysis.
+Temperatures have continued to increase since 2002 and winters are experiencing less extreme cold while summers are experiencing more extreme heat.
+Further, pollution levels for NO (Nitric Oxide), NO2 (Nitrogen Dioxide), CO (Carbon Monoxide), and SO2 (Sulfur Dioxide) have decreased since 2002 which may be related to laws and regulations.
+Data visualizations showed that these pollutants appear to have higher levels in winter months while O3 (Ozone) pollution levels have remained mostly the same since 2002 and O3 levels have higher maximums in summer months.
+The presence of these pollutants has detrimental health effects for the Toronto population including lung irritation, increase in allergens for those with asthma, respiratory illnesses, heart disease, cardiovascular diseases, and reduced lung function.
+Understanding which pollutants are most problematic during particular months can enable further research which would reduce their effects on health."
 toc: true
@@
 
-Urgent action is needed and the science shows that it will make a tangible difference [@].
-The health impacts from climate change can be mitigated by improving efforts both in societies and health systems in Canada [@].
-Changes in measures can prevent health impacts and build health systems and facilities that are more resilient to climate change [@].
+Urgent action is needed and the science shows that it will make a tangible difference \[\@\].
+The health impacts from climate change can be mitigated by improving efforts both in societies and health systems in Canada \[\@\].
+Changes in measures can prevent health impacts and build health systems and facilities that are more resilient to climate change \[\@\].
 

### 2022-08-04 scripts/plots.R

@@ -309,5 +309,74 @@ df %>% filter(pollutant == "NO") %>% ggplot(
 
 
 
+level_order <- c("Dec", "Jan", "Feb","Mar")
+
+temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
+                             month_name == "Mar") & (year >= 2002 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, 
+                                                                                                  color=factor(month_name, level = level_order))) + 
+  geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + 
+  theme_bw() + facet_grid(. ~ month_name)
+
+data_new <- temp_2002_2022
+data_new$month_name <- factor(data_new$month_name,
+                         levels = c("Dec", "Jan", "Feb","Mar"))
+data_new %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
+                             month_name == "Mar") & (year >= 2002 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, 
+                                                                                                  color=factor(month_name, level = level_order))) + 
+  geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + 
+  theme_bw() + facet_grid(. ~ month_name)
+
+data_new %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
+                       month_name == "Mar") & (year >= 2007 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, 
+                                                                                            color=factor(month_name, level = level_order))) + 
+  geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + 
+  theme_bw() + facet_grid(. ~ month_name)
 
 
+
+
+
+group.colors <- c("#4253AF", "#8BDEF2", "#D9EDF5", "#5A69A9")
+data_new %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
+                       month_name == "Mar") & (year >= 2014 & year <= 2020)) %>% ggplot(aes(x=year, y=max_temp, 
[…]

### 2022-08-04 outputs/paper/paper.qmd
Updates to paper

@@
 
-Temperature variables of interest:
-
 ```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="", fig.width=9, fig.height=3.2}
@@
 
-df <- data.frame(temp_var)
+pollu_var <- c("City", "Pollutant", "Mean Pollution Level", "Day", "Month", "Year", "")
 
-df %>% select(temp_var) %>%
+df <- data.frame(temp_var, pollu_var)
+
+df %>% select(temp_var, pollu_var) %>%
   kable(
-    col.names=c("Variables"),
-    linesep = "", digits = 2, booktabs=TRUE) %>%
-  kable_styling(latex_options = "HOLD_position", font_size = 10)
-
-```
-
-Pollution variables of interest:
-
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="", fig.width=9, fig.height=3.2}
-
-pollu_var <- c("City", "Pollutant", "Mean Pollution Level", "Day", "Month", "Year")
-
-df <- data.frame(pollu_var)
-
-df %>% select(pollu_var) %>%
-  kable(
-    col.names=c("Variables"),
+    col.names=c("Temperature Variables of Interest", "Pollution Variables of Interest"),
     linesep = "", digits = 2, booktabs=TRUE) %>%
@@
 
-temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
-                            month_name == "Mar") & (year >= 2002 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, 
-                            color=factor(month_name, level = level_order))) + 
-  geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + 
-  theme_bw() + facet_grid(. ~ month_name)
+temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") & (year >= 2002 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
+~ month_name) + scale_x_discrete(breaks = 0:2100)
 
@@
   geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + 
-  theme_bw() + facet_grid(. ~ month_name)
+  theme_bw() + facet_grid(. ~ month_name) + scale_x_discrete(breaks = 0:2100)
 
@@
 temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") & (year >= 2002 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
-~ month_name)
+~ month_name) + scale_x_discrete(breaks = 0:2100)
 
@@
 temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") & (year > 2012 & year <= 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
-~ month_name)
+~ month_name) + scale_x_discrete(breaks = 0:2100)
 
@@
 temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") & (year >= 2002 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
-~ month_name)
+~ month_name) + scale_x_discrete(breaks = 0:2100)
 
@@
 temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") & (year > 2012 & year <= 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
-~ month_name)
+~ month_name) + scale_x_discrete(breaks = 0:2100)
 

### 2022-08-04 scripts/plots.R

@@ -281,102 +281,3 @@ pollution %>% filter((pollutant == "SO2") & (year > 2012 & year <= 2020))
 pollution %>% filter((pollutant == "NO") & (year >= 2002 & year <= 2012) & (month == "Dec" | 
                          month == "Jan" | month == "Feb" | month == "Mar"))
 
-
-
-
-
-
-
-
-
-
-
-
-
-# pollution NA months
-
-df <- pollution
-df <- df %>% drop_na(month_name)
-
-level_order <- c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug","Sep", "Oct","Nov", "Dec")
-
-df %>% filter(pollutant == "NO") %>% ggplot(
-  aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point(color = "gray") +
-  xlab("Month") + ylab("Mean Pollution (ppb)") + theme_bw()
-
-
-
-
-
-
-level_order <- c("Dec", "Jan", "Feb","Mar")
-
-temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
-                             month_name == "Mar") & (year >= 2002 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, 
-                                                                                                  color=factor(month_name, level = level_order))) + 
-  geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + 
-  theme_bw() + facet_grid(. ~ month_name)
-
-data_new <- temp_2002_2022
-data_new$month_name <- factor(data_new$month_name,
-                         levels = c("Dec", "Jan", "Feb","Mar"))
-data_new %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
-                             month_name == "Mar") & (year >= 2002 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, 
-                                                                                                  color=factor(month_name, level = level_order))) + 
-  geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + 
-  theme_bw() + facet_grid(. ~ month_name)
-
-data_new %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
[…]

### 2022-08-08 outputs/paper/paper.qmd
Updates to paper

@@
 
-As climate change worsens, the danger it presents to human health becomes more and more dangerous [@citenasaclimatewarn].
+As climate change worsens, the danger it presents to human health becomes more dangerous [@citenasaclimatewarn].
 The health impacts due to climate change are already being demonstrated in parts of Canada.
@@
 
-Urgent action is needed and the science shows that it will make a tangible difference \[\@\].
-The health impacts from climate change can be mitigated by improving efforts both in societies and health systems in Canada \[\@\].
-Changes in measures can prevent health impacts and build health systems and facilities that are more resilient to climate change \[\@\].
+Urgent action is needed and the science shows that it will make a tangible difference \[\@citeipcc\].
+The health impacts from climate change can be mitigated by improving efforts both in societies and health systems in Canada \[\@citeipcc\].
+Changes in measures can prevent health impacts and build health systems and facilities that are more resilient to climate change \[\@citeipcc\].
 
@@
 From these the city, date, year, month, day, maximum temperature, minimum temperature, and mean temperature were selected using the dplyr package [@citedplyr].
-Then any NA values from the maximum, minimum, and mean temperature columns were dropped.
 
@@
 
-Table 1 shows some of the rows of the Temperature data from 2002 to 2022.
-
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
@@
 
+Table 1 shows some rows of the Temperature data from 2002 to 2022.
 Table 2 shows some of the rows of the Pollution data from 2002 to 2020.
@@
 
-temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") & (year >= 2002 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
-~ month_name) + scale_x_discrete(breaks = 0:2100)
-
+temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") & (year >= 2002 & year < 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
+~ month_name)
 ```
@@
 
-temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | 
-                             month_name == "Mar") & (year > 2012 & year <= 2022)) %>% ggplot(aes(x=year, y=max_temp, 
-                                                                                                  color=factor(month_name, level = level_order))) + 
-  geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + 
-  theme_bw() + facet_grid(. ~ month_name) + scale_x_discrete(breaks = 0:2100)
+temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") & (year >= 2012 & year < 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
+~ month_name)
 
@@
 
-temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") & (year >= 2002 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
-~ month_name) + scale_x_discrete(breaks = 0:2100)
-
+temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") & (year >= 2002 & year < 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
+~ month_name)
 ```
@@
 
-temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") & (year > 2012 & year <= 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
-~ month_name) + scale_x_discrete(breaks = 0:2100)
-
+temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") & (year >= 2012 & year < 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
+~ month_name)
 ```
@@
 
-temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") & (year >= 2002 & year <= 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
-~ month_name) + scale_x_discrete(breaks = 0:2100)
-
+temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") & (year >= 2002 & year < 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
+~ month_name)
 ```
@@
 
-temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") & (year > 2012 & year <= 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
-~ month_name) + scale_x_discrete(breaks = 0:2100)
-
+temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") & (year >= 2012 & year < 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
+~ month_name)
 ```
@@
 
-pollution %>% filter(pollutant == "NO") %>% ggplot(
+pollution %>% filter(pollutant == "NO" & is.na(month_name) == "FALSE") %>% ggplot(
   aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point(color = "gray") +
@@
 
-pollution %>% filter(pollutant == "NO2") %>% ggplot(
+pollution %>% filter(pollutant == "NO2" & is.na(month_name) == "FALSE") %>% ggplot(
   aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point(color = "light blue") +
@@
 
-pollution %>% filter(pollutant == "O3") %>% ggplot(
+pollution %>% filter(pollutant == "O3" & is.na(month_name) == "FALSE") %>% ggplot(
   aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point(color = "light pink") +
@@
 
-pollution %>% filter(pollutant == "CO") %>% ggplot(
+pollution %>% filter(pollutant == "CO" & is.na(month_name) == "FALSE") %>% ggplot(
   aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point(color = "#9AD67F") +
@@
 
-pollution %>% filter(pollutant == "SO2") %>% ggplot(
+pollution %>% filter(pollutant == "SO2" & is.na(month_name) == "FALSE") %>% ggplot(
   aes(x = factor(month_name, level = level_order), y = mean_pollution)) + geom_point(color = "#847FD6") +

### 2022-08-08 scripts/01-data_cleaning.R

@@ -926,7 +926,7 @@ write_csv(
 pollu_temp <- dplyr::full_join(temp_2002_2022, pollution)
 
 pollu_temp <- pollu_temp %>% drop_na(pollutant)
-pollu_temp <- pollu_temp %>% drop_na(month_name)
+#pollu_temp <- pollu_temp %>% drop_na(month_name)
 pollu_temp <- pollu_temp %>% select(city, pollutant, mean_pollution, max_temp, min_temp, mean_temp, 
                                     day, month, year, month_name)
 

### 2022-08-08 scripts/plots.R

@@ -281,3 +281,4 @@ pollution %>% filter((pollutant == "SO2") & (year > 2012 & year <= 2020))
 pollution %>% filter((pollutant == "NO") & (year >= 2002 & year <= 2012) & (month == "Dec" | 
                          month == "Jan" | month == "Feb" | month == "Mar"))
 
+

### 2022-08-08 outputs/paper/paper.qmd
Updates to paper

@@
 
-Urgent action is needed and the science shows that it will make a tangible difference \[\@citeipcc\].
-The health impacts from climate change can be mitigated by improving efforts both in societies and health systems in Canada \[\@citeipcc\].
-Changes in measures can prevent health impacts and build health systems and facilities that are more resilient to climate change \[\@citeipcc\].
+Urgent action is needed and the science shows that it will make a tangible difference [@citeipcc].
+The health impacts from climate change can be mitigated by improving efforts both in societies and health systems in Canada [@citeipcc].
+Changes in measures can prevent health impacts and build health systems and facilities that are more resilient to climate change [@citeipcc].
 
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2012 for December to March", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2011 for December to March", fig.width=9, fig.height=3.2}
 
 level_order <- c("Dec", "Jan", "Feb","Mar")
+temp_2002_2022$month_name_f = factor(temp_2002_2022$month_name, levels = c("Dec", "Jan", "Feb", "Mar"))
 
 temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") & (year >= 2002 & year < 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
-~ month_name)
+~ month_name_f)
 ```
 
-Figure 1 shows an increasing pattern for the maximum temperature from year 2002 to 2012 from December to March.
+Figure 1 shows an increasing pattern for the maximum temperature from year 2002 to 2011 from December to March.
 Temperatures in March continue to be increasingly higher than any of the other months in this plot.
-It is noticeable that the temperatures for all months are the highest in 2012.
-According to NASA, the average temperatures in 2012 was warmer than the mid-20th century baseline and is considered to be an outlying of seasonal extremes that are warmer [@citenasaclimatewarn].
-The lowest maximum temperature in January is higher than other years in 2012.
+The lowest maximum temperature in January is higher than other years in 2011.
 The plot shows that the average maximum temperature from December to March has been between -19.2 degrees Celsius in February of 2015 and 24.3 degrees Celsius in March of 2012.
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2013 to 2022 for December to March", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2012 to 2022 for December to March", fig.width=9, fig.height=3.2}
 
 temp_2002_2022 %>% filter((month_name == "Dec" | month_name == "Jan"| month_name == "Feb" | month_name == "Mar") & (year >= 2012 & year < 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
-~ month_name)
+~ month_name_f)
 
@@
 Figure 2 shows an increasing pattern for the maximum temperature from year 2013 to 2022 from December to March.
+It is noticeable that the temperatures for all months are the highest in 2012.
+According to NASA, the average temperatures in 2012 was warmer than the mid-20th century baseline and is considered to be an outlying of seasonal extremes that are warmer [@citenasaclimatewarn].
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2012 for April to July", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2011 for April to July", fig.width=9, fig.height=3.2}
 
 level_order <- c("Apr", "May", "Jun", "Jul")
+temp_2002_2022$month_name_f = factor(temp_2002_2022$month_name, levels = c("Apr", "May", "Jun", "Jul"))
 
 temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") & (year >= 2002 & year < 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
-~ month_name)
+~ month_name_f)
 ```
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2013 to 2022 for April to July", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2012 to 2022 for April to July", fig.width=9, fig.height=3.2}
 
@@
 temp_2002_2022 %>% filter((month_name == "Apr" | month_name == "May" | month_name == "Jun" | month_name == "Jul") & (year >= 2012 & year < 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
-~ month_name)
+~ month_name_f)
 ```
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2012 for August to November", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2002 to 2011 for August to November", fig.width=9, fig.height=3.2}
 
 level_order <- c("Aug","Sep", "Oct","Nov")
+temp_2002_2022$month_name_f = factor(temp_2002_2022$month_name, levels = c("Aug","Sep", "Oct","Nov"))
 
 temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") & (year >= 2002 & year < 2012)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
-~ month_name)
+~ month_name_f)
 ```
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2013 to 2022 for August to November", fig.width=9, fig.height=3.2}
+```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Temperature 2012 to 2022 for August to November", fig.width=9, fig.height=3.2}
 
@@
 temp_2002_2022 %>% filter((month_name == "Aug" | month_name == "Sep" | month_name == "Oct" | month_name == "Nov") & (year >= 2012 & year < 2022)) %>% ggplot(aes(x=year, y=max_temp, color=factor(month_name, level = level_order))) + geom_point(alpha = 0.4) + labs(color="Month") + xlab("Year") + ylab("Maximum Temperature") + theme_bw() + facet_grid(.
-~ month_name)
+~ month_name_f)
 ```
